### 2021-06-04 survey/code/deconv_corr.Rmd
Deconv correlation experiments

---
title: Correlation analysis to inferred infection rates (deconvolved case rates)
---

```{r, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, autodep = TRUE, cache.comments = TRUE, 
                      message = FALSE, warning = FALSE)
```

```{r, messages = FALSE}
library(covidcast)
library(dplyr)
library(ggplot2)

cases <- covidcast_signal(data_source = "usa-facts",
                           signal = "confirmed_7dav_incidence_prop",
                           start_day = "2020-04-15",
                           end_day = "2021-04-15", 
                           geo_type = "state")
survey <- covidcast_signal(data_source = "fb-survey",
                           signal = "smoothed_whh_cmnty_cli",
                           start_day = "2020-04-15",
                           end_day = "2021-04-15", 
                           geo_type = "state")

deconv <- read.csv("deconv_data.csv")
deconv <- as.covidcast_signal(deconv, signal = deconv$signal[1])
deconv$time_value <- as.Date(deconv$time_value)
```

## Geo-wise correlations (sliced by time)

```{r}
# Compute correlation per time, over all counties
cor_geowise1 <- covidcast_cor(deconv, survey, by = "time_value")
cor_geowise2 <- covidcast_cor(cases, survey, by = "time_value")
cor_geowise <- rbind(cor_geowise1, cor_geowise2)
cor_geowise$signal = as.factor(
  c(rep("deconv-survey", nrow(cor_geowise1)),
    rep("cases-survey", nrow(cor_geowise2))))

ggplot(cor_geowise, aes(x = time_value, y = value)) +
  geom_line(aes(color = signal)) +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") + 
  labs(x = "Date", y = "Correlation") + theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
#ggsave("../fig/deconv_geowise_corr.pdf", width = 5, height = 3.5)
```

## Time-wise correlations (sliced by location)

```{r}
# Compute correlation per location, over all time
cor_timewise1 <- covidcast_cor(deconv, survey, by = "geo_value")
cor_timewise2 <- covidcast_cor(cases, survey, by = "geo_value")
cor_timewise <- rbind(cor_timewise1, cor_timewise2)
cor_timewise$signal = as.factor(
  c(rep("deconv-survey", nrow(cor_timewise1)),
    rep("cases-survey", nrow(cor_timewise2))))

ggplot(cor_timewise, aes(value)) +
  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
  labs(x = "Correlation", y = "Density") + theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
#ggsave("../fig/deconv_timewise_corr.pdf", width = 5, height = 3.5)
```

## More systematic lag analysis

```{r}
# Loop over values for dt, and compute correlations per state
dt_vec <- (-20):10
cor_timewise_list1 <- vector("list", length(dt_vec))
for (i in 1:length(dt_vec)) {
  cor_timewise_list1[[i]] <- covidcast_cor(deconv, survey, dt_x = dt_vec[i],
                               by = "geo_value")
  cor_timewise_list1[[i]]$dt <- dt_vec[i]
}

# Stack into one big data frame, and then plot the median correlation by dt
cor_timewise_dt1 <- do.call(rbind, cor_timewise_list1)
cor_timewise_dt1 %>%
  group_by(dt) %>%
  summarize(value = mean(value, na.rm = TRUE), .groups = "drop_last") %>%
  ggplot(aes(x = dt, y = value)) + geom_line() + geom_point() +
  labs(x = "dt", y = "Correlation") + theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r}
# Loop over values for dt, and compute correlations per state
cor_timewise_list2 <- vector("list", length(dt_vec))
for (i in 1:length(dt_vec)) {
  cor_timewise_list2[[i]] <- covidcast_cor(cases, survey, dt_x = dt_vec[i],
                               by = "geo_value")
  cor_timewise_list2[[i]]$dt <- dt_vec[i]
}

# Stack into one big data frame, and then plot the median correlation by dt
cor_timewise_dt2 <- do.call(rbind, cor_timewise_list2)
cor_timewise_dt2 %>%
  group_by(dt) %>%
  summarize(value = mean(value, na.rm = TRUE), .groups = "drop_last") %>%
  ggplot(aes(x = dt, y = value)) + geom_line() + geom_point() +
  labs(x = "dt", y = "Correlation") + theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
```