### 2021-06-18 forecast/paper/pnas-suppl-template.tex
rebuild leadingness; reflow supplement for ease of edits

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}

\input{../../common/prelim.tex}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

$for(header-includes)$
$header-includes$
$endfor$

\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
  Prediction?}

\author{Daniel J.\ McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries, Sangwon Hyun, Natalia L.\ Oliveira, James Sharpnack, Jingjing Tan, Robert Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J.\ Tibshirani}
\correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}

\begin{document}

\maketitle

% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

$body$

\bibliography{$for(bibliography)$$bibliography$$sep$,$endfor$}

\end{document}