### 2021-06-19 forecast/paper/paper.tex
Copyedit abstract

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\input{../../common/prelim.tex}

\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
  Prediction?}

% Use letters for affiliations, numbers to show equal authorship (if applicable)
% and to indicate the corresponding author

% Writing group
\author[a,1]{Daniel J.\ McDonald}
\author[b,2]{Jacob Bien}
\author[c,2]{Alden Green}
\author[c,d,2]{Addison Hu}

% Middle authors
\author[d]{Nat DeFries}
\author[b]{Sangwon Hyun}
\author[c,d]{Natalia L.\ Oliveira}
\author[e]{James Sharpnack}
\author[f]{Jingjing Tang}
\author[g]{Robert Tibshirani}
\author[c]{Valerie Ventura}
\author[c,d]{Larry Wasserman}

% Senior authors
\author[c,d]{Ryan J.\ Tibshirani}

\affil[a]{Department of Statistics, University of British Columbia}
\affil[b]{Data Sciences and Operations, University of Southern California}
\affil[c]{Department of Statistics \& Data Science, Carnegie Mellon University}
\affil[d]{Machine Learning Department, Carnegie Mellon University}
\affil[e]{Department of Statistics, University of California, Davis}
\affil[f]{Computational Biology Department, Carnegie Mellon University}
\affil[g]{Department of Statistics, Stanford University}

% Please give the surname of the lead author for the running footer
\leadauthor{McDonald}

% Please add a significance statement to explain the relevance of your work
\significancestatement{Forecasting is a critical element of the public health
  response to any fast-moving epidemic or pandemic.  Predicting the future
  spread of some disease process, based on disease activity in the recent
  past---say, using what is known as autoregressive modeling in the time series
  literature---is a basic and widely-used method.  In this paper we study
  whether including auxiliary indicators as additional features in such an
  autoregressive model can be useful in COVID-19 case forecasting and hotspot
  prediction.  We find that indicators based on de-identified medical insurance 
  claims, self-reported symptoms through online surveys, and COVID-related
  internet search activity all provide a nontrivial boost in accuracy, though
  their utility generally depends the pandemic's local dynamics (upswing,
  downswing, or flat period) at prediction time.
}

% Please include corresponding author, author contribution and author
% declaration information 
\authorcontributions{Please provide details of author contributions here.}
\authordeclaration{Please declare any competing interests here.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be
  addressed. E-mail: daniel@stat.ubc.ca} 
\equalauthors{\textsuperscript{2}J.B., A.G., and A.H.\ contributed equally to
  this work.}

% At least three keywords are required at submission. Please provide three to
% five keywords, separated by the pipe symbol. 
\keywords{COVID-19 $|$ forecasting $|$ hotspot prediction $|$ time series | 
  digital surveillance}

\begin{abstract}
  Reliable, short-term forecasts of traditional public health reporting streams
  (such as cases, hospitalizations, and deaths) are a key ingredient in
  effective public health decision-making during a pandemic. Since April 2020,
  our research group has worked with data partners to collect, curate, and make
  publicly available numerous real-time COVID-19 indicators, providing multiple
  views of pandemic activity.  This paper studies the utility of these
  indicators from a forecasting perspective. We focus on five indicators,
  derived from medical insurance claims data, web search queries, and online
  survey responses. For each indicator, we ask whether its inclusion in a simple
  model leads to improved predictive accuracy relative to a similar model
  excluding it. We consider both probabilistic forecasting of confirmed COVID-19 
  case rates and binary prediction of case ``hotspots''.  Since the values of 
  indicators (and case rates) are commonly revised over time, we take special
  care to ensure that the data provided to a forecaster is the version that
  would have been available at the time the forecast was made.   
  % We demonstrate how less careful backtesting can lead to misleading
  % retrospective evaluations.   
  Our  analysis shows that consistent but modest gains in predictive accuracy 
  are obtained by using these indicators, and furthermore, these gains are
  related to periods in which the auxiliary indicators behave as ``leading
  indicators'' of case rates.     
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, formulating public policy responses,
and anticipating future public health resource needs.  In a companion paper, we
describe our research group's (Delphi's) efforts in curating and maintaining a 
database of real-time indicators that track COVID-19 activity and other relevant
phenomena. The signals (a term we use synonomously with ``indicators'') in this
database are accessible through the COVIDcast API \cite{CovidcastAPI}, with
associated R \cite{CovidcastR} and Python \cite{CovidcastPy} packages for
convenient data fetching and processing tools. In the current paper, we aim to
quantify the utility provided by a core set of these indicators for two
fundamental prediction tasks: probabilistic forecasting of COVID-19 case 
rates and prediction of future COVID-19 case hotspots (defined by the event that
a relative increase in COVID-19 cases exceeds a certain threshold).

 At the outset, we should be clear that our intent in this paper is \textit{not}
to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
Instead, our purpose here is to provide a rigorous, quantitative assessment of
the utility that several auxiliary indicators---such as those derived from
internet surveys or medical insurance claims---provide in tasks that involve
predicting future trends in confirmed COVID-19 cases. To assess such utility in
as simple terms as possible, we center our study in the framework of a basic
autoregressive model (in which COVID cases in the near future are predicted from  
a linear combination of COVID cases in the near past), and ask whether the 
inclusion of an auxiliary indicator as an additional feature in such a model
improves its predictions.

While forecasting carries a rich literature that offers a wide range of
techniques, see e.g., \cite{Hyndman:2018}, we purposely constrain ourselves to
very simple models, avoiding common enhancements such as order selection,
correction of outliers/anomalies in the data, and inclusion of regularization or
nonlinearities. That said, analyses of forecasts submitted to the COVID-19
Forecast Hub \cite{ForecastHub} by a large community of modelers have shown that
simple, robust models have consistently been among the best-performing over the
pandemic \cite{Cramer:2021}, including time series models similar to those we
consider in what follows.

In our companion paper, we analyze correlations between various indicators and   
COVID case rates. These correlations are natural summaries of the
contemporaneous association between an indicator and COVID cases, but they fall 
short of delivering a satisfactory answer to the question that motivates the
current article: is the information contained in an indicator demonstrably
useful for the prediction tasks we care about? For such a question, lagged
correlations (e.g., measuring the correlation between an indicator and COVID
case rates several days in the future) move us in the right direction, but still
do not move us all the way there. The question about \textit{utility for 
  prediction} is focused on a much higher standard than simply asking 
about correlations; to be useful in forecast or hotspot models, an indicator
must provide relevant information that is not otherwise contained in past values
of the case rate series itself. We will assess this in the most direct way
possible: by inspecting the difference in predictive performance of simple
autoregressive models trained with and without access to past values of a
particular indicator.

There is another, more subtle issue in evaluating predictive utility that 
deserves explicit mention, as it will play a key role in our analysis.
Signals computed from surveillance streams will often be subject to  
latency and/or revision. For example, a signal based on aggregated medical
insurance claims may be available after just a few days, but it can then be
substantially revised over the next several weeks as additional claims are
submitted and/or processed late. Correlations between such a signal and case
rates calculated ``after the fact'' (i.e., computed retrospectively, using the
finalized values of this signal) will not deliver an honest answer to the    
question of whether this signal would have been useful in real time. Instead,
we build predictive models using only the data that would have been available
\textit{as of} the prediction date, and compare the ensuing predictions in terms 
of accuracy. To do so, we leverage Delphi's \texttt{evalcast} R package 
\cite{EvalcastR}, which plugs into the COVIDcast API's data versioning system,  
and facilitates honest backtesting.

Finally, it is worth noting that examining the importance of additional features
for prediction is a core question in inferential statistics and econometrics,
with work dating back to at least \cite{Granger:1969}. Still today, drawing
rigorous inference based on predictions, without (or with lean) assumptions, is
an active field of research from both the applied and theoretical angles;
see, e.g., \cite{Diebold:2002, McCraken:2007, Diebold:2015, Stokes:2017,
  Lei:2018, Rinaldo:2019, Williamsom:2020, Zhang:2020, Dai:2021, Fryer:2021}.
Our take on this problem is in line with much of this literature; however, in
order to avoid making any explicit assumptions, we do not attempt to make formal 
significance statements, and instead, broadly examine the stability of our
conclusions with respect to numerous modes of analysis.

\section{Methods}

\subsection{Signals and Locations}

We consider prediction of future COVID-19 case rates or case hotspots (to be
defined precisely shortly).  By case rate, we mean the case count per 100,000
people (the standard in epidemiology).  We use reported case data aggregated by 
JHU CSSE \cite{Dong:2020}, which, like the auxiliary indicators that we use to supplement the
basic purely autoregressive models, is accessible through the COVIDcast API
\cite{CovidcastAPI}.

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey COVID-like illness in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community that is sick, based on Delphi's surveys of
  Facebook users, called the COVID Trends and Impact Survey. 
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google Search Trends for anosmia and ageusia (Google-AA): A measure of 
  Google search volume for queries relating to anosmia or
  ageusia (loss of smell or taste), based on Google's Search Trends data set. 
\end{itemize}
In short, we choose these indicators because, conceptually speaking, they
measure aspects of an individual's disease progression that would plausibly
precede the occurence of (at worst, co-occur  with) the report of a positive
COVID-19 test, through standard public health reporting streams.

For more details on the five indicators (including how these are precisely
computed from the underlying data streams) we refer to
\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
which documents all of the signals in the COVIDcast API, and our companion paper
on the API and database. For CTIS in particular, we refer to our companion paper
on this survey. For the Google Search Trends data set, see
\cite{GoogleSymptoms}, and see \cite{Klopfen:2020, Vaira:2020} for justification of the
relevance of anosmia or ageusia to COVID-19 infection.

As for geographic resolution, we consider the prediction of COVID-19 case rates
and hotspots aggregated at the level of an individual \textit{hospital referral
  region} (HRR). HRRs correspond to groups of counties in the United States
within the same hospital referral system. The Dartmouth Atlas of Healthcare 
Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
characteristics. They are contiguous regions such that most of the hospital
services for the underlying population are performed by hospitals within the
region. Each HRR also contains at least one city where major procedures
(cardiovascular or neurological) are performed. The smallest HRR has a
population of about 125,000. While some are quite large (such as the one
containing Los Angeles, which has more than 10 million people), generally HRRs 
are much more homogenous in size than the (approximately) 3200 counties,  
and they serve as a nice middle ground in between counties and states.

HRRs, by their definition, would be most relevant for forecasting hospital
demand.  We have chosen to focus on cases (forecasting and predicting
hotspots) at the HRR level because the indicators considered should be more 
useful in predicting case activity rather than hospital demand. The \attn{former?} is
intuitively more contemporaneous to the events that are measured by
the given five indicators. Predicting case rates (and hotspots) at the HRR level
is still a reasonable goal in its own right; and moreover, it could be used to
feed predicted case information into downstream hospitalization models.

\subsection{Vintage Training Data}

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 28, 2020, we train models 
using only the data that would have been available to us \textit{as of}
September 28 (imagine that we can ``rewind'' the clock to September 28 and query
the COVIDcast API to get the latest data it would have had available at that
point in time.)  This is made possible by the COVIDcast API's comprehensive data   
versioning system (described in more detail in our companion paper).  We also
use the \texttt{evalcast} R package \cite{EvalcastR}, which streamlines the
process of training arbitrary prediction models over a sequence of prediction
dates, by constructing the proper sequence of vintage training data sets.

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
  \caption{This figure displays revision behavior for two signals in the HRR
    containing New York 
    City.  Each colored line corresponds to the data as reported on a particular 
    date (\textit{as of} dates varying from September 28 through October
    19). The left panel shows the DV-CLI signal, which was regularly revised
    throughout the period, though the effects fade as we look further back in   
    time. In contrast, the right panel shows case rates reported by JHU CSSE,
    which remain ``as reported'' on September 28, with a huge spike towards the 
    end of this period, until a major correction is made on October 19, which
    brings down this spike and affects all prior data as well.}
  \label{fig:vintage}
\end{figure}

Vintage training data means different things, in practice, for different
signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
DV-CLI, are most often 3-5 days latent, and subject to a considerable but
regular degree of revision or ``backfill'' after their initial publication date.
The survey-based signal, CTIS-CLI-in-community, is 2 days latent, and rarely
undergoes any revision at all.  The target variable itself, reported COVID-19
case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
after initial publication.  Compared to the revision
pattern in the medical claims signals, which are much more systematic in nature,
revisions in case reports can be highly erratic. Big spikes or other anomalies
can occur in the data as reporting backlogs are cleared, changes in case
definitions are made, etc. Groups like JHU CSSE then work tirelessly to correct
such anomalies after first publication (e.g., they will attempt to
back-distribute a spike when a reporting backlog is cleared, by working with a
local authority to figure out how this should best be done), which can result
in very nontrivial revisions.  See
Figure~\ref{fig:vintage} for an example.

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then.  Furthermore, the
latency of the signal was always at least one week through 2020.  However, this
signal is never revised after initial publication (confirmed via personal
communication with the Google team that produces this signal) and furthermore
the latency of the signal is not an unavoidable property of the data type, so we
simply use finalized signal values, with zero latency, in our analysis.

\subsection{Analysis Tasks}

\begin{table*}[t]
\centering
\caption{Summary of forecasting and hotspot prediction tasks considered in
  this paper.}
\begin{tabular}{l p{2.75in} p{2.75in}}
  \toprule
  & \textbf{Forecasting} & \textbf{Hotspot prediction} \\
  \midrule
  Response variable & $Y_{\ell,t}$ (7-day trailing average of COVID-19 case 
incidence rates, per location $\ell$ and time $t$) & $Z_{\ell,t} =
\indicator{Y_{\ell,t} \geq 1.25 \cdot Y_{\ell,t-7}}$ (indicator that 
$Y_{\ell,t}$ grows by more than 25\% relative to the preceding week) \\ 
  Geographic resolution & Hospital referral region (HRR) & Hospital referral
region (HRR) \\ 
  Forecast period & June 9--December 31, 2020 & June 16--December 31, 2020 \\  
  Model type & Quantile regression & Logistic regression \\
  Evaluation metric & Weighted interval score (WIS) & Area under curve (AUC) \\
  \bottomrule
\end{tabular}
\label{tab:analysis_tasks}
\end{table*}

To fix notation, let $Y_{\ell,t}$ denote the 7-day trailing average of COVID-19
case incidence rates in location (HRR) $\ell$ and at time (day) $t$.  To be
clear, this is the number of new daily reported cases per 100,000
people, averaged over the 7-day period $t-6, \ldots, t$. The first task we
consider---\textit{forecasting}---is to predict $Y_{\ell,t+a}$ for each
``ahead'' value $a=7,\ldots,21$.  The second task---\textit{hotspot
  prediction}---is to predict a binary variable defined in terms of the relative
change of $Y_{\ell,t+a}$ (relative to its value one week prior,
$Y_{\ell,t+a-7}$), again for each $a=7,\ldots,21$.

Why do we define the response variables via 7-day averaging? The short answer 
is robustness: averaging stabilizes the case time series, and accounts for 
uninteresting artifacts like day-of-the-week effects in the series.  Note
that we can equivalently view this (equivalent up to a constant
factor) as predicting the HRR-level COVID-19 case incidence rate \textit{summed}
over some 7-day period in the future, and predicting a binary variable derived
from this.

In what follows, we cover more details on our two analysis
tasks. Table~\ref{tab:analysis_tasks} presents a summary.

\paragraph{Dynamic Re-Training}

For each prediction date $t$, we use a 21-day trailing window of data to train
our forecast or hotspot prediction models (so, e.g., the trained models will
differ from those at prediction date $t-1$).  This is done to account for
(potential) nonstationarity.  For simplicity, the forecasting and hotspot
prediction models are always trained on data across all HRRs (i.e., the
coefficients in the models do not account for location-specific effects).

\paragraph{Prediction Period}

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data and lagged values
of signals for features, as we will detail shortly) June 9, 2020 for forecasting,
and June 16, 2020 for hotspot prediction. (The bottleneck here is the
CTIS-CLI-in-community signal, which does not exist before early April 2020, when
the survey was first launched). The end date was chosen again with a
consideration to align both tasks as best as possible, and because few hotspots
exist post December 31, 2020, due to the general and gradual decline of the
pandemic in 2021.

\paragraph{Forecasting Models}

Recall $Y_{\ell,t}$ denotes the 7-day trailing average of COVID-19 case
incidence rates in location $\ell$ and at time $t$.  Separately for each
$a=7,\ldots,21$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
simple probabilistic forecasting model of the form:     
\begin{equation}
\label{eq:forecast_ar}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ s \leq t)  
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
\end{equation}
This model uses current case rates, and the case rates 7 and 14 days ago, in
order to predict (the quantiles of) case rates in the future.  We consider a
total of 7 quantile levels (chosen in accordance with the county-level quantile
levels suggested by the COVID-19 Forecast Hub),      
\begin{equation}
\label{eq:quantile_levels}
\tau \in \{0.025,\ 0.1,\ 0.25,\ 0.5,\ 0.75,\ 0.9,\ 0.975 \}.
\end{equation}
We fit \eqref{eq:forecast_ar} using \textit{quantile regression}
\cite{Koenker:1978, Koenker:2005, Koenker:2006} separately for each $\tau$, 
using data from all 306 HRRs, and within each HRR, using the most recent 21 days
of training data.  This gives us 6,426 training samples for each quantile 
regression problem.

In addition to this pure autoregressive model, we also consider five
probabilistic forecasting models of the form:  
\begin{multline}
\label{eq:forecast_ar_x}
\Quantile_\tau(Y_{\ell,t+a} \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t)  
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
\end{multline}
where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
CHNG-COVID, CTIS-CLI-in-community, DV-CLI, or Google-AA---at location $\ell$ and
time $t$. Note that we apply the same lags (current value, along with the values
7 and 14 days ago) for the auxiliary indicators as we do for the case
rates. Training then proceeds just as before: we use the same 7 quantile levels
in \eqref{eq:quantile_levels}, and fit quantile regression separately for each 
level $\tau$, using data from all 306 HRRs and a trailing window of 21 days of
training data.

At prediction time, in order to avoid crossing violations (that is,
for two levels $\tau' > \tau$,
the predicted quantile at level $\tau$ exceeds the predicted quantile at level
$\tau'$), we apply a simple post-hoc sorting.  See Figure~\ref{fig:trajectory}
for an example forecast.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
\caption{This figure displays a forecast from an autoregressive model for the
  HRR containing
  New York City made  
  on October 15 (vertical line).  The fan displays 50\%, 80\% and 95\% intervals
  while the orange curve shows the median forecast. The black curve represents
  ``finalized'' data, as reported in May 2021.} 
\label{fig:trajectory}
\end{figure}

\paragraph{Hotspot Prediction Models}

Define the binary indicator:
$$
Z_{\ell,t} = \indicator{Y^\Delta_{\ell,t} \geq 0.25},
$$
where we use the notation \smash{$Y^\Delta_{\ell,t} = (Y_{\ell,t} - Y_{\ell,
    t-7})/(Y_{\ell,t-7})$}. In other words, $Z_{\ell,t}=1$ if the number of
newly reported cases over the past 7 days has increased by at least 25\%
compared to the preceding week.  When this occurs, we say location $\ell$ is a
\textit{hotspot} at time $t$.  Empirically, this rule labels about 27\% of
location-time pairs as hotspots, during the prediction period (June 16--December
31, 2020).

We treat hotspot prediction as a binary classification problem and use a setup 
altogether quite similar to the forecasting setup described previously.
Separately for each $a=7,\ldots,21$, to predict $Z_{\ell,t+a}$, we consider a
simple logistic model:
\begin{equation}
\label{eq:hotspot_ar}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ s \leq t) \big) 
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
\end{equation}
where $\logit(p) = \log(p/(1-p))$, the log-odds of $p$.

In addition to this pure autoregressive model, we also consider five
logistic models of the form:  
\begin{multline}
\label{eq:hotspot_ar_x}
\logit \big( \P(Z_{\ell,t+a} = 1 \,|\, Y_{\ell,s},\ X_{\ell,s},\ s \leq t) \big)
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit the above models,
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x}, using logistic regression,   
pooling all 306 HRRs and using a 21-day trailing window for the training data.

An important detail is that in hotspot prediction we remove all data from
training and evaluation where, on average, fewer than 30 cases (this refers to a
count, not a rate) are observed over the preceding 7 days. This avoids having to
make arbitrary calls for a hotspot (or lack thereof) based on small counts.

\subsection{Evaluation Metrics}

For forecasting, we evaluate the probabilistic forecasts produced by the
quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using 
\textit{weighted interval score} (WIS), a quantile-based scoring rule; see e.g.,
\cite{Gneiting:2007}.  WIS is a proper score, meaning that its expectation
is minimized by the population quantiles of the target variable.  The use of WIS
in COVID-19 forecast scoring is discussed in \cite{Bracher:2021}; WIS is also
the main evaluation metric used in the COVID-19 Forecast Hub.

WIS is typically defined for quantile-based forecasts where the quantile levels
are symmetric around 0.5.  This is the case for our choice in
\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the case of
symmetric quantile levels, this is equivalent to a collection of central
prediction intervals $(\ell_\alpha, u_\alpha)$, parametrized by an exclusion
probability $\alpha$. The WIS of the forecaster $F$, evaluated at the target
variable $Y$, is defined by:
\begin{equation}
\label{eq:wis_intervals}
\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
\end{equation}
where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
smallest distance between $a$ and an element of $S$).  Note that, corresponding 
to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
\{0.05,\ 0.2,\ 0.5,\ 1\}$, resulting in 4 terms in the above sum.  By 
straightforward algebra, it is not hard to see WIS has an alternative 
representation in terms of the predicted quantiles themselves:
\begin{equation}
\label{eq:wis_quantiles}
\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
\end{equation}
where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
presentation, because the score nicely decouples into a ``sharpness'' component
(first term in each summand) and an ``under/overprediction'' component (second term
in each summand).  The second form given in \eqref{eq:wis_quantiles} is
especially noteworthy in our current study because it reveals WIS to be the same as the
quantile regression loss that we use to train our forecasting models (i.e., our
models are fit to optimize WIS averaged over the training set).

For hotspot prediction, we evaluate the probabilistic classifiers produced by
the logistic models in \eqref{eq:hotspot_ar} and \eqref{eq:hotspot_ar_x} using the 
area under the curve (AUC) of their true positive versus false positive rate
curve (which is traced out by varying the discrimination threshold).

The primary aggregation scheme that we will use in model evaluation and
comparisons will be to average WIS per forecaster at ahead value $a$ over all
forecast dates $t$ and locations $\ell$; and similarly, to compute AUC per
classifier at ahead value $a$ over all forecast dates $t$ and locations
$\ell$.

\subsection{Other Considerations}

\paragraph{Missing Data Imputation}

Over the prediction period, all auxiliary indicators are available (in the
proper vintage sense) for all locations and prediction times, except for the
Google-AA signal, which is only observed for an average of 105 (of 306) HRRs.
Such missingness occurs because the COVID-19 search trends data is
constructed using differential privacy methods \cite{Bavadekar:2020}, and a 
missing signal value means that the level of noise added in the differential
privacy mechanism is high compared to the underyling search count.  In other
words, values of the Google-AA signal are clearly \textit{not} missing at
random.  It seems most appropriate to impute missing values by zero, and this is
what we do in our analysis.

\paragraph{Backfill and Nowcasting}

As described previously, the auxiliary indicators defined in terms of medical
claims (CHNG-CLI, CHNG-COVID, and DV-CLI) undergo a significant and systematic
pattern of revision, or ``backfill'', after their initial publication.  Given
their somewhat statistically-regular backfill profiles, it would be reasonable
to attempt to estimate their finalized values based on vintage data---a problem
we refer to as \textit{nowcasting}---as a pre-processing step before using them
as features in the models in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}. 
Nowcasting is itself a highly nontrivial modeling problem, and we do not attempt
it in this paper (it is a topic of ongoing work in our research group), but we
note that nowcasting would likely improve the performance of the models
involving claims-based signals in particular.

\paragraph{Spatial Heterogeneity}

Some signals have a significant amount of spatial heterogeneity, by which we
mean their values across different geographic locations are not comparable.
This is the case for the Google-AA signal (due to the way in which the
underlying search trends time series is self-normalized, see
\cite{GoogleSymptoms}) and the claims-based signals (see the discussion in our
companion paper on the API and database).  Such spatial heterogeneity likely
hurts the performance of the predictive models that rely on these signals,
because we train the models on data pooled over all locations.  In the current
paper, we do not attempt to address this issue (it is again a topic of ongoing
work in our group), and we simply note that location-specific effects (or
pre-processing to remove spatial bias) would likely improve the performance of
the models involving Google-AA and the claims-based indicators.

\section{Results}

Here, and in the
following subsections, we will use ``AR'' to refer to the pure autoregressive
model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot prediction,
\eqref{eq:hotspot_ar} (the reference to the prediction task should always be
clear from the context). We will also use the name of an auxiliary
indicator---namely ``CHNG-CLI'',  ``CHNG-COVID'', ``CTIS-CLI-in-community'',
``DV-CLI'', or ``Google-AA''---interchangeably with the model in forecasting, 
\eqref{eq:forecast_ar_x}, or hotspot prediction, \eqref{eq:hotspot_ar_x}, that
uses this particular indicator as a feature (the meaning should be clear from
the context).  So, for example, the CHNG-CLI model in forecasting is the one in  
\eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI 
indicator at location $\ell$ and time $t$.
% fits the coefficients based on the training data, and makes its predictions by
% setting $X_{\ell,t}$ to the appropriate CHNG-CLI values.  
Finally, we use the term ``indicator model'' to refer to any one of the ten
models of the form \eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five from
each of the forecasting and hotspot prediction tasks). We begin with a summary
of the high-level conclusions.

\begin{itemize}
\item Stratifying predictions by the ahead value ($a=7,\ldots,21$), and
  aggregating results over the prediction period (early June through end of
  December 2020), we find that each of the indicator models generally gives a
  boost in predictive accuracy over the AR model, in both the forecasting and
  hotspot prediction tasks.  The gains in accuracy generally attenuate as the
  ahead value grows.

\item In the same aggregate view, CHNG-COVID and DV-CLI offer the biggest 
  gains in both forecasting and hotspot prediction.  CHNG-CLI is
  inconsistent: it provides a big gain in hotspot prediction, but little gain 
  in forecasting (it seems to be hurt by a notable lack of robustness, due
  to backfill).  CTIS-CLI-in-community and Google-AA each provide decent 
  gains in forecasting and hotspot prediction.  The former's performance in 
  forecasting is notable in that it clearly improves on AR even at the largest
  ahead values.

\item In a more detailed analysis of forecasting performance, we find that the
  indicator models tend to be
  better than AR when case rates
  are flat or decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community),
  but they are
  worse than AR when case rates are increasing (this
  is most notable in CHNG-CLI and DV-CLI).
  More rarely does an indicator model tend to beat AR when case rates are 
  increasing, but there appears to be some evidence of this for the Google-AA
  model. 
  % In an increasing period, when an indicator model is worse than AR, it tends
  % to underpredict relative to AR (its median forecast is lower).  In a flat
  % or decreasing period, when an indicator model is better than AR, AR tends
  % to overpredict relative to it (the median AR forecast is higher).

\item In this same analysis, when an indicator model performs better than AR in
  a decreasing period, this tends to co-occur with instances in which the
  indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
  given location, its behavior mimics that of future case rate behavior).  On
  the other hand, if an an indicator model does better in periods of
  increase, or worse in periods of increase or decrease, its performance is
  unrelated to leadingness.   
\end{itemize}

Finally, to quantify the importance of training and making predictions using
proper vintage data, we ran a parallel set of forecasting and hotspot prediction 
experiments using finalized data. The results, given in the supplement, show
that training and making predictions on finalized data can result in overly
optimistic estimates of true test-time performance (up to 10\% better in terms
of average WIS or AUC). Furthermore, since indicators can have greatly different
backfill profiles, the use of finalized data in retrospective evaluations
changes the relative ranking of models.  For example, CHNG-CLI and DV-CLI,
when trained on finalized data, perform very similarly for the forecasting task.
This makes sense since they are both claims-based indicators supposedly measuring
the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage
data, reflecting its has a less severe backfill profile.

Code to reproduce all results (which uses the \texttt{evalcast} R package) can 
be found at
\url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/forecast/code}.

\subsection{Aggregate Results by Ahead Value}

Figure~\ref{fig:forecast} (left panel) displays evaluation results for
forecasting, stratified by ahead value and averaged over all HRRs and forecast
dates.  Shown is the average WIS for each forecast model divided by that from a
baseline model, which is basically a flat-line forecaster (its median forecast
for $Y_{\ell,t+a}$ is always $Y_{\ell,t}$, with predicted quantiles defined
around this based on historical variation). This is the same baseline model as
in the COVID-19 Forecast Hub. Here, we use the baseline model in order to scale
mean WIS so that it gets put on an interpretable, unitless scale.  In the
figure, we can see that all curves are below 1, which means (smaller WIS is
better) that all of the models, including AR, outperform the baseline on average
over the forecasting period.  On the other hand, the models deliver at best an
improvement of about 20\% in average WIS over the baseline model, with this gap
narrowing to about 10\% at the largest ahead values, illustrating the difficulty 
of the forecasting problem.

We can also see from the figure that CHNG-COVID and DV-CLI offer the biggest
gains over AR at small ahead values, followed by CTIS-CLI-in-community and
Google-AA, with the former providing the biggest gains at large ahead values.
The CHNG-CLI model performs basically the same as AR.  This is likely due to the
fact that CHNG-CLI suffers from spatially heterogeneous volatility due to
backfill.  The evidence for this explanation is twofold: (1) the CHNG-CLI model benefits from a
more robust method of aggregating WIS (geometric mean; shown in the supplement);
and (2) when we train and make predictions on finalized data, it handily beats
AR, on par with the best-performing models (also shown in the 
supplement).
% TODO note what we said here about the supplement

Figure~\ref{fig:forecast} (right panel) displays the results for hotspot
prediction, again stratified by ahead value and averaged over all HRRs and
prediction dates.  We can see many similarities to the forecasting results (now
larger AUC is better). For example, CHNG-COVID and DV-CLI offer the biggest
improvement over AR, and all models, including AR, degrade in performance towards
the baseline (in this context, a classifier based on random guessing, which
achieves an AUC of 0.5) as the ahead values grow, illustrating the difficulty of
the hotspot prediction problem. A clear difference, however, is that the CHNG-CLI
model performs quite well in hotspot prediction, close to the best-performing
indicator models for many of the ahead values. This may be because volatility in
the CHNG-CLI indicator plays less of a role in the associated logistic model's
predicted probabilities (in general, a sigmoid function can absorb a lot of the
variability in its input).

\begin{figure*}[t]
  \includegraphics[width=\textwidth]{fig/fcast-hot-combo-1.pdf}
  \caption{Main results for both tasks. Left: average WIS for each forecast
    model, over all forecast dates 
    and all HRRs, divided by the average WIS achieved by a baseline model (a
    probabilistic version of the flat-line forecaster).  Right: area under the
    curve for each hotspot prediction model, calculated over all prediction
    dates and all HRRs.  Here and in all figures we abbreviate
    CTIS-CLI-in-community by CTIS-CLIIC.  
  }  
  \label{fig:forecast}
\end{figure*}

\subsection{Implicit Regularization Hypothesis}

One might ask if the benefits we observe in forecasting and hotspot
prediction have anything to do with the actual indicator themselves. A plausible
alternative explanation is that the indicators are simply providing \textit{implicit
  regularization} on top of  the basic AR model when we include them as lagged
features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.

To test this hypothesis, we reran all of the prediction experiments but with
$X_{\ell,t}$ in the each indicator models
%in \eqref{eq:forecast_ar_x} \eqref{eq:hotspot_ar_x}
replaced with suitable random noise (bootstrap samples
from a signal's historical distribution).  The results, shown and explained more
precisely in the supplement, are vastly different (worse) than the original set
of results.  In both forecasting and hotspot prediction, the ``fake'' indicator
models
% (with random noise as auxiliary features)
offered essentially no
improvement over the pure AR model, which provides a strong rejection
(informally speaking) of the implicit regularization hypothesis.

On the topic of regularization, it is also worth noting that the use of 
$\ell_1$ regularization (tuned by cross-validation) in fitting any of the models
in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x}, \eqref{eq:hotspot_ar}, or 
\eqref{eq:hotspot_ar_x} did not generally improve their performance
(experiments not shown).  This is likely due to the fact that the number of
training samples is large compared to the number of features (6,426 training
samples and only 3--6 features). [DJM: and that proper CV requires longer
training windows, exacerbating the nonstationarity, a scenario in which its
behavior is poorly understood.]

% \begin{figure}[t]
% \includegraphics[width=\columnwidth]{fig/hot-1.pdf}
% \caption{Area under the curve for each hotspot prediction model, calculated over
%   all prediction dates and all HRRs.} 
% \label{fig:hotspot}
% \end{figure}

\subsection{Evaluation in Up, Down, and Flat Periods}

The course of the pandemic has played out quite differently across space and
time. Aggregating case rates nationally shows three pronounced waves, but the
behavior is more nuanced at the HRR level.  Figure \ref{fig:trajectory} is a
single example of a forecast in a period of relatively flat case trends, as New
York City enters what would become its second wave. The AR forecaster's 50\%
prediction interval contains this upswing, but its forecasted median is clearly
below the finalized case data.  Unfortunately, this behavior is fairly typical
of all forecasters: during upswings, the forecasted median tends to fall below 
the target, while the reverse is true during downswings.

Figure~\ref{fig:up_down_flat} shows histograms of the differences in WIS of the 
AR model and each indicator model, where we stratify these differences by
whether the target occurs during a period of increasing cases rates (up),
decreasing case rates (down), or flat case rates (flat). To define the
increasing period, we use the same definition we used for the hotspot task in
Table \ref{tab:analysis_tasks}.  Therefore all hotspots are ``up'', while all
non-hotspots are either ``flat'' or ``down''.  For the ``down'' scenario, we 
simply use the opposite of the hotspot definition: $Y_{\ell,t}$ decreases by
more than 20\% relative to the preceding week.

While the performance of all forecasters, including AR, will generally degrade
in up periods, different models exhibit different and interesting patterns.
CHNG-CLI, CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have
large right tails (displaying improvements over AR) during the down periods.
Google-AA and CTIS-CLI-in-community have large right tails during the flat
periods. CHNG-CLI and DV-CLI have large left tails (poor forecasts relative to
AR) in flat and up periods.  Google-AA is the only model that outperforms
the AR model, on average, over up periods.  Overall, the indicators seem to help
more during flat or down periods than up periods, with the exception of
Google-AA.

\begin{figure*}[t]
  \includegraphics[width=\textwidth]{fig/upswing-histogram-1.pdf}
  \caption{Histogram of the difference in WIS for the AR model and that for each
    indicator model, stratified by up, down, or flat period, measured in terms
    of case trends. Note that \textit{larger} differences here are better for
    each indicator model.  The y-axis is on the log scale to emphasize tail 
    behavior.}   
  \label{fig:up_down_flat}
\end{figure*}

The supplement pursues this analysis further.  For example, we examine
classification accuracy and log-likelihood for the hotspot task and find a
similar phenomenon: the indicators considerably improve accuracy or
log-likelihood during flat or down periods, with more mixed behavior during up
periods when  CHNG-CLI, CHNG-COVID, and DV-CLI, in particular, lead to decreased
performance.

% Essentially, during downswings, improved performance is strongly correlated
% with predicing larger declines than the AR model. The AR model tends to
% overpredict relative to the decline, but adding indicators tends to diminish
% this effect. During upswings, the AR model tends to underpredict, but when the
% addition of an indicator improves performance (a more rare occurrence) it is 
% rarely due to higher median predictions.

% RJT: Commented out the last paragraph for now, because this is kind of a tough
% thing to explain succintly. Can revisit if we really want to include it here,
% or just let them find it in the supplement?
% DJM: Agreed. In the supplement.

\subsection{Effects of Leading or Lagging Behavior}

As described in the methods section, each of the indicators we examine could be
said to measure aspects of disease progression that would precede a positive
test. That is, we imagine that these signals should ``lead'' cases. It is
entirely reasonable to imagine that, prior to an increase of confirmed COVID-19
tests reported by a public health authority in a particular location, we would
see an increase in medical insurance claims for COVID-related outpatient
visits. However, it may well be the case that such behavior is different
during different periods. In fact, we find empirically that the ``leadingness''
of an indicator (degree to which it leads case activity) tends to be more
pronounced in down or flat periods than in up periods, a plausible explanation
for the decreased performance in up periods noted above.

In the supplement, we show how to define a quantitative score to measure the
leadingness of an indicator, at any time $t$ and any location $\ell$, based on
cross correlations to case rates over a short time window around $t$.
The higher this score, the greater it ``leads'' case activity.  Figure 
\ref{fig:leading} displays correlations between the leadingness score of an 
indicator and the WIS difference (AR model minus an indicator model), stratified
by whether the target is classified as up, down, or flat.  One would expect that  
the WIS difference would be positively correlated with leadingness.  Somewhat
surprisingly, this relationship turns out to be strongest in down periods
and weakest in up periods.  In fact, it is very nearly the case that for each
indicator, the strength of correlations only decreases as we move from down to
flat to up  periods.  In the supplement, we extend this analysis by examining
analogous ``laggingness'' scores, but we do not find as clear patterns.

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/leading-only-1.pdf}
  \caption{Correlation of the difference in WIS with the ``leadingness'' of the
    indicator at the target date, stratified by up, down, or flat period.} 
  \label{fig:leading}
\end{figure}

\section{Discussion}

Can auxiliary indicators improve COVID-19 forecasting and hotspot prediction
models?  Our answer, based on analyzing five auxiliary indicators from the
COVIDcast API (defined using from medical insurance claims, internet-based
surveys, and internet search trends) is undoubtedly ``yes''. However, there are
levels of nuance to such an answer that must be explained.  None of the
indicators that we have investigated appear to be the ``silver bullet'' that one
might have hoped for, revolutionizing the tractability of the prediction
problem, rendering it easy when it was once hard (in the absence of auxiliary
information).  Rather, the gains in accuracy from the indicator models (over an
autoregressive model based only on past case rates) appear to be nontrivial, and
consistent across modes of analysis, but modest.  In forecasting, the indicator
models are found to be most useful in periods in which case rates are flat or
trending down, rather than periods in which case rates are trending up (as one
might hope to see is the benefit provided by a hypothetical ``leading
indicator'').

As described previously, it is likely that we could improve the indicator models
by using location-specific effects, as well as using nowcasting techniques to
estimate finalized indicator values before we use them as features (to account
for backfill in the claims-based signals in particular).  Beyond this, it is
certainly possible that more sophisticated models for forecasting or hotspot
prediction would lead to different results, and possibly even different
insights. Natural directions to explore include using multiple indicators in a
single model, allowing for interaction terms, and leveraging HRR demographics or
mobility patterns.  That said, we are doubtful that more sophisticated modeling
techniques would change the ``topline'' conclusion---that auxiliary indicators
can provide nontrivial, consistent, but modest gains in forecasting and hotspot
prediction. Whether a more sophisticated model would be able to leverage the
indicators in such a way as to change some of the finer conclusions (e.g., by 
offering clear improvements in periods in which cases are trending up) is less 
clear to us.

We reiterate the importance of using vintage data for rigorous backtesting. Data
sources that are relevant to public health surveillance are often subject to
revision, sometimes regularly (such as medical claims data) and sometimes
unpredictably (such as COVID-19 case reports).  When analyzing models that are
designed to predict future events, if we train these models and make predictions
using finalized data, then we are missing a big part of the story.  Not only
will our sense of accuracy be unrealistic, but certain models may degrade by a
greater or lesser extent when they are forced to work with vintage data, so 
backtesting them with finalized data may lead us to make modeling decisions that
are suboptimal for true test-time performance.

In this paper, we have chosen to consider only very simple forecasting models,
while devoting most of our effort to accounting for as much of the complexity of
the data and evaluation as possible.  In fact, our paper is as much about
demonstrating how one might address questions about model comparisons and
evaluation in forecasting and hotspot prediction in general, as it is about
providing rigorous answers to such questions in the context of COVID-19 case
rates in particular.  We hope that others will leverage our framework, and build
on it, so that it can be used to guide work that advances the frontier of
predictive modeling for epidemics and pandemics.

\acknow{We thank Matthew Biggerstaff, Logan Brooks, Johannes Bracher, Michael 
  Johansson, Evan Ray, Nicholas Reich, and Roni Rosenfeld for several
  enlightening conversations about forecast evaluation. The research was
  supported by Centers for Disease Control and Prevention (CDC) grant
  U01IP001121; gifts from Facebook, Google, McCune, and Optum; the National
  Science Foundation Graduate Research Fellowship Program (NSF GRFP) award
  DGE175016; the Canadian Statistical Sciences Institute; and the National
  Sciences and Engineering Research Council of Canada (NSERC) RGPIN-2021-02618.}

\showacknow{} % Display the acknowledgments section

% Bibliography
\bibliography{../../common/covidcast.bib}

\end{document}