### 2021-06-21 indicators/paper/supplement.Rmd
Tweak non-caption text between hosp-related figs in indicators supplement

---
bibliography: [../../common/covidcast.bib, pnas-materials/pnas-sample.bib]
output:
  pdf_document:
    keep_tex: true
    template: pnas-suppl-template.tex
params:
  fd_casefloor: 30
  flag_jumps: 15
---

<!-- PNAS materials -->

<!-- Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found". -->
\instructionspage

<!-- we need this -->
\maketitle

<!-- Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text. -->

\SItext

All our text goes here. Reference figures with the R chuck label as Figure~\ref{fig:fcast-finalized}.

All figures go on their own page after all of the text...

Edit `pnas-suppl-template.tex` for the correct Author list and Title.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6.5, fig.height = 4.5,
                      fig.align = "center",
                      fig.path = "fig/",
                      cache = TRUE,
                      out.width = "\\textwidth",
                      autodep = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(covidcast)
library(gridExtra)
library(ggplot2)
library(ggforce) # for splitting across pages
library(scales)
library(dplyr)
```

\begin{figure}
\includegraphics[width=\textwidth]{fig/severity-pyramid.pdf}
\caption{The epidemiological "severity pyramid" represents the progression of
cases from the public, through infection, through increasingly severe stages of
disease. The annotations here represent the data sources collected by Delphi's
COVIDcast Epidata API. TODO Edit to use consistent language to paper}
\end{figure}

\clearpage

```{r state-trends}
SOURCES <- c("jhu-csse", "chng", "chng", "fb-survey")
SIGNALS <- c("confirmed_7dav_incidence_prop",
             "smoothed_adj_outpatient_cli",
             "smoothed_adj_outpatient_covid",
             "smoothed_whh_cmnty_cli")
colors <- c("black", scales::hue_pal()(5))
names <- c("Cases", "CHNG-CLI", "CHNG-COVID", "CTIS-CLI-in-community")

# geo_labeler is a function from geo_value to human-readable name
get_trend_grid <- function(trends, geo_labeler) {
  case_max <- max(trends[[1]]$value)

  trends <- aggregate_signals(trends, format = "long") %>%
    group_by(signal) %>%
    mutate(value = value / max(value, na.rm = TRUE) * case_max) %>%
    ungroup() %>%
    mutate(signal = factor(signal, levels = SIGNALS, labels = names),
           geo_value = geo_labeler(geo_value))

  g <- ggplot(trends, aes(x = time_value, y = value, color = signal)) +
    geom_line() +
    scale_color_manual(breaks = names, values = colors) +
    labs(x = "Date", y = "Signal value (scaled)") +
    scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
                 date_minor_breaks = "1 month", guide = guide_axis(angle = 45)) +
    theme_bw() +
    theme(legend.position = "bottom", legend.title = element_blank())

  return(g)
}

state_trends <- covidcast_signals(
  SOURCES,
  SIGNALS,
  "2020-04-15", "2021-04-15",
  geo_type = "state"
)

g <- get_trend_grid(state_trends, function(val) { abbr_to_name(toupper(val)) })
```

```{r state-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. Cases are reported cases per 100,000 population. Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range. (part 1 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 1)
```

\clearpage

```{r state-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 2 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 2)
```

\clearpage

```{r state-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 3 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 3)
```

\clearpage

```{r state-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for states across the United States. (part 4 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 4)
```

\clearpage

```{r county-trends}
# Same thing again, but the 50 largest counties
top_counties <- county_census %>%
  filter(COUNTY != 0) %>% # these are states
  arrange(desc(POPESTIMATE2019)) %>%
  head(n = 50) %>%
  pull(FIPS)

# Work around an API problem: specifying geo_values in the query causes it to
# fail on the CHNG data because of bad database indexing. Filter after the
# fact instead. DANGER: Very slow.
county_trends_nochng <- covidcast_signals(
  SOURCES[c(1,4)], SIGNALS[c(1,4)],
  "2020-04-15", "2021-04-15",
  geo_type = "county",
  geo_values = top_counties
)

county_trends_chng <- covidcast_signals(
  SOURCES[2:3], SIGNALS[2:3],
  "2020-04-15", "2021-04-15",
  geo_type = "county"
)

county_trends_chng <- lapply(
  county_trends_chng,
  function(df) {
    filter(df, geo_value %in% top_counties)
  }
)

county_trends <- append(county_trends_nochng, county_trends_chng[1])
county_trends <- append(county_trends, county_trends_chng[2])
```

```{r county-trend-grids-1, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. Cases are reported cases per 100,000 population. Other signals are scaled to have the same global maximum across all counties and times, so they can be presented in the same range. (part 1 of 4)", fig.height=8}
g <- get_trend_grid(
  county_trends,
  function(fipsen) {
    paste0(county_fips_to_name(fipsen), ", ", name_to_abbr(state_fips_to_name(fipsen)))
  }
)

g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 1)
```

\clearpage

```{r county-trend-grids-2, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 2 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 2)
```

\clearpage

```{r county-trend-grids-3, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 3 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 3)
```

\clearpage

```{r county-trend-grids-4, fig.cap="Trends of cases, CHNG-CLI, CHNG-COVID, and CTIS-CLI-in-community for the 50 most populous counties in the United States. (part 4 of 4)", fig.height=8}
g + facet_wrap_paginate(vars(geo_value), ncol = 3, nrow = 5, page = 4)
```

\clearpage

```{r hospitalization_time_trends_national, fig.cap="National trends, from April 2020 to April 2021, of HHS-reported confirmed COVID-19 hospital admissions, plus another hospitalization signal in the COVIDcast API. Here, the HHS data has been limited to values from 2020-08-01 onward due to changes in reporting behavior, and processed into a 7-day trailing average rate per 100,000 resident population using Census Bureau estimates for 2019. HSP-Hosp is the percentage of new hospital admissions with COVID-associated diagnoses, based on claims data from health system partners, smoothed in time and adjusted for systematic day-of-week effects (HSP-Hosp is available from an earlier date when working at finer geographical resolutions). HSP-Hosp has been scaled to have the same maximum value as the HHS data."}
start_day = "2020-04-15"
end_day = "2021-04-15"
as_of_day = "2021-06-01"

HOSP_SOURCES <- c("hhs", "hospital-admissions")
HOSP_SIGNALS <- c("confirmed_admissions_covid_1d", "smoothed_adj_covid19_from_claims")
hosp_names = c("Hospitalizations", "HSP-Hosp")
hosp_signal_has_national = c(TRUE, TRUE)
hhs_start_date = as.Date("2020-08-01")

hosp_colors = c("black", scales::hue_pal()(length(HOSP_SIGNALS)-1))

hosp_raw_signal_data <- covidcast_signals(
  HOSP_SOURCES[hosp_signal_has_national],
  HOSP_SIGNALS[hosp_signal_has_national],
  start_day, end_day,
  as_of=as_of_day,
  geo_type="nation"
)

## For HHS data, remove pre-cutoff-date period, and take a 7-day rolling average rate:
postprocess_hhs_hosp_df <- function(hhs_hosp_df, geo_type, population_df) {
  stopifnot(nrow(distinct(hhs_hosp_df, data_source, signal)) == 1)
  hhs_hosp_df %>%
    filter(time_value >= hhs_start_date) %>%
    group_by(geo_value) %>%
    arrange(time_value) %>%
    mutate(value = c(rep(NA, 6), sum(value[1:7]) + c(0, cumsum(tail(value,-7) - head(value, -7))))/7) %>% # (could also use zoo::rollmeanr)
    slice(-(1:6)) %>%
    left_join(population_df) %>%
    mutate(value = value/population*100000, population=NULL) %>%
    ungroup() %>%
    {as.covidcast_signal(., signal=unique(.$signal), geo_type=geo_type, time_type="day")}
}
hosp_signal_data <- hosp_raw_signal_data
hosp_signal_data[[1]] <- postprocess_hhs_hosp_df(
  hosp_raw_signal_data[[1]], "nation",
  ## use national entry appearing in state_census
  state_census %>% transmute(geo_value=tolower(ABBR), population=POPESTIMATE2019)
)

hosp_main_signal_max <- max(hosp_signal_data[[1]]$value, na.rm=TRUE)

hosp_rescaled <- aggregate_signals(hosp_signal_data, format = "long") %>%
  group_by(signal) %>%
  mutate(value = value / max(value, na.rm=TRUE) * hosp_main_signal_max) %>%
  ungroup()

hosp_rescaled$signal <- factor(hosp_rescaled$signal, levels=HOSP_SIGNALS, labels=hosp_names)

ggplot(hosp_rescaled, aes(x = time_value, y = value, color = signal)) +
  geom_line() +
  scale_color_manual(breaks = hosp_names, values = hosp_colors) +
  labs(x = "Date", y = "Signal value (scaled)") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

\clearpage

```{r hosp-correlation-data}
df_hosp_signals = vector("list", length(HOSP_SIGNALS))
for (i in seq_along(HOSP_SIGNALS)) {
  df_hosp_signals[[i]] =
    covidcast_signal(HOSP_SOURCES[i], HOSP_SIGNALS[i],
                     start_day, end_day, as_of=as_of_day,
                     geo_type="state")
}
state_populations = state_census %>%
  transmute(geo_value=tolower(ABBR), population=POPESTIMATE2019)
df_hosp_signals[[1]] <- postprocess_hhs_hosp_df(df_hosp_signals[[1]], "state", state_populations)

## Fetch HHS-reported confirmed hospital admissions and smooth with 7-day
## trailing average rate:
df_raw_hospitalizations =
  covidcast_signal("hhs", "confirmed_admissions_covid_1d",
                   start_day, end_day, as_of=as_of_day,
                   geo_type="state")
df_hospitalizations = postprocess_hhs_hosp_df(df_raw_hospitalizations, "state", state_populations)

## Consider only state-like jurisdictions for which all signals of interest have at least 50 rows:
hosp_geo_values = unique(df_hospitalizations$geo_value)
for (i in seq_along(HOSP_SIGNALS)) {
  hosp_geo_values = intersect(hosp_geo_values, df_hosp_signals[[i]] %>% count(geo_value) %>% filter(n > 50) %>% pull(geo_value))
}

cat('Number of unique jurisdictions included in correlation analysis below:\n')
length(hosp_geo_values)
```

```{r hosp-correlation-instances}
## Find common set of (geo_value, time_value) instances across all signals:
hosp_common_instances =
  select(df_hospitalizations, geo_value, time_value) %>%
  filter(geo_value %in% hosp_geo_values)

for (i in seq_along(HOSP_SIGNALS)) {
  hosp_common_instances <- inner_join(hosp_common_instances,
                                      select(df_hosp_signals[[i]], geo_value, time_value),
                                      by=c("geo_value","time_value"))
}

cat('Number of common instances among all signals of interest in correlation analysis below:\n')
nrow(hosp_common_instances)
```

```{r hosp-correlations-by-time, fig.cap="Geo-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated for all times with sufficient available data within this period, over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available."}
n_hosp_signals = length(HOSP_SIGNALS)
df_list_hosp_geowise_cor = vector("list", n_hosp_signals)
for (i in 1:n_hosp_signals) {
  df_list_hosp_geowise_cor[[i]] = covidcast_cor(
    df_hosp_signals[[i]] %>% inner_join(hosp_common_instances, by=c("geo_value","time_value")),
    df_hospitalizations %>% inner_join(hosp_common_instances, by=c("geo_value","time_value")),
    by = "time_value", method = "spearman"
  )
  df_list_hosp_geowise_cor[[i]]$signal = hosp_names[[i]]
}
df_long_hosp_geowise_cor = do.call(rbind, df_list_hosp_geowise_cor[-1])
df_long_hosp_geowise_cor$signal = factor(df_long_hosp_geowise_cor$signal, hosp_names[-1])

ggplot(df_long_hosp_geowise_cor, aes(x = time_value, y = value)) +
  geom_line(aes(color = signal)) +
  scale_color_manual(breaks = hosp_names, values = hosp_colors) +
  expand_limits(y=0) +
  labs(x = "Date", y = "Correlation") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r hosp-correlations-by-state, fig.cap="Time-wise correlations with hospitalization rates derived from HHS data, from April 15, 2020 to April 15, 2021, calculated over all state-like jurisdictions for which each signal was reported on at least 50 days during this period, limited to state-day combinations for which both signals are available."}
df_list_hosp_timewise_cor = vector("list", length(HOSP_SIGNALS))
for (i in seq_along(HOSP_SIGNALS)) {
  df_list_hosp_timewise_cor[[i]] = covidcast_cor(
    df_hosp_signals[[i]] %>% inner_join(hosp_common_instances, by=c("geo_value","time_value")),
    df_hospitalizations %>% inner_join(hosp_common_instances, by=c("geo_value","time_value")),
    by = "geo_value", method = "spearman"
  )
  df_list_hosp_timewise_cor[[i]]$signal = hosp_names[[i]]
}
df_long_hosp_timewise_cor = do.call(rbind, df_list_hosp_timewise_cor[-1])
df_long_hosp_timewise_cor$signal = factor(df_long_hosp_timewise_cor$signal, hosp_names[-1])

ggplot(df_long_hosp_timewise_cor, aes(value)) +
  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
  labs(x = "Correlation", y = "Density") +
  expand_limits(x=c(-1,1)) +
  scale_color_manual(breaks = hosp_names, values = hosp_colors) +
  scale_fill_manual(breaks = hosp_names, values = hosp_colors) +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

<!-- PNAS says: Add this line AFTER all your figures and tables -->
\FloatBarrier