### 2021-06-21 forecast/paper/pnas-suppl-template.tex
Minor typos.

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}

\input{../../common/prelim.tex}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

$for(header-includes)$
$header-includes$
$endfor$

\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
  Prediction?}

\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison Hu, Nat DeFries,
  Sangwon Hyun, Natalia L. Oliveira, James Sharpnack, Jingjing Tang, Robert
  Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J. Tibshirani} 
\correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}

\begin{document}

\maketitle

\SItext

$body$

\FloatBarrier

\dataset{predictions.zip}{Archived {\tt .RDS} (R objects) files containing
  all predictions for forecasting and hotspots using vintage and finalized data.
Persistent DOI to be added at publication.}

\dataset{evaluations.zip}{Archived {\tt .RDS} (R objects) files containing
  all evaluations for forecasting and hotspots using vintage and finalized data.
  Persistent DOI to be added at publication.}

\dataset{analysis.zip}{Archived {\tt .RDS} (R objects) files containing
  additional data used to produce graphics and conclusions in the manuscript.
  Persistent DOI to be added at publication.}

\dataset{code.zip}{R script files containing
  all code used to reproduce the analysis described in the manuscript.
  Persistent DOI to be added at publication.}

%\bibliography{$for(bibliography)$$bibliography$$sep$,$endfor$}

\end{document}