### 2021-06-22 forecast/paper/supplement-text.tex

\section{Examining the relative advantage of using finalized rather than vintage data}

The goal of this section is to quantify the effect of
not properly accounting for the question of ``what was known when'' in
performing retrospective evaluations of forecasters.   Figures
\ref{fig:fcast-finalized} and \ref{fig:hot-finalized} show 
what Figures 3 and 4 in the main paper would have looked like if we
had simply trained all models using the finalized data rather than
using vintage data.  This comparison can be seen more
straightforwardly in Figures \ref{fig:fcast-honest-v-finalized} and
\ref{fig:hot-honest-v-finalized}, which show the ratio in performance between
the vintage and finalized 
versions.  When methods are 
given the finalized version of the data rather than the version available at the
time that the forecast would have been made, all methods 
appear (misleadingly) to have better performance than they would
have had if run prospectively.  For example, for forecasting case rates
7-days ahead, the WIS of all methods is at least 8\% larger than what would have been
achieved using finalized data.  This effect
diminishes as the forecasting horizon increases, reflecting the fact
that longer-horizon forecasters rely less heavily on recent data than very
short-horizon forecasters.  Crucially, some methods are
``helped'' more than others by the less scrupulous retrospective
evaluation, underscoring the difficulty of avoiding misleading
conclusions when performing retrospective evaluations of forecasters.

\chngcli~(and, to a lesser extent, the other claims-based
signals) is the most  
affected by this distinction, reflecting the latency in claims-based
reporting.  This underscores the importance of efforts to provide ``nowcasts'' for
claims signals 
(which corresponds to a 0-ahead forecast of what the claims
signal's value will be once all data has been collected). Looking at
the \chngcli~and \dv~curves in Figure \ref{fig:fcast-finalized}, we
can see that they perform very similarly when trained on the finalized data.
This is reassuring 
because they are, in principle, measuring the same thing (namely, the percentage of
outpatient visits that are primarily about COVID-related symptoms).
The substantial difference in their curves in Figure 3 of the
main paper must therefore reflect their having very different
backfill profiles.

While using finalized rather than vintage data affects \dv~the least for forecasting,
it is one of the most affected methods for the
hotspot problem. 
This is a reminder that the forecasting and hotspot problems are fundamentally
distinct.  For example, the hotspot problem does not measure
the ability to distinguish between flat and downward trends.

Even the \ar~model is affected by this distinction, reflecting the
fact that the case rates themselves (i.e., the response values) are also
subject to revision.  The forecasters based on indicators are thus
affected both by revisions to the indicators and by revisions to the
case rates.  In the case of the \gs~model, in which we only used finalized
values for the \gs~indicator, the difference in performance can be
wholly attributed to revisions of case rates.

\section{Aggregating with geometric mean}

In this section, we consider using the geometric mean instead of the
arithmetic mean when aggregating the weighted interval score (WIS) across
location-time pairs. 
There are three reasons why using the geometric mean may be desirable.
\begin{enumerate}
\item  WIS is right-skewed, being bounded below by zero and having
  occasional very large values.   Figure \ref{fig:wis-densities} illustrates that the
  densities appear roughly log-Gaussian.  The geometric mean is a
  natural choice in such a context since the relative ordering of forecasters is
  determined by the arithmetic mean of the {\em logarithm} of their WIS
  values.
\item In the main paper, we report the ratio of the mean WIS of a
  forecaster to the mean WIS of the baseline forecaster. Another
  choice could be to take the mean of the ratio of WIS values for the
  two methods. This latter choice would penalize a method less for
  doing poorly where the baseline forecaster also does poorly.
  Using instead the geometric mean makes the order of aggregation and
  scaling immaterial since the ratio of geometric means is the same as
  the geometric mean of ratios.
\item If one imagines that a forecaster's WIS is composed of
  multiplicative space-time effects $S_{\ell,t}$ shared across all forecasters,
  i.e. $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{f,t}$ a
  forecaster-specific error, then
  taking the ratio of two forecasters' geometric mean WIS values will
  effectively cancel these space-time effects.
\end{enumerate}

Figure \ref{fig:fcast-adjusted} uses the geometric
mean for aggregation.  Comparing this with Figure 3 of the main paper,
we see that the main conclusions are largely unchanged; however,
\chngcli~now appears better than \ar.  This behavior would be
expected if \chngcli's poor performance is attributable to a
relatively small number of large errors (as opposed to a large number
of moderate errors).  Indeed, Figure 5 of the main paper further
corroborates this, in which we see the heaviest left tails occuring
for \chngcli.

\section{Bootstrap results}

As explained in Section 2.B. of the main paper, a (somewhat cynical)
hypothesis for why we see benefits in forecasting and hotspot
prediction is that the indicators are not actually providing useful
information but they are instead acting as a sort of ``implicit
regularization,''  leading to shrinkage on the autoregressive
coefficients and therefore to less volatile predictions.  To investigate
this hypothesis, we consider fitting  ``noise features'' that in truth
should have zero coefficients.  Recall (from the main paper) that at each
forecast date, we 
train a model on 6,426 location-time pairs.  Indicator models are
based on six features, corresponding to the three autoregressive terms
and the three lagged indicator values.  To form noise indicator features,
we replace their values with those from a randomly chosen time-space pair
(while keeping the autoregressive 
features fixed).  In particular, at each location $\ell$ and
time $t$, for the forecasting task we replace the triplet $(X_{\ell,t}, X_{\ell,t-7},
X_{\ell,t-14})$ in Eq. (3) of the main paper with the triplet $(X_{\ell^*,t^*},
X_{\ell^*,t^*-7}, 
X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is a location-time pair
sampled with replacement from the 6,426 location-time pairs.  
Likewise in the hotspot prediction task, we replace the triplet
$(X_{\ell,t}^\Delta, X_{\ell,t-7}^\Delta, 
X_{\ell,t-14}^\Delta)$ in Eq. (5) of the main paper with
$(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta, 
X_{\ell^*,t^*-14}^\Delta)$.
Figures
\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No
method exhibits a noticeable performance gain over the \ar~method,
leading us to dismiss the implicit regularization hypothesis.

\section{Upswings and Downswings}

In this section we provide extra details about the upswing / flat / downswing 
analysis described in the main text. Figure \ref{fig:upswing-summary} shows
the overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in 
period. Figure \ref{fig:hotspots-upswing-downswing} shows the same information
for the hotspot task. On average, during downswings and flat periods, the 
indicator-assisted models have lower classification error and higher 
log likelihood than the AR model. For hotspots, both Google-AA and CTIS-CLIIC
perform better than the AR model during upswings, in contrast to the forecasting
task, where only Google-AA improves. For a related analysis, Figure 
\ref{fig:cor-wis-ratio} shows histograms of the Spearman
correlation (Spearman's $\rho$, a rank-based measure of association) between
the $\WIS(F)/\WIS(AR)$ and the magnitude of the swing. Again we see that case
rate increases are positively related to diminished performance of the 
indicator models.

One hypothesis for diminished relative performance during upswings is that 
the AR model tends to overpredict downswings and underpredict upswings. Adding
indicators appears to help avoid this behavior on the downswing but not as much 
on upswings. Figure \ref{fig:upswing-corr-table} shows the correlation between
between $\WIS(AR) - \WIS(F)$ and the difference of their median forecasts.
During 
downswings, this correlation is large, implying that improved relative
performance 
of $F$ is related to making lower forecasts than the AR model. The opposite is
true during upswings. This is largely to be expected. However, the relationship
attenuates in flat periods and during upswings. That is, when performance is
better 
in those cases, it may be due to other factors than simply making predictions
in the correct direction, for example, narrower confidence intervals.

\section{Leadingness and laggingness}

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness at location $\ell$, we use the cross correlation
function (CCF) 
between the two time series. The $\CCF_\ell(a)$ of an indicator $X_{\ell}$ and case
rates $Y_{\ell}$ is defined as
% \[
% \CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
%   \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
%   \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
% \]
% where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
% shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
% are the sample mean and
% standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
their Pearson correlation where $X_\ell$ has been
% sequence of Pearson correlations for each selected $a$, where we
aligned
with the values of $Y_{\ell}$ that occurred $a$ days earlier.
Thus, for any $a>0$, $\CCF_{\ell}(a)>0$ indicates that
$Y_{\ell,t}$ is moving together with $X_{\ell,t+a}$.
% days in the future.
In this case we say that $X_{\ell}$ is lagging
$Y_{\ell}$. For $a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is
positively correlated with $X_{\ell,t-a}$, so we say that
$X_{\ell}$ leads $Y_{\ell}$.

Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
containing Charlotte, North Carolina, from August 1, 2020 until the end of
September. These are the same signals shown in Figure 1 in the manuscript but
using finalized data. To
define ``leadingness'' we compute $\CCF_{\ell}(a)$ (as implemented with the R
function {\tt ccf()}) for each $a\in
\{-15,\ldots,15\}$ using the 56 days leading up to the target date. This is the same
amount of data used to train the forecasters: 21 days of training data, 21 days
to get the response at $a=21$, and 14 days for the longest lagged value. The
orange dashed horizontal line represents the 95\% significance threshold for
correlations based on 56 observations. Any correlations larger in magnitude than
this value are considered statistically significant under the null hypothesis of
no relationship. We define leadingness to be the sum of the significant
correlations that are leading (those above the dashed line with $a<0$) while
laggingness is the same but for $a>0$. In the figure, there are three
significant correlations on the ``leading'' side (at $a = -5, -4, -3$), so
leadingness will be the sum of those values while laggingness is 0:
%correlations, so both scores will be 0:
on September 28 in Charlotte, \dv~is leading cases
leading but not lagging.

Figure \ref{fig:lagging-only} shows the correlation between laggingness and the
difference in indicator WIS and AR WIS. Unlike leadingness (Figure 5 in the
manuscript) there is no obvious relationship that holds consistently
across indicators. This is heartening as laggingness
should not aid forecasting performance. On the other hand, if an indicator is
more lagging than it is leading, this may suggest diminished performance.
Figure~\ref{fig:diff-in-lead-lag} shows the correlation of the difference in
leadingness and laggingness with the difference in WIS. The pattern here is
largely similar to the pattern in leadingness described in the manuscript: the
relationship is strongest in down periods and weakest in up periods with the
strength diminishing as we move from down to flat to up for all indicators.

In calculating the CCF and the associated leadingness and laggingness scores, we
have used the finalized data, and we look at the behavior at the 
target date of the forecast. That is we are using the same data to
evaluate predictive accuracy as to determine
leadingness and laggingness. It should be noted that the leadingness of the
indicator at the time the model is trained may also be important. Thus, we could
calculate separate leadingness and laggingness scores for the trained model and
for the evaluation data and examine their combination in some way. We do not
pursue this combination further and leave this investigation for future work.

\section{Examining data in 2021}

In this section, we investigate the sensitivity of the results to the
period over which we train and evaluate the models.  In the main
paper, we end all evaluation on December 31, 2020.  Figures
\ref{fig:fcast-alldates}--\ref{fig:hot-alldates} show how the
results would differ if we extended this analysis through March
  31, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
the main paper, one sees that as ahead increases most methods now improve
relative to the baseline forecaster. When compared to other methods, 
\chngcli~appears much better than
it had previously; however, all forecasters other than \chngcov~and
\dv~are performing less well relative to the baseline than before.
These changes are likely due to the differing nature of the pandemic
in 2021, with flat and downward trends much more common than upward
trajectories.  Indeed, the nature of the hotspot prediction problem is
quite different in this period.  With a 21-day training window, it is
common for there to be many fewer hotspots in training.