### 2021-06-22 indicators/paper/pnas-suppl-template.tex
Fill out supplement template

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}

\input{../../common/prelim.tex}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

$for(header-includes)$
$header-includes$
$endfor$

\title{An Open Repository of Real-Time COVID-19 Indicators}
\author{Alex Reinhart, Logan Brooks, Maria Jahja, Aaron Rumack, Jingjing Tang,
  et al.}
\correspondingauthor{Alex Reinhart.\\E-mail: areinhar@stat.cmu.edu}

\begin{document}

$body$

\bibliography{$for(bibliography)$$bibliography$$sep$,$endfor$}

\end{document}