### 2021-06-23 indicators/code/indicator-paper-figures.Rmd

---
title: COVIDcast Indicators figures and tables
author: Alex Reinhart, et al.
date: June 23, 2021
---

This code accompanies the paper "An Open Repository of Real-Time COVID-19
Indicators" by Reinhart et al. It requires the [covidcast
package](https://cran.r-project.org/package=covidcast) for R, plus the packages
listed below.

```{r setup, message=FALSE}
library(covidcast)
library(gridExtra)
library(ggplot2)
library(scales)
library(dplyr)
library(purrr)
```

This file can be knit to HTML using
`rmarkdown::render("indicator-paper-figures.Rmd")` from within R, using the
[rmarkdown package](https://cran.r-project.org/package=rmarkdown). By default it
also saves publication-ready PDF versions of figures to the `../paper/fig/`
directory.

## Figure 1

National trends, from April 2020 to April 2021, of four signals in the COVIDcast
API.

```{r figure-1, message=FALSE, cache=TRUE}
SOURCES <- c("jhu-csse", "chng", "chng", "fb-survey")
SIGNALS <- c("confirmed_7dav_incidence_prop",
             "smoothed_adj_outpatient_cli",
             "smoothed_adj_outpatient_covid",
             "smoothed_wcli")
cs <- covidcast_signals(
  SOURCES,
  SIGNALS,
  "2020-04-15", "2021-04-15",
  geo_type = "nation"
)

case_max <- max(cs[[1]]$value)

rescaled <- aggregate_signals(cs, format = "long") %>%
  group_by(signal) %>%
  mutate(value = value / max(value, na.rm = TRUE) * case_max) %>%
  ungroup()

colors <- c("black", scales::hue_pal()(5))
names <- c("Cases", "CHNG-CLI", "CHNG-COVID", "CTIS-CLI",
          "CTIS-CLI-in-community", "Quidel-TPR")
rescaled$signal <- factor(rescaled$signal, labels = names[1:4])

ggplot(rescaled, aes(x = time_value, y = value, color = signal)) +
  geom_line() +
  scale_color_manual(breaks = names[1:4], values = colors[1:4]) +
  labs(x = "Date", y = "Signal value (scaled)") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())

ggsave("../paper/fig/time_trends_national.pdf", width = 5, height = 3)
```

# Figure 2

Geo-wise correlations with case rates, from April 15, 2020 to April 15, 2021,
calculated over all counties for which all signals were available and which had
at least 500 cumulative cases by the end of this period.

First, we fetch the required data from the API. This may take a while, since it
fetches the entire county-level history of several signals with large coverage.

```{r figure-2-fetch, message=FALSE, warning=FALSE, cache=TRUE}
sources <- c("chng", "chng", "fb-survey", "fb-survey", "quidel")
signals <- c("smoothed_adj_outpatient_cli",
             "smoothed_adj_outpatient_covid",
             "smoothed_cli",
             "smoothed_whh_cmnty_cli",
             "covid_ag_smoothed_pct_positive")

start_day <- "2020-04-15"
end_day <- "2021-04-15"
as_of_day <- "2021-06-01"

df_signals <- vector("list", length(signals))
for (i in seq_along(signals)) {
  df_signals[[i]] <- covidcast_signal(sources[i], signals[i], start_day,
                                      end_day, as_of = as_of_day)
}

# Fetch JHU-CSSE confirmed case incidence proportion (smoothed with 7-day
# trailing average)
df_cases <- covidcast_signal("jhu-csse", "confirmed_7dav_incidence_prop",
                             start_day, end_day, as_of = as_of_day)
n <- length(signals)
```

Next, we find counties with at least 500 cumulative cases.

```{r figure-2-counties, message=FALSE, cache=TRUE}
case_num <- 500
geo_values <- covidcast_signal("jhu-csse", "confirmed_cumulative_num",
                               max(df_cases$time_value),
                               max(df_cases$time_value),
                               as_of = as_of_day) %>%
  filter(value >= case_num) %>%
  pull(geo_value)

for (i in seq_along(signals)) {
  geo_values <- intersect(geo_values, df_signals[[i]]$geo_value)
}

colors <- c("black", scales::hue_pal()(5))
names <- c("Cases", "CHNG-CLI", "CHNG-COVID", "CTIS-CLI",
           "CTIS-CLI-in-community", "Quidel-TPR")
```

There are `r length(geo_values)` counties that have at least 500 cumulative
cases and have measurements for all the signals desired.

```{r figure-2, cache=TRUE}
df_cor <- vector("list", n)
for (i in 1:n) {
  df_cor[[i]] <- covidcast_cor(df_signals[[i]] %>%
                                 filter(geo_value %in% geo_values),
                               df_cases %>%
                                 filter(geo_value %in% geo_values),
                               by = "time_value", method = "spearman")
  df_cor[[i]]$signal <- names[i]
}

df <- do.call(rbind, df_cor)
df$signal <- factor(df$signal, labels = names[2:6])

ggplot(df, aes(x = time_value, y = value)) +
  geom_line(aes(color = signal)) +
  scale_color_manual(breaks = names[2:6], values = colors[2:6]) +
  ylim(0, NA) +
  labs(x = "Date", y = "Correlation") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  theme_bw() +
  guides(col = guide_legend(nrow = 2)) +
  theme(legend.position = "bottom", legend.title = element_blank())

ggsave("../paper/fig/geo_wise_corr.pdf", width = 5, height = 3.5)
```

## Figure 3

Time-wise correlations with case rates, from April 15, 2020 to April 15, 2021.

```{r figure-3, warning=FALSE, cache=TRUE}
df_cor <- vector("list", length(signals))

for (i in seq_along(signals)) {
  df_cor[[i]] <- covidcast_cor(df_signals[[i]] %>%
                                 filter(geo_value %in% geo_values),
                               df_cases %>%
                                 filter(geo_value %in% geo_values),
                               by = "geo_value", method = "spearman")
  df_cor[[i]]$signal <- names[i]
}

df <- do.call(rbind, df_cor)
df$signal <- factor(df$signal, labels = names[2:6])

ggplot(df, aes(value)) +
  geom_density(aes(color = signal, fill = signal), alpha = 0.4) +
  labs(x = "Correlation", y = "Density") +
  scale_color_manual(breaks = names[2:6], values = colors[2:6]) +
  scale_fill_manual(breaks = names[2:6], values = colors[2:6]) +
  theme_bw() +
  guides(col = guide_legend(nrow = 2)) +
  theme(legend.position = "bottom", legend.title = element_blank())

ggsave("../paper/fig/time_wise_correlation.pdf", width = 5, height = 3.5)
```

## Figure 4

```{r figure-4, message=FALSE, cache=TRUE}
sa_fips <- "48029"
sa_start <- "2020-06-15"
sa_end <- "2020-08-15"
sa_anomaly_date <- as.Date("2020-07-16")

sa_cases <- covidcast_signal("jhu-csse", "confirmed_7dav_incidence_num",
                             sa_start, sa_end, geo_values = sa_fips)

sa_cli <- covidcast_signal("fb-survey", "smoothed_whh_cmnty_cli",
                            sa_start, sa_end, geo_values = sa_fips)

g1 <- ggplot(sa_cases, aes(x = time_value, y = value)) +
  geom_vline(xintercept = sa_anomaly_date, color = "gray", linetype = "dashed",
             size = 0.75, alpha = 0.75) +
  geom_line() +
  labs(x = "", y = "Cases", title = "Cases") +
  theme_bw()

g2 <- ggplot(sa_cli, aes(x = time_value, y = value)) +
  geom_vline(xintercept = sa_anomaly_date, color = "gray", linetype = "dashed",
             size = 0.75, alpha = 0.75) +
  geom_line() +
  labs(x = "", y = "% CLI-in-community", title = "CTIS-CLI-in-community") +
  scale_y_continuous(labels = label_percent(scale = 1, accuracy = 1),
                     limits = c(NA, 75)) +
  theme_bw()

grid.arrange(g1, g2, ncol = 2)

g <- arrangeGrob(g1, g2, ncol = 2)

ggsave("../paper/fig/bexar_compare.pdf", g, width = 6, height = 4)
```

## Figure 5

Estimated percentage of doctor's visits due to COVID-like illness displayed
across multiple issue dates, with later issue dates adding additional data and
revising past data from prior issue dates.

```{r figure-5, message=FALSE, warning=FALSE, cache=TRUE}
as_ofs <- seq(as.Date("2020-06-08"), as.Date("2020-08-01"), by = "week")
states <- c("az", "ca", "ny", "tx")
dv_as_of <- map_dfr(as_ofs, function(as_of) {
  covidcast_signal(data_source = "doctor-visits", signal = "smoothed_adj_cli",
                   start_day = "2020-06-01", end_day = as_of - 1L,
                   geo_type = "state", geo_values = states, as_of = as_of) %>%
    mutate(as_of = as_of)
})

## Now plot the each "as of" time series curve, faceted by state
dv_as_of %>%
  mutate(geo_label = factor(
    geo_value, levels = states,
    labels = abbr_to_name(states, ignore.case = TRUE))) %>%
  mutate(as_of = factor(format(as.Date(as_of), "%b %d"))) %>%
  ggplot(aes(x = time_value, y = value, color = as_of)) +
  geom_line() +
  facet_wrap(vars(geo_label), nrow = 2) +
  labs(color = "As of:", x = "Date", y = "% doctor's visits due to CLI") +
  expand_limits(y = 0) +
  scale_y_continuous(labels = label_percent(scale = 1, accuracy = 1),
                     breaks = seq(0, 12, by = 4)) +
  theme_bw() +
  theme(legend.pos = "bottom", legend.direction = "horizontal")

ggsave("../paper/fig/dv_as_of.pdf", width = 5, height = 4)
```

## Figure 6

Percentage point increase in SafeGraph estimates of the percentage of people
staying completely at home, based on aggregated mobile device data, between
March 1, 2020 and April 15, 2020.

```{r figure-6, message=FALSE, cache=TRUE}
begin_date <- "2020-03-01"
end_date <- "2020-04-15"

begin_data <- covidcast_signal("safegraph", "completely_home_prop_7dav",
                               begin_date, begin_date, geo_type = "state")
end_data <- covidcast_signal("safegraph", "completely_home_prop_7dav",
                             end_date, end_date, geo_type = "state")

diffs <- begin_data %>%
  inner_join(end_data, by = "geo_value", suffix = c(".begin", ".end")) %>%
  mutate(value = 100 * (value.end - value.begin)) %>%
  select(data_source = data_source.end, signal = signal.end, geo_value, value, issue = issue.end) %>%
  mutate(time_value = end_date) %>%
  as.data.frame %>%
  as.covidcast_signal(signal = "safegraph-change",
                      geo_type = "state")

plot(diffs, plot_type = "choro", range = c(10, 30),
     title = "", choro_params = list(legend_n = 5))

ggsave("../paper/fig/mobility-drop.pdf", width = 5, height = 4)
```