### 2021-09-20 survey/code/paper-tables.Rmd
Tidy up code and figures
1. Convert everything to put figures in the right directory.

2. Use consistent population data.

3. Move legends on contacts plot so the plots have consistent x axes.

4. Finish signal time series plot.

---
title: Survey paper tables and figures
author: Alex Reinhart, Alyssa Bilinski, and coauthors
date: July 12, 2021
---

This file contains the code used to generate Tables 2 and 3 and Figures 1, 3, 4,
and 5 of "COVID Trends and Impact Survey in the United States, 2020-2021:
Continuous real-time measurement of COVID-19 symptoms, risks, protective
behaviors, testing and vaccination" by Salomon et al. Figure 2 is generated
separately in `survey-paper-exhibits.ipynb`.

To knit this file: `rmarkdown::render("paper-tables.Rmd")`. This file requires
the packages listed below, plus unified versions of the complete survey
microdata; microdata is only available under [data use
agreement](https://cmu-delphi.github.io/delphi-epidata/symptom-survey/data-access.html).

```{r setup, message=FALSE}
library(data.table)
library(dplyr)
library(tidyr)
library(epitools)
library(tidycensus)
library(zoo)

library(ggplot2)
library(gridExtra)
library(scales)
library(geofacet)

library(knitr)
library(lubridate)
library(covidcast)

source("survey-utils.R")
```

## Setup

These categories are in order based on their coding in Q64, so indices
correspond to coding:

```{r more-setup}
occupations <- c(
  "Community and social service",
  "Education, training, and library",
  "Arts, design, entertainment, sports, and media",
  "Healthcare practitioners and technicians",
  "Healthcare support",
  "Protective service",
  "Food preparation and serving related",
  "Building and grounds cleaning and maintenance",
  "Personal care and service",
  "Sales and related",
  "Office and administrative support",
  "Construction and extraction",
  "Installation, maintenance, and repair",
  "Production",
  "Transportation and material moving",
  "Other"
)
```

We load the replication data files. Each object takes quite a lot of RAM, so
removing them and garbage-collecting is essential for this to run in the
available RAM on our server.

```{r load-data}
d1 <- fread("unified-1-year-part-1.csv.gz")
d2 <- fread("unified-1-year-part-2.csv.gz")

data <- rbind(d1, d2, fill = TRUE)

rm(d1, d2)
gc()
```

Some essential cleaning. We use 5-digit FIPS codes with leading zeros by
convention, and we'll want region names for plots:

```{r clean-data}
data$fips <- ifelse(is.na(data$fips), NA_character_, sprintf("%05d", data$fips))

state_census$REGION <- case_when(
  state_census$REGION == 1 ~ "Northeast",
  state_census$REGION == 2 ~ "Midwest",
  state_census$REGION == 3 ~ "South",
  state_census$REGION == 4 ~ "West"
)
```

## Metadata

This is based on data in the following time range:

```{r date-range}
range(data$StartDatetime)
```

Each wave has received this many responses:

```{r wave-count}
data %>%
  select(wave) %>%
  group_by(wave) %>%
  summarize(n = n()) %>%
  arrange(wave) %>%
  kable(col.names = c("Wave", "N"))
```

In total, that is:

```{r sample-size}
nrow(data)
```

Design effects:

```{r deff}
deff <- data %>%
  select(StartDatetime, weight) %>%
  mutate(StartDatetime = as.Date(StartDatetime),
         week = floor_date(StartDatetime, unit = "week"),
         epiweek = epiweek(StartDatetime),
         epiyear = epiyear(StartDatetime)) %>%
  group_by(week, epiyear, epiweek) %>%
  summarize(deff = mean(weight**2) / (mean(weight)**2),
            n = n(),
            .groups = "keep") %>%
  arrange(week) %>%
  select(week, epiyear, epiweek, deff, n)

write_csv(deff, "deff.csv")

ggplot(deff, aes(x = week, y = deff)) +
  geom_line() +
  labs(x = "", y = "Design effect") +
  scale_x_date(date_breaks = "2 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  scale_y_continuous(limits = c(0, NA)) +
  theme_bw()
```

## Respondent demographics

We'll produce unweighted and weighted proportions for each demographic variable.

```{r demo-setup}
proportionate <- function(data, var) {
  var <- ensym(var)

  data %>%
    filter(!is.na(!!var)) %>%
    group_by(!!var) %>%
    summarize(n = n(), sum_weight = sum(weight)) %>%
    mutate(unweighted_prop = n / sum(n) * 100,
           weighted_prop = sum_weight / sum(sum_weight) * 100) %>%
    select(!!var, n, unweighted_prop, weighted_prop)
}
```

### Gender

```{r demo-gender}
data %>%
  select(D1, weight) %>%
  mutate(gender = case_when(
    D1 == "1" ~ "Male",
    D1 == "2" ~ "Female",
    D1 == "3" | D1 == "4" ~ "Non-binary/self-describe",
    D1 == "5" | is.na(D1) ~ NA_character_
  )) %>%
  proportionate(gender) %>%
  kable(col.names = c("Gender", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Age

```{r demo-age}
gc()

data %>%
  select(D2, weight) %>%
  mutate(age = case_when(
    D2 == "1" ~ "18-24",
    D2 == "2" ~ "25-34",
    D2 == "3" ~ "35-44",
    D2 == "4" ~ "45-54",
    D2 == "5" ~ "55-64",
    D2 == "6" ~ "65-74",
    D2 == "7" ~ "75 plus",
    is.na(D2) ~ NA_character_
  )) %>%
  proportionate(age) %>%
  kable(col.names = c("Age", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Region

```{r demo-region}
data %>%
  select(fips, weight) %>%
  mutate(state_abbr = toupper(name_to_abbr(state_fips_to_name(fips)))) %>%
  inner_join(select(state_census, REGION, ABBR),
             by = c("state_abbr" =  "ABBR")) %>%
  proportionate(REGION) %>%
  kable(col.names = c("Region", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Household size

```{r demo-hh-size}
gc()

data %>%
  select(A2b, A5_1, A5_2, A5_3, wave, weight) %>%
  mutate(
    A2b = as.integer(A2b),
    A5_1 = as.integer(A5_1),
    A5_2 = as.integer(A5_2),
    A5_3 = as.integer(A5_3),
    hh_size = case_when(
      wave < 4 ~ A2b,
      wave >= 4 ~ ifelse(is.na(A5_1) + is.na(A5_2) + is.na(A5_3) < 3,
                         ifelse(is.na(A5_1), 0L, A5_1) +
                           ifelse(is.na(A5_2), 0L, A5_2) +
                           ifelse(is.na(A5_3), 0L, A5_3),
                         NA_integer_)
    ),
    hh_size_bucket = case_when(
      hh_size == 1 ~ "1",
      hh_size == 2 ~ "2",
      hh_size >= 3 & hh_size <= 5 ~ "3-5",
      hh_size >= 6 & hh_size <= 10 ~ "6-10",
      hh_size > 10 ~ ">10"
      )) %>%
  proportionate(hh_size_bucket) %>%
  kable(col.names = c("Household size", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Date of completion

```{r demo-completion-date}
gc()

data %>%
  select(StartDatetime, weight) %>%
  mutate(date = as.Date(StartDatetime),
         quarter = paste0(year(date), "Q", quarter(date))) %>%
  proportionate(quarter) %>%
  kable(col.names = c("Quarter", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

### Education

ACS educational attainment data is for adults age 25 and older.

```{r demo-education}
data %>%
  filter(D2 >= 2) %>%
  select(D8, weight) %>%
  mutate(education = case_when(
    D8 == 1 ~ "Less than high school",
    D8 == 2 ~ "High school or equivalent",
    D8 == 3 ~ "Some college, no degree",
    D8 == 4 ~ "Associate's degree",
    D8 == 5 ~ "Bachelor's degree",
    D8 %in% 6:8 ~ "Graduate or professional degree"
  )) %>%
  proportionate(education) %>%
  kable(col.names = c("Education", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

This does not match the ACS data (our respondents are more educated), so it may
be valuable to look at this breakdown monthly. If the bias is consistent over
time, at least we know it is not messing up time trends.

```{r demo-education-time}
data %>%
  filter(D2 >= 2) %>%
  select(D8, StartDatetime, weight) %>%
  mutate(date = as.Date(StartDatetime),
         month = paste0(year(date), "M", month(date)),
         education = case_when(
           D8 == 1 ~ "Less than high school",
           D8 == 2 ~ "High school or equivalent",
           D8 == 3 ~ "Some college, no degree",
           D8 == 4 ~ "Associate's degree",
           D8 == 5 ~ "Bachelor's degree",
           D8 %in% 6:8 ~ "Graduate or professional degree"
         )) %>%
  group_by(education, month) %>%
  summarize(sum_weight = sum(weight), .groups = "drop") %>%
  group_by(month) %>%
  mutate(prop = 100 * sum_weight / sum(sum_weight)) %>%
  ungroup() %>%
  select(month, education, prop) %>%
  pivot_wider(names_from = month, values_from = prop)
```

### Comorbidities

The paper only reports on "at least 1 comorbidity".

```{r comorbidities}
data %>%
  filter(!is.na(C1)) %>%
  select(C1, weight) %>%
  mutate(at_least_one = case_when(
    C1 == "9" ~ "None",
    TRUE ~ "At least one comorbidity"
  )) %>%
  proportionate(at_least_one) %>%
  kable(col.names = c("Comorbidities", "Number", "Unweighted %", "Weighted %"),
        digits = 1, format.args = list(big.mark = ","))
```

## Trends in each signal, by Census region

Figure 2:

```{r signal-anosmia-trends, message=FALSE}
gc()

start_day <- "2020-04-06"
end_day <- "2021-04-05"

anosmia <- data %>%
  select(StartDatetime, B2, fips, weight) %>%
  filter(!is.na(fips)) %>%
  mutate(date = as.Date(StartDatetime),
         anosmia = is_selected(split_options(B2), "13"),
         state = toupper(name_to_abbr(state_fips_to_name(fips)))) %>%
  group_by(date, state) %>%
  summarize(anosmia = 100 * weighted.mean(anosmia, weight, na.rm = TRUE),
            sample_size = sum(!is.na(anosmia))) %>%
  select(time_value = date, value = anosmia, geo_value = state, sample_size) %>%
  as.covidcast_signal(signal = "anosmia", geo_type = "state")

signals <- covidcast_signals(
  c("fb-survey", "fb-survey", "jhu-csse", "hhs", "jhu-csse"),
  c("raw_wcli", "raw_whh_cmnty_cli", "confirmed_incidence_prop",
    "confirmed_admissions_covid_1d_prop", "deaths_incidence_prop"),
  start_day, end_day,
  geo_type = "state")

signals <- append(signals, list(anosmia)) %>%
  aggregate_signals(format = "long")

signals$signal <- factor(
  signals$signal,
  levels = c("anosmia", "raw_wcli", "raw_whh_cmnty_cli", "confirmed_incidence_prop",
             "confirmed_admissions_covid_1d_prop", "deaths_incidence_prop"),
  labels = c("Anosmia %", "CLI %", "Community CLI %", "Cases per 100,000",
             "Hospitalizations per 100,000", "Deaths per 100,000"))

to_plot <- signals %>%
  mutate(geo_value = toupper(geo_value)) %>%
  inner_join(select(state_census, ABBR, REGION, POPESTIMATE2019),
             by = c("geo_value" = "ABBR")) %>%
  group_by(REGION, time_value, signal) %>%
  summarize(value = weighted.mean(value,
                                  ifelse(data_source %in% c("jhu-csse", "hhs"),
                                         POPESTIMATE2019, sample_size),
                                  na.rm = TRUE))

to_plot <- to_plot %>%
  filter(REGION != "X") %>%
  group_by(REGION, signal) %>%
  arrange(time_value) %>%
  mutate(value = rollmean(value, 7, fill = NA))

ggplot(to_plot, aes(x = time_value, y = value, color = REGION)) +
  geom_line() +
  facet_wrap(vars(signal), nrow = 2, ncol = 3, scales = "free_y") +
  labs(x = "Date", y = "Value", color = "Census Region") +
  scale_x_date(date_breaks = "3 months",
               date_labels = "%b %Y",
               date_minor_breaks = "1 month",
               guide = guide_axis(angle = 45)) +
  theme_bw() +
  theme(legend.position = "bottom")

ggsave("../paper/fig/signal-time-series.pdf",
       width = 7, height = 5)
```

## Occupation

Table columns:

- % tested positive: The % of respondents in each occupation category who have
  ever tested positive. P(ever tested positive | occupation group). Note no
  conditioning on testing, so respondents who were never tested count as not
  having tested positive.
- % working with symptoms: In each occupation category, % of respondents who (a)
  have worked outside their home in the past few days but (b) have unusual
  symptoms. That is, P(symptoms & worked outside home | occupation group)
- % working outside/not tested: In each occupation category, % of respondents
  who (a) have worked outside their home in the past 4 weeks but (b) have never
  been tested. (There are some pretty stark disparities in this table). That is,
  P(never tested & worked outside home | occupation group).

Calculated on all responses in January 2020 with the relevant items completed.

**Note:** From here on, all analyses will be for Wave 4 onward (since Wave 4 has
the needed variables).

```{r occupation}
data <- data %>%
  filter(wave >= 4)

data$C13_logic <- split_options(data$C13)

data %>%
  filter(!is.na(Q64),
         StartDatetime >= "2021-01-01 00:00:00",
         StartDatetime < "2021-02-01 00:00:00") %>%
  mutate(
    ever_pos = case_when(
      B10a == "1" ~ TRUE,
      B11 == "1" ~ TRUE,
      B11 == "2" ~ FALSE,
      B8 == "2" ~ FALSE,
      TRUE ~ NA
    ),
    work_sym = case_when(
      is.na(C13) ~ NA, # No info on activities outside home
      is.na(B2) ~ NA, # No info on symptoms, unusual or not
      B2 == "15" ~ FALSE, # No symptoms, unusual or not
      is.na(B2c) ~ FALSE, # No unusual symptoms listed
      !is_selected(C13_logic, "1") ~ FALSE, # Did not work outside home
      is_selected(C13_logic, "1") ~ TRUE, # Worked outside home
      TRUE ~ NA
    ),
    work_not_tested = case_when(
      D10 == "2" ~ FALSE,
      B8 == "1" ~ FALSE,
      D10 == "1" & B8 == "2" ~ TRUE,
      TRUE ~ NA
    ),
    occupation = occupations[Q64]) %>%
  select(occupation, ever_pos, work_sym, work_not_tested, weight) %>%
  group_by(occupation) %>%
  summarize(work_pos_rate = 100 * weighted.mean(ever_pos, weight, na.rm = TRUE),
            work_sym_rate = 100 * weighted.mean(work_sym, weight, na.rm = TRUE),
            work_not_tested_rate = 100 * weighted.mean(work_not_tested, weight, na.rm = TRUE),
            n = n(),
            .groups = "drop_last") %>%
  kable(col.names = c("Occupation group", "% tested positive",
                      "% working with symptoms", "% working outside/not tested",
                      "N"),
        digits = 1, format.args = list(big.mark = ","))
```

## Symptoms

```{r symptom-rates}
data$positive_14d <- case_when(
  data$B10a == 1 ~ "Tested positive",
  TRUE ~ "All others"
)

symptoms <- list(
  "1" = "Fever",
  "2" = "Cough",
  "3" = "Shortness of breath",
  "4" = "Difficulty breathing",
  "5" = "Tiredness or exhaustion",
  "6" = "Nasal congestion",
  "7" = "Runny nose",
  "8" = "Muscle or joint aches",
  "9" = "Sore throat",
  "10" = "Chest pain",
  "11" = "Nausea or vomiting",
  "12" = "Diarrhea",
  "13" = "Loss of smell/taste",
  "16" = "Eye pain",
  "17" = "Chills",
  "18" = "Headaches",
  "19" = "Changes in sleep",
  "14" = "Other"
)

## P(unusual symptoms | tested positive in past 14 days, age group)
symptom_data <- data %>%
  select(positive_14d, B2, B2c, weight)

symptom_data$B2c_logic <- split_options(symptom_data$B2c)

for (ii in names(symptoms)) {
  sym_name <- paste0("sym_", symptoms[[ii]])

  symptom_data[[sym_name]] <- case_when(
    symptom_data$B2 == "15" ~ FALSE, # no symptoms, unusual or not
    is.na(symptom_data$B2) ~ NA, # left symptoms blank
    is.na(symptom_data$B2c) ~ FALSE, # symptom given in B2, no unusual symptoms
    TRUE ~ is_selected(symptom_data$B2c_logic, ii)
  )
}

symptom_rates_overall <- symptom_data %>%
  summarize(across(starts_with("sym_"), ~ weighted.mean(.x, weight, na.rm = TRUE))) %>%
  pivot_longer(cols = starts_with("sym_"), names_to = "symptom",
               names_prefix = "sym_", values_to = "rate")

symptom_rates <- symptom_data %>%
  group_by(positive_14d) %>%
  summarize(across(starts_with("sym_"), ~ weighted.mean(.x, weight, na.rm = TRUE))) %>%
  pivot_longer(cols = starts_with("sym_"), names_to = "symptom",
               names_prefix = "sym_", values_to = "rate")
```

Here are overall symptoms among all respondents:

```{r symptoms-overall-table}
symptom_rates_overall %>%
  mutate(rate = 100 * rate) %>%
  arrange(desc(rate)) %>%
  kable(col.names = c("Symptom", "% with symptom"),
        digits = 1)
```

Here we have unusual symptoms, split by whether the respondent reported testing
positive in the past 14 days. Calculate risk ratios and their CIs.

```{r symptoms-compare}
# Reorder factor levels to order by symptoms among those who tested positive
ordered_symptoms <- symptom_rates %>%
  filter(positive_14d == "Tested positive") %>%
  arrange(desc(rate)) %>%
  pull(symptom)
symptom_rates$symptom <- factor(
  symptom_rates$symptom,
  levels = ordered_symptoms)

# Get relative risks using epitools. We need raw counts, not weighted estimates,
# unfortunately.
symptom_counts <- symptom_data %>%
  group_by(positive_14d) %>%
  summarize(across(starts_with("sym_"), ~ sum(.x, na.rm = TRUE))) %>%
  pivot_longer(cols = starts_with("sym_"), names_to = "symptom",
               names_prefix = "sym_", values_to = "rate")

n_positive <- sum(symptom_data$positive_14d == "Tested positive", na.rm = TRUE)
n_negative <- sum(symptom_data$positive_14d == "All others", na.rm = TRUE)

symptom_rrs <- symptom_counts %>%
  pivot_wider(id_cols = symptom, names_from = positive_14d, values_from = rate)

symptom_rrs$rr <- NA
symptom_rrs$rr_lower <- NA
symptom_rrs$rr_upper <- NA
for (ii in seq_len(nrow(symptom_rrs))) {
  res <- riskratio(c(
    n_negative - symptom_rrs[[ii, "All others"]], # without symptom, negative
    symptom_rrs[[ii, "All others"]], # with symptom, negative
    n_positive - symptom_rrs[[ii, "Tested positive"]], # without symptom, positive
    symptom_rrs[[ii, "Tested positive"]]))$measure

  symptom_rrs$rr[ii] <- res["Exposed2", "estimate"]
  symptom_rrs$rr_lower[ii] <- res["Exposed2", "lower"]
  symptom_rrs$rr_upper[ii] <- res["Exposed2", "upper"]
}
```

Now plot (Figure 1):

```{r symptoms-compare-plot}
rr_scale_factor <- 70

ggplot(symptom_rates, aes(x = symptom, y = rate, fill = positive_14d)) +
  geom_col(position = position_dodge()) +
  geom_pointrange(
    aes(x = symptom, y = rr / rr_scale_factor,
        ymin = rr_lower / rr_scale_factor,
        ymax = rr_upper / rr_scale_factor),
    data = symptom_rrs,
    inherit.aes = FALSE) +
  scale_y_continuous(labels = label_percent(accuracy = 1),
                     sec.axis = sec_axis(~ . * rr_scale_factor,
                                         name = "Relative risk")) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(x = "", y = "Frequency among respondents",
       fill = "Group") +
  theme_bw() +
  theme(legend.position = "bottom")

ggsave("../paper/fig/symptoms-comparison.pdf")
```

In table form, so we can quote it:

```{r symptoms-table}
symptom_rates %>%
  mutate(rate = 100 * rate) %>%
  pivot_wider(names_from = "positive_14d", values_from = "rate") %>%
  select(symptom, positive = `Tested positive`, others = `All others`) %>%
  arrange(desc(positive)) %>%
  kable(col.names = c("Symptom", "Tested positive", "All others"),
        digits = 1)
```

And the table of relative risks and their CIs:

```{r rr-table}
symptom_rrs %>%
  select(symptom, rr, rr_lower, rr_upper) %>%
  arrange(desc(rr)) %>%
  kable(col.names = c("Symptom", "RR", "lower", "upper"),
        digits = 2)
```

## Cumulative test positivity

Figure 3. Based on code from Minttu Rönn.

We use 7-day average cumulative incidence (per 100,000 population) to match the
7-day averaging used in the survey estimates. The survey data will have one
estimate per week, rather than per day, and so when we join the two data frames
we'll get one JHU case rate per week.

```{r case-rate-compare, message=FALSE}
# Estimates of the Total Resident Population and Resident Population Age 18
# Years and Older for the United States, States, and Puerto Rico: July 1, 2019
# (SCPRC-EST2019-18+POP-RES)
state_pops <- read.csv("state-age-pops.csv")

# State-level age distributions of cases, from CDC data
case_age_dist <- read.csv("case-age-distribution.csv")

start_day <- floor_date(as.Date(min(data$StartDatetime)), unit = "week")
state_case_rate <- covidcast_signal("jhu-csse", "confirmed_7dav_cumulative_num",
                                    start_day = start_day,
                                    end_day = as.Date(max(data$StartDatetime)),
                                    geo_type = "state") %>%
  mutate(geo_value = toupper(geo_value),
         year = year(time_value),
         month = month(time_value)) %>%
  select(state = geo_value, cumulative_cases = value, time_value, year, month) %>%
  inner_join(state_pops, by = "state") %>%
  inner_join(case_age_dist, by = c("year", "month")) %>%
  mutate(case_rate_adult = cumulative_cases * frac_adult_cases / adult_pop * 100)

state_case_compare <- data %>%
  filter(!is.na(fips)) %>%
  mutate(
    ever_pos = case_when(
      B10a == "1" ~ TRUE,
      B11 == "1" ~ TRUE,
      B11 == "2" ~ FALSE,
      B8 == "2" ~ FALSE,
      TRUE ~ NA
    ),
    week = floor_date(as.Date(StartDatetime), unit = "week"),
    state_abbr = name_to_abbr(state_fips_to_name(fips))
  ) %>%
  group_by(week, state_abbr) %>%
  summarize(ever_pos = 100 * weighted.mean(ever_pos, weight, na.rm = TRUE),
            .groups = "drop") %>%
  left_join(state_case_rate,
            by = c("state_abbr" = "state", "week" = "time_value"))

g <- ggplot(state_case_compare) +
  facet_geo(~ state_abbr, grid = "us_state_with_DC_PR_grid1", label = "code") +
  geom_line(aes(x = week, y = case_rate_adult, color = "Surveillance"),
            linetype = 1, size = 0.5) +
  geom_line(aes(x = week, y = ever_pos, color = "CTIS"),
            linetype = 1, size = 0.5) +
  scale_color_manual(values = c("lightblue", "gray50")) +
  scale_y_continuous(breaks = c(0, 10),
                     minor_breaks = c(5, 15),
                     labels = label_percent(scale = 1,
                                            accuracy = 1)) +
  scale_x_date(date_breaks = "4 months", date_labels = "%b",
               date_minor_breaks = "1 month") +
  labs(x = "", y = "", color = "") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("../paper/fig/cumulative-case-compare.pdf", g, width = 10, height = 7)

g
```

We should also calculate correlations between state reports and survey estimates
during the last week of data:

```{r case-rate-correlations}
max(state_case_compare$week)

final_week <- state_case_compare %>%
  filter(week == max(week))

# States with the lowest proportion of ever-diagnosed in survey data
final_week %>%
  arrange(ever_pos) %>%
  select(state_abbr, ever_pos, case_rate_adult) %>%
  head()

# States with the highest proportion
final_week %>%
  arrange(desc(ever_pos)) %>%
  select(state_abbr, ever_pos, case_rate_adult) %>%
  head()

# Pearson correlation between surveillance and CTIS
cor(final_week$case_rate_adult, final_week$ever_pos)

# Spearman (rank) correlation between surveillance and CTIS
cor(final_week$case_rate_adult, final_week$ever_pos,
    method = "spearman")
```

## Contacts, masks, and public transport

Figure 4. Based on code from Alyssa Bilinski.

```{r external-data}
make_df <- function(f, var = SVI_cat) {
  var <- ensym(var)

  f %>%
    select(!!var, week, C13_options, C14, contacts_tot, weight) %>%
    filter(!is.na(!!var)) %>%
    group_by(!!var, week) %>%
    summarize(
      pub_trans = 100 * weighted.mean(is_selected(C13_options, "6"),
                                      weight, na.rm = TRUE),
      masks_not_always = 100 * weighted.mean(
        case_when(
          C14 == 1 ~ FALSE,
          C14 %in% 2:5 ~ TRUE,
          C14 == 6 ~ NA), # exclude those not in public in past 5/7 days
        weight, na.rm = TRUE),
      contacts_avg_mod = weighted.mean(contacts_tot, weight, na.rm = TRUE)) %>%
    pivot_longer(c(contacts_avg_mod, pub_trans, masks_not_always),
                 names_to = "var", values_to = "value") %>%
    mutate(var_cat = case_when(
      var == "contacts_avg_mod" ~ "Average contacts (past 24h)",
      var == "pub_trans" ~ "Rode public transit (past 24h, %)",
      var == "masks_not_always" ~ "Not always masked in public (%)"))
}

# We don't actually need populations -- just region codes
pops <- state_census %>%
  select(state = NAME, region = REGION)

# SVI
svi <- read.csv("data/SVI2018_US_COUNTY.csv") %>%
  mutate(SVI_cat = case_when(RPL_THEMES >= 0.75 ~ "1st quartile",
                             RPL_THEMES >= 0.5 & RPL_THEMES < 0.75 ~ "2nd quartile",
                             RPL_THEMES >= 0.25 & RPL_THEMES < 0.5 ~ "3rd quartile",
                             RPL_THEMES < 0.25 ~ "4th quartile"),
         FIPS = sprintf("%05d", FIPS)) %>%
  select(FIPS, SVI_cat)
```

Produce the variables and data:

```{r contacts-data}
contacts_data <- data %>%
  # Exclude ludicrous numbers of contacts, since they screw up the means
  filter(chk_work_contacts & chk_shopping_contacts & chk_social_contacts & chk_other_contacts) %>%
  mutate(
    C13 = ifelse(!is.na(C13), C13, C13b), # Replaced in Wave 10; treat as identical
    C14 = ifelse(!is.na(C14), C14, C14a), # Replaced in Wave 8; treat as identical
    C13_options = split_options(C13),
    state = state_fips_to_name(fips),
    week = floor_date(as.Date(StartDatetime), unit = "week"),
    any_work = ifelse(is.na(C10_1_1), 0, C10_1_1),
    any_shopping = ifelse(is.na(C10_2_1), 0, C10_2_1),
    any_social = ifelse(is.na(C10_3_1), 0, C10_3_1),
    any_other = ifelse(is.na(C10_4_1), 0, C10_4_1),
    chk = !is.na(C10_1_1) | !is.na(C10_2_1) | !is.na(C10_3_1) | !is.na(C10_4_1),
    work_mod2 = ifelse(chk, any_work, NA),
    social_mod2 = ifelse(chk, any_social, NA),
    shopping_mod2 = ifelse(chk, any_shopping, NA),
    other_mod2 = ifelse(chk, any_other, NA),
    contacts_tot = (as.numeric(work_mod2) + as.numeric(shopping_mod2) +
                      as.numeric(social_mod2) + as.numeric(other_mod2)),
    age_cat = case_when(D2 <= 2 ~ "18-34",
                        D2 %in% 3:5 ~ "35-64",
                        D2 > 5 ~ "65+")
  ) %>%
  left_join(svi, by = c("fips" = "FIPS")) %>%
  left_join(pops, by = c("state"))
```

Produce the plots:

```{r contacts-plots}
g1 <- make_df(contacts_data, var = SVI_cat)
g2 <- make_df(contacts_data, var = age_cat)
g3 <- make_df(contacts_data, var = region)

# Pull the legends up towards the plots to eliminate wasted space. ggplot's
# default unit is half a line height, and a line height is 11 pt, so work in
# multiples of this.
legend_top_margin <- -11 * 3 / 2

a <- ggplot(g1, aes(x = week, y = value, col = SVI_cat, group = SVI_cat)) +
  geom_line() +
  facet_wrap(. ~ var_cat, scales = "free", ncol = 4) +
  scale_color_discrete(name = "SVI") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  labs(x = "", y = "") +
  ylim(0,  NA) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.margin = margin(legend_top_margin, 0, 11 / 2, 0),
        plot.margin = margin(11 / 2, 11 / 2, 0, 11 / 2))

b <- ggplot(g2, aes(x = week, y = value, col = age_cat, group = age_cat)) +
  geom_line() +
  facet_wrap(. ~ var_cat, scales = "free", ncol = 4) +
  scale_color_discrete(name = "Age") +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  labs(x = "", y = "") +
  ylim(0, NA) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.margin = margin(legend_top_margin, 0, 11 / 2, 0),
        plot.margin = margin(11 / 2, 11 / 2, 0, 11 / 2))

c <- ggplot(g3, aes(x = week, y = value, col = region, group = region)) +
  geom_line() +
  facet_wrap(. ~ var_cat, scales = "free", ncol = 4) +
  scale_color_discrete(name = "Region")  +
  scale_x_date(date_breaks = "3 months", date_labels = "%b %Y",
               date_minor_breaks = "1 month") +
  labs(x = "", y = "") +
  ylim(0, NA) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.margin = margin(legend_top_margin, 0, 11 / 2, 0),
        plot.margin = margin(11 / 2, 11 / 2, 0, 11 / 2))

g <- arrangeGrob(a, b, c, nrow = 3)

ggsave("../paper/fig/contacts-masks-transport.pdf", g, width = 8, height = 5)
```

## Vaccine acceptance and hesitancy

Figure 5B. Based on code from Marissa Reitsma.

```{r vaccine-hesitancy-map}
# Get data into the form Marissa's code expects
df <- data %>%
  select(StartDatetime, fips, V1, V3, weight) %>%
  mutate(date = as.Date(StartDatetime)) %>%
  filter(!is.na(fips),
         date >= as.Date("2021-03-01"),
         date <= as.Date("2021-04-05")) %>%
  mutate(hesitant = case_when(
             V1 == 1 ~ FALSE,
             V3 %in% 1:2 ~ FALSE,
             V3 %in% 3:4 ~ TRUE))

# Get all FIPSen
fips <- as.data.table(tidycensus::fips_codes)
fips <- fips[, fips:=paste0(state_code, county_code)]

# Collapse into summaries. This uses megacounty logic to get summaries by state
# if sample size per county is inadequate
df <- merge(df, fips, by = "fips", all.y = TRUE)

# Compute percent vaccinated or accepting
df <- df[, sample_size := .N, by = c("fips")]
df <- df[, geo_id := ifelse(sample_size >= 50, fips, state_code)]

df <- df[, pct_hesitant := weighted.mean(hesitant, weight, na.rm = TRUE), by = c("geo_id")]
df <- df[, pct_hesitant := mean(pct_hesitant, na.rm = TRUE), by = c("geo_id")]

# Collapse the dataset
df <- unique(df[, .(pct_hesitant, geo_id, fips, sample_size)])

# Code the colors
df <- df[pct_hesitant < 0.1, plot_var := "<10%"]
df <- df[pct_hesitant >= 0.1 & pct_hesitant < 0.14, plot_var := "10-14%"]
df <- df[pct_hesitant >= 0.14 & pct_hesitant < 0.18, plot_var := "14-18%"]
df <- df[pct_hesitant >= 0.18 & pct_hesitant < 0.22, plot_var := "18-22%"]
df <- df[pct_hesitant >= 0.22 & pct_hesitant < 0.26, plot_var := "22-26%"]
df <- df[pct_hesitant >= 0.26 & pct_hesitant < 0.3, plot_var := "26-30%"]
df <- df[pct_hesitant >= 0.3, plot_var := "30+%"]

df <- df[, plot_var := factor(plot_var, levels = c("<10%", "10-14%", "14-18%", "18-22%", "22-26%", "26-30%", "30+%"))]

df <- df[, GEOID := fips]

counties <- get_acs(geography = "county", geometry = TRUE,
                    shift_geo = TRUE, variables = "B19013_001")
states <- get_acs(geography = "state", geometry = TRUE,
                  shift_geo = TRUE, variables = "B19013_001")

plot_data <- merge(counties, df, by = "GEOID", all.x = TRUE)

g <- ggplot(data = plot_data, aes(fill = plot_var)) +
  geom_sf(data = plot_data[plot_data$geo_id == plot_data$fips, ],
          size = 0.2, color = "gray") +
  geom_sf(data = plot_data[plot_data$geo_id != plot_data$fips, ],
          size = 0.2, color = "NA") +
  geom_sf(data = states, fill = NA, color = "#444444") +
  scale_fill_manual(
    values = c("#0d2c84", "#225ea8", "#1d91c0", "#41b6c4",
               "#7fcdbb", "#c7e9b4", "#ffffcc", "#dedede"),
    na.value = "#dedede") +
  theme_void() +
  labs(fill = "% hesitant", title = "")

ggsave("../paper/fig/hesitancy-county-map.pdf", g, width = 6, height = 4)

g
```