### 2021-09-27 forecast/paper/supplement.Rmd

---
bibliography: [../../common/covidcast.bib]
output: 
  pdf_document:
    keep_tex: true
    template: pnas-suppl-template.tex
params:
  fd_casefloor: 30
  flag_jumps: 15
---

\input{supplement-text.tex}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6.5, fig.height = 4.5,
                      fig.align = "center",
                      fig.path = "fig/",
                      cache = TRUE, 
                      out.width = "\\textwidth",
                      autodep = TRUE,
                      message = FALSE, 
                      warning = FALSE)
```

```{r source-funs, cache=FALSE}
library(scales)
library(cowplot)
library(covidcast)
library(tidyverse)
library(lubridate)
library(evalcast)
library(here)

source(here("code", "figures", "results_processing_and_plotting.R"))
source(here("code", "figures", "eval_funs.R"))
source(here("code", "figures", "hotspot_funs.R"))

```

```{r data-loading}
actuals <- readRDS(here("data", "confirmed_7dav_incidence_prop.RDS"))
fcasts_honest <- readRDS(here("data", "results_honest.RDS")) %>%
  process_res_cases(actuals) %>%
  filter(forecaster != "gs_inherit") 
fcasts_finalized <- readRDS(here("data", "results_dishonest.RDS")) %>%
  process_res_cases(actuals) %>%
  filter(forecaster != "gs_inherit")
hotspots_honest <- readRDS(here("data", "hotspots_honest.RDS")) %>%
  process_res_hotspots() %>%
  filter(forecaster != "gs_inherit")
hotspots_finalized <- readRDS(here("data", "hotspots_dishonest.RDS")) %>%
  process_res_hotspots() %>%
  filter(forecaster != "gs_inherit")
cases7dav <- readRDS(here("data", "confirmed_7dav_incidence_num.RDS")) %>%
  rename(case_num = value)
if (params$fd_casefloor >= 0) {
  fcasts_honest <- filter_case_floor(fcasts_honest, cases7dav, params)
  fcasts_finalized <- filter_case_floor(fcasts_finalized, cases7dav, params)
}
if (params$flag_jumps > 0) {
  fcasts_honest <- filter_jumps(fcasts_honest, actuals, params)
  fcasts_finalized <- filter_jumps(fcasts_finalized, actuals, params)
  hotspots_honest <- filter_jumps(hotspots_honest, actuals, params)
  hotspots_finalized <- filter_jumps(hotspots_finalized, actuals, params)
}
```

```{r gs-processing}
# only honest
fcasts_gs <- intersect_averagers(
  fcasts_honest, "forecaster", 
  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
  filter(forecaster != "gs_subset") %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
fcasts_honest <- filter(fcasts_honest, forecaster != "gs_subset") %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
fcasts_finalized <- filter(fcasts_finalized, forecaster != "gs_subset")%>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
hotspots_gs <- intersect_averagers(
  hotspots_honest %>% filter(!is.na(value), !is.na(actual)), 
  "forecaster", 
  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
  filter(forecaster != "gs_subset") %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
hotspots_honest <- intersect_averagers( # Deal with DV NaN's
  hotspots_honest %>% 
    filter(! is.na(value), !is.na(actual), forecaster != "gs_subset"), 
  "forecaster", 
  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
hotspots_finalized <- intersect_averagers( 
  hotspots_finalized %>% 
    filter(! is.na(value), !is.na(actual), forecaster != "gs_subset"), 
  "forecaster", 
  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
```

```{r hrr-names}
hrr_names <- read_csv(here("data", "Hospital_Referral_Regions.csv"), 
                      col_types = "_ccc__") %>%
  mutate(state = substr(hrrcity, 1, 2),
         hrrname = paste(HRR_lbl, state)) %>%
  select(hrrnum, hrrname)
hrr_tab <- pull(hrr_names, hrrname)
names(hrr_tab) <- pull(hrr_names, hrrnum)
```

<!-- All figures and tables below this line -->

<!-- Finalized v. vintage -->

```{r fcast, include=FALSE}
# Figure 3 in the paper
fig3 <- plotter(fcasts_honest %>% filter(period != "jm"), 
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
fig3
```

```{r hot, include=FALSE}
# Figure 4 in the manuscript
fig4 <- plotter_hotspots(hotspots_honest %>% filter(period != "jm")) + 
  geom_hline(yintercept = 0.5)
fig4
```

```{r fcast-hot-combo, include=FALSE, fig.width=9, fig.height=4.5}
leg <- get_legend(fig3)
comb1 <- plot_grid(fig3 + theme(legend.position = "none"),
                   fig4 + theme(legend.position = "none"), 
                   nrow = 1)
plot_grid(comb1, leg, ncol = 1, rel_heights = c(1,.1))
```

```{r fcast-finalized, fig.cap="Forecasting performance using finalized data. Compare to Figure 3 in the manuscript."}
plotter(fcasts_finalized %>% filter(period != "jm"), 
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

\clearpage

```{r hot-finalized, fig.cap="Hotspot prediction performance using finalized data. Compare to Figure 4 in the manuscript."}
plotter_hotspots(hotspots_finalized %>% filter(period != "jm")) +
  geom_hline(yintercept = 0.5)
```

\clearpage

```{r fcast-honest-v-finalized, fig.cap="Relative forecast WIS with vintage compared to finalized data. Using finalized data leads to overly optimistic performance."}
plotter(
  left_join(
    fcasts_honest %>% filter(period != "jm") %>% select(forecaster:wis),
    fcasts_finalized %>% filter(period != "jm") %>% select(forecaster:wis) %>%
      rename(finalized_wis = wis)
    ), 
  "wis", Mean, scaler = "finalized_wis", 
  order_of_operations = c("aggr","scale")) +
  geom_hline(yintercept = 1) +
  ylab("Mean WIS (vintage / finalized)")

```

\clearpage

```{r hot-honest-v-finalized, fig.cap="Relative AUC with vintage compared to finalized data. Using finalized data leads to overly optimistic hotspot performance."}
left_join(
  hotspots_honest %>%
    filter(period != "jm") %>%
    group_by(forecaster, ahead) %>%
    summarise(auc_honest = auc(value, actual)),
  hotspots_finalized %>%
    filter(period != "jm") %>%
    group_by(forecaster, ahead) %>%
    summarise(auc_finalized = auc(value, actual))) %>%
  ggplot(aes(ahead, auc_honest / auc_finalized, color = forecaster)) + 
  geom_line() + 
  geom_point() +
  theme_bw() + 
  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow=1)) +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  geom_hline(yintercept = 1) +
  ylab("AUC (vintage / finalized)")
```

\clearpage

<!-- Geometric mean justifications -->

```{r wis-densities, fig.cap="Weighted interval score appears to more closely resemble a log-Gaussian distribution."}
ggplot(fcasts_honest %>% filter(period != "jm"), 
       aes(wis, fill = forecaster)) +
  geom_density() +
  scale_x_log10() +
  theme_bw() +
  xlab("WIS") +
  scale_fill_manual(values = fcast_colors) +
  facet_wrap(~forecaster) +
  theme(legend.position = "none")
```

\clearpage

```{r fcast-adjusted, fig.cap="Relative forecast performance using vintage data and summarizing with the more robust geometric mean."}
plotter(fcasts_honest %>% filter(period != "jm"), 
        "wis", GeoMean, scaler = "strawman_wis", 
        order_of_operations = c("scale","aggr")) +
  ylab("Geometric mean of WIS (relative to baseline)")
```

\clearpage

<!-- Statistical significance -->

```{r sign-test, fig.cap="P-values for a sign test that the WIS of the AR forecaster is smaller than that of the indicator-assisted model. Each P-value corresponds a particular forecast date."}
fcast_colors2 <- fcast_colors[names(fcast_colors) != "AR"]

st <- fcasts_honest %>%
  filter(period != "jm") %>%
  mutate(wis = wis / strawman_wis) %>%
  select(forecaster, geo_value, ahead, forecast_date, wis) %>%
  pivot_wider(names_from = forecaster, values_from = wis) %>%
  mutate(across(AR:`Google-AA`, ~ AR - .x)) %>%
  select(-AR) %>%
  pivot_longer(`CHNG-CLI`:`Google-AA`, names_to = "forecaster", values_to = "dif") %>%
  group_by(forecaster, forecast_date, geo_value) %>%
  summarise(dif = sum(dif)) %>%
  group_by(forecaster, forecast_date) %>%
  summarise(p = binom.test(
    x = sum(dif > 0),
    n = n(),
    alternative = "greater")$p.val)

ggplot(st, aes(p)) +
  geom_histogram(
    aes(y = after_stat(density), color = forecaster, fill = forecaster), 
    alpha = 0.4,
    bins = 40) +
  scale_color_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
  scale_fill_manual(values = fcast_colors2, guide = guide_legend(nrow = 1)) +
  facet_wrap(~forecaster) +
  ylab("Frequency") +
  xlab("P-value") +
  scale_y_continuous(
    labels = scales::percent_format(scale = 1, accuracy = 1)) +
  theme_bw() +
  theme(legend.position = "bottom")

```

```{r dm-test}
errs <- fcasts_honest %>% 
  filter(period != "jm") %>%
  group_by(forecaster, forecast_date) %>%
  summarise(
    `Mean Relative WIS` = mean(wis) / mean(strawman_wis),
    # mwis = mean(wis), 
    `Geometric Mean Relative WIS` = GeoMean(wis / strawman_wis) #,
    # gmwis = GeoMean(wis)
    ) %>%
  ungroup()
ar <- errs %>% 
  filter(forecaster == "AR") %>% 
  select(-forecaster) %>%
  pivot_longer(-forecast_date, names_to = "metric", values_to = "AR")
notar <- errs %>%
  filter(forecaster != "AR") %>% 
  pivot_longer(-c("forecaster", "forecast_date"), names_to = "metric", values_to = "err")

all_errs <- left_join(notar, ar)
dms <- all_errs %>% 
  group_by(forecaster, metric) %>% 
  summarise(dm = forecast::dm.test(AR, err, "greater", h = 7 , power = 1)$p.value)

kableExtra::kable(
  bind_rows(dms) %>% 
    pivot_wider(names_from = forecaster, values_from = dm), 
  digits = 3, booktabs = TRUE, caption = "$P$-values for a Diebold-Mariano test for differences in forecast error. The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.")
```

<!-- Bootstrap stuff -->

```{r bootstrap-loading}
fcasts_booted <- readRDS(here("data","results_honest_bootstrapped.RDS")) %>%
  process_res_cases(actuals) %>%
  filter(forecaster != "gs_inherit")
if (params$fd_casefloor >= 0) {
  fcasts_booted <- filter_case_floor(fcasts_booted, cases7dav, params)
}
hotspots_booted <- readRDS(here("data", "hotspots_honest_bootstrapped.RDS")) %>%
  process_res_hotspots() %>%
  filter(forecaster != "gs_inherit")
if (params$flag_jumps > 0) {
  hotspots_booted <- filter_jumps(hotspots_booted, actuals, params)
  fcasts_booted <- filter_jumps(fcasts_booted, actuals, params)
}
fcasts_booted <- filter(fcasts_booted, forecaster != "gs_subset") %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
hotspots_booted <- intersect_averagers( # Deal with DV NaN's
  hotspots_booted %>% 
    filter(! is.na(value), !is.na(actual), forecaster != "gs_subset"), 
  "forecaster", 
  c("geo_value", "ahead", "forecast_date", "target_end_date")) %>%
  mutate(forecaster = recode(forecaster, gs = "Google-AA"))
```

```{r fcast-booted, fig.cap = "Forecast performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model."}
plotter(fcasts_booted %>% filter(period != "jm"), 
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

\clearpage

```{r fcast-booted-adjusted, fig.cap = "Forecast performance as measured with the geometric mean when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model."}
plotter(fcasts_booted %>% filter(period != "jm"), 
        "wis", GeoMean, scaler = "strawman_wis", 
        order_of_operations = c("scale","aggr")) +
  ylab("Geometric mean of WIS (relative to baseline)")
```

\clearpage

```{r hot-booted, fig.cap = "Hotspot prediction performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model."}
plotter_hotspots(hotspots_booted %>% filter(period != "jm")) + 
  geom_hline(yintercept = 0.5)
```

\clearpage

<!-- Upswings vs. downswings -->

```{r up-down-processing}
preds <- readRDS(here("data", "predictions_honest.RDS"))
preds <- preds %>% 
  filter(abs(quantile - 0.5) < 1e-6) %>% 
  select(-quantile) %>%
  mutate(forecaster = recode(forecaster,
                             AR3 = "AR",
                             AR3CHCLI3 = "CHNG-CLI",
                             AR3CHCOV3 = "CHNG-COVID",
                             AR3DVCLI3 = "DV-CLI",
                             AR3FBCLI3 = "CTIS-CLIIC",
                             AR3GSSAA3_Subset = "gs_subset",
                             AR3GSSAA3_Zero = "Google-AA",
                             Baseline = "strawman"
  )) %>%
  filter(! forecaster %in% c("gs_subset", "strawman"))

up_down <- actuals %>%
  arrange(geo_value, target_end_date) %>%
  group_by(geo_value) %>%
  left_join(cases7dav, c("target_end_date" = "time_value", "geo_value")) %>%
  mutate(pct_change = (actual - lag(actual,7)) / lag(actual,7),
         udf = case_when(
           pct_change >= .25 ~ "up",
           pct_change <= -.2 ~ "down",
           TRUE ~ "flat")) %>%
  filter(case_num > 30, !is.na(pct_change)) #%>%
  #rename(forecast_date = target_end_date)

corr_df <- left_join(fcasts_honest, preds, 
                     by = c("geo_value", "ahead", "forecaster", 
                            "forecast_date")) %>%
  select(-actual_fd, -starts_with("strawman"), -ae) %>%
  rename(med = value)
rm(preds)

up_down_df <- left_join(corr_df %>% 
                  ungroup() %>%
                  filter(forecaster != "AR"),
                corr_df %>% 
                  ungroup() %>%
                  filter(forecaster == "AR") %>%
                  select(-forecaster, -actual_td) %>%
                  rename(AR_wis = wis, AR_med = med)) %>%
  inner_join(up_down, by = c("geo_value","target_end_date"))
```

```{r upswing-summary, fig.cap="Average difference between the WIS of the AR model and the WIS of the other forecasters. The indicator-assisted forecasters do best during down and flat periods."}
# Figure 5 in the manuscript
up_down_df_summary <- up_down_df %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(mean = Mean(AR_wis - wis))
up_down_df_summary %>%
  ggplot(aes(udf, mean , fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
  scale_y_continuous() +
  theme_bw() +
  geom_hline(yintercept = 0) +
  ylab("Mean of WIS(AR) - WIS(F)") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r upswing-histogram, include=FALSE, fig.width=9, fig.height=4.5}
# Figure 4
up_down_df %>% 
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  ggplot(aes(AR_wis - wis, fill = forecaster)) +
  geom_histogram(bins = 100) +
  facet_grid(udf ~ forecaster) +
  theme_bw() +
  ylab("Count") +
  theme(legend.position = "none") +
  scale_fill_manual(values = fcast_colors) +
  scale_y_log10(breaks = c(1,100,10000), 
                labels = trans_format("log10", math_format(10^.x))) +
  coord_cartesian(ylim=c(1,2e4)) +
  xlab("AR WIS - forecaster WIS") +
  geom_vline(xintercept = 0)
```

```{r upswing-histogram-logged, eval = FALSE, fig.cap="Not sure if we want this here. Similar to Figure 5 in the manuscript but taking logs. "}
up_down_df %>% 
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  ggplot(aes((AR_wis + 1) / (wis + 1), fill = forecaster)) +
  geom_histogram(bins = 100) +
  facet_grid(udf ~ forecaster) +
  theme_bw() +
  ylab("Count") +
  theme(legend.position = "none") +
  scale_fill_manual(values = fcast_colors) +
  scale_x_log10() +
  scale_y_log10(breaks = c(10,1000,100000), 
                labels = trans_format("log10", math_format(10^.x))) +
  xlab("AR WIS - forecaster WIS") +
  geom_vline(xintercept = 1)
```

\clearpage

```{r hotspots-upswing-downswing, fig.cap="Classification and loglikelihood separated into periods of upswing, downswing, and flat cases. Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods."}
hot_udf <- inner_join(
  hotspots_honest %>% filter(period != "jm"),
  up_down %>% select(geo_value, target_end_date, udf))

cutoff <- mean(hot_udf %>% filter(forecaster == "AR") %>% pull(actual))

con_tab <- hot_udf %>% 
  filter(!is.na(actual)) %>%
  mutate(pred = value > cutoff) %>%
  group_by(forecaster, udf) %>%
  summarise(m = mean(pred != actual)) %>%
  ungroup()

con_tab <- left_join(
  con_tab %>% filter(forecaster != "AR"),
  con_tab %>% filter(forecaster == "AR") %>%
    select(-forecaster) %>%
    rename(mar = m))

llike_tab <- hot_udf %>%
  filter(!is.na(actual)) %>%
  mutate( # kill prob 0-1 predictions
    value = case_when(
      value < 1e-8 ~ 1e-8,
      value > 1-1e-8 ~ 1-1e-8,
      TRUE ~ value
    ),
    llike = (actual == 1) * log(value) + (actual == 0) * log(1 - value)) %>%
  group_by(forecaster, udf) %>%
  summarise(m = mean(llike)) %>%
  ungroup()

llike_tab <- left_join(
  llike_tab %>% filter(forecaster != "AR"),
  llike_tab %>% filter(forecaster == "AR") %>%
    select(-forecaster) %>%
    rename(mar = m))

bind_rows(con_tab %>% mutate(err = "Classification error"), 
          llike_tab %>% mutate(err = "Log likelihood")) %>%
  ggplot(aes(udf, (m - mar) / abs(mar) , fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  facet_wrap(~err) +
  ylab("Change relative to AR") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

\clearpage

<!-- Correlations with lagged actuals -->

```{r cor-with-actuals}
comparison_df <- left_join(
  fcasts_honest %>% filter(forecaster != "AR", period != "jm"),
  fcasts_honest %>% filter(forecaster == "AR", period != "jm") %>%
    select(-c(forecaster,target_end_date)) %>%
    rename(ar3_wis = wis),
  by = c("geo_value", "ahead", "forecast_date")) %>%
  group_by(forecaster, forecast_date, geo_value) %>%
  summarize(wis_ratio = mean(wis) / mean(ar3_wis), .groups = "drop")

pct_chng_df <- readRDS(here(
  "data", 
  "confirmed_7dav_incidence_num.RDS")) %>%
  pct_change(n = 7) %>%
  select(geo_value, time_value, pct_change) %>%
  filter(time_value %in% unique(comparison_df$forecast_date))

comparison_df <- comparison_df %>% 
  left_join(pct_chng_df, 
            by = c("forecast_date" = "time_value", "geo_value"))

pct_chng_df <- comparison_df %>% 
  group_by(forecaster, geo_value) %>%
  summarize(
    cor_wisratio_pctchange = cor(wis_ratio, pct_change, method = "spearman"),
    cor_abs_wisratio_minus_1_pctchange = cor(
      abs(wis_ratio - 1), pct_change,method = "spearman"), .groups = "drop")
```

```{r cor-wis-ratio, fig.cap="Histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in smoothed case rates relative to 7 days earlier."}
ggplot(pct_chng_df %>% 
         group_by(forecaster) %>% 
         mutate(median = median(cor_wisratio_pctchange)), 
       aes(x = cor_wisratio_pctchange)) +
  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), bins = 50) +
  scale_fill_manual(values = fcast_colors) +
  theme_bw() +
  theme(legend.position = "none") +
  geom_vline(aes(xintercept = median), linetype = "dotted") + 
  geom_vline(xintercept = 0) +
  facet_wrap(~forecaster) + 
  xlab("Spearman correlation") + ylab("Relative frequency") +
  scale_y_continuous(labels = scales::label_percent(accuracy = .1))
```

\clearpage

```{r cor-wis-ratio-m1, eval = FALSE, fig.cap="This is Alden's second set of histograms. Here we have the correlation of the absolute value of WIS ratio - 1 with the percent change in 7dav cases relative to 7 days earlier"}
ggplot(pct_chng_df %>% 
         group_by(forecaster) %>% 
         mutate(median = median(cor_abs_wisratio_minus_1_pctchange)), 
       aes(x = cor_abs_wisratio_minus_1_pctchange)) +
  geom_histogram(aes(y = ..count.. / sum(..count..), fill = forecaster), bins = 50) +
  scale_fill_manual(values = fcast_colors) +
  theme_bw() +
  theme(legend.position = "none") +
  geom_vline(aes(xintercept = median), linetype = "dotted") + 
  geom_vline(xintercept = 0) +
  facet_wrap(~forecaster) + 
  xlab("Spearman correlation") + ylab("Relative frequency") +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1))
```

\clearpage

```{r upswing-corr-table, fig.cap="Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster. In down periods, improvements in forecast risk are highlycorrelated with lower median predictions. The opposite is true in up periods. This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model. This conclusion is stronger in down periods then in up periods."}
up_down_df %>% 
  filter(period != "jm") %>% 
  group_by(forecaster, udf) %>%
  summarise(cor = cor(AR_wis - wis, AR_med - med)) %>%
  ggplot(aes(udf, cor , fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
  theme_bw() +
  geom_hline(yintercept = 0) +
  theme(legend.position = "bottom", legend.title = element_blank())

# up_down_df %>% 
#   filter(period != "jm") %>% 
#   group_by(forecaster, udf) %>%
#   summarise(cor = cor(AR_wis - wis, AR_med - med)) %>%
#   pivot_wider(names_from = forecaster, values_from = cor) %>%
#   kableExtra::kbl(
#     booktabs = TRUE, digits = 2, centering = TRUE,
#     caption = paste("Correlation of the difference in WIS between the AR",
#                    "model with the difference in median predictions. In down",
 #                   "periods, improvements in forecast risk are highly",
#                    "correlated with lower median predictions. The opposite",
#                    "is true in up periods. This suggests, as one might expect",
#                    "that improved performance of the indicator-assisted",
#                    "model is attributable to being closer to the truth then",
#                    "the AR model. This conclusion is stronger in down",
#                    "periods then in up periods."))
```

\clearpage

<!-- Leadingness and laggingness -->

```{r lead-lag-processing}
# see lead-lag-analysis.R for signal downloading and processing
lead_lag <- function(x, y, lag.max = NULL, na.len = 14) {
  if (sum(is.na(x)) >= na.len || sum(is.na(y)) >= na.len) {
    return(c(leading = NA, lagging = NA))
  }
  out <- ccf(x, y, lag.max = lag.max, plot = FALSE, na.action = na.omit)
  clim <- qnorm((1 + .9) / 2) / sqrt(out$n.used)
  lag_vec <- drop(out$lag)
  cc <- drop(out$acf)
  cc <- cc * (cc > clim)
  return(c(leading = mean(cc[lag_vec < 0]),
           lagging = mean(cc[lag_vec > 0])))
}

slide_lead_lag <- function(x, y, min_len = 56, lag.max = NULL, na.len = 14) {
  n <- length(x)
  outlist <- list()
  idx <- 1:min_len
  for (i in min_len:n) {
    outlist[[i]] <- lead_lag(x[idx], y[idx], lag.max, na.len)
    idx <- idx + 1
  }
  outlist <- bind_rows(outlist)
  bind_rows(
    tibble(leading = rep(NA, min_len-1), lagging = rep(NA, min_len -1)),
    outlist)
}

all_sigs <- readRDS(here("data", "all_signals_wide_as_of_2020-05-18.RDS"))

all_sigs_imputed <- all_sigs %>% 
  group_by(geo_value) %>% 
  arrange(time_value) %>%
  mutate(across(contains("value+0"), 
                ~zoo::na.locf(.x, na.rm = FALSE, maxgap = 13))) %>%
  pivot_longer(contains("value+0"))

sigs_cor <- left_join(
  all_sigs_imputed %>% filter(! str_detect(name, "jhu-csse")),
  all_sigs_imputed %>% filter(str_detect(name, "jhu-csse")) %>%
    select(-name) %>%
    rename(cases = value))

lead_lag_metric <- sigs_cor %>% 
  group_by(geo_value, name) %>%
  arrange(time_value) %>%
  group_modify(~ {
    bind_cols(time_value = .x$time_value, slide_lead_lag(.x$value, .x$cases))
  })

lead_lag_metric <- lead_lag_metric %>%
  mutate(
    forecaster = recode(
      name,
      `value+0:chng_smoothed_adj_outpatient_cli` = "CHNG-CLI",
      `value+0:chng_smoothed_adj_outpatient_covid` = "CHNG-COVID",
      `value+0:doctor-visits_smoothed_adj_cli` = "DV-CLI",
      `value+0:fb-survey_smoothed_hh_cmnty_cli` = "CTIS-CLIIC",
      `value+0:google-symptoms_sum_anosmia_ageusia_smoothed_search` = "Google-AA")) %>%
  ungroup() %>%
  select(-name)

df2 <- left_join(
  up_down_df, 
  lead_lag_metric,
  by = c("geo_value","forecaster","target_end_date" = "time_value"))
```

```{r leading-only, include=FALSE, fig.width=6, fig.height=4}
# Figure 6 in the manuscript
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, leadingness, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors) +
  guides(fill = guide_legend(nrow = 1)) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = c(1,-.15), 
        legend.justification = "right",
        legend.title = element_blank(), 
        legend.text = element_text(size = 10),
        plot.margin = margin(b = 20, l = 4),
        legend.background = element_rect(fill = "transparent"))
```

```{r leading-and-lagging, eval=FALSE}
# Deprecated
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  pivot_longer(leadingness:laggingness) %>%
  ggplot(aes(udf, value, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors) +
  facet_wrap(~name) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r ccf-dv-finalized, eval=TRUE, fig.cap="Illustration of the cross-correlation function between DV-CLI and cases. The left panel shows the standardized signals over the period from August 1 to September 28 (as of May 15, 2021). The right panel shows $\\CCF_{\\ell}(a)$ for different values of $a$ as vertical blue bars. The orange dashed lines indicate the 95\\% significance threshold. By our leadingness/laggingness metric, DV-CLI is leading (but notlagging) cases over this period."}
source(here("code", "figures", "ccf-dv-finalized.R"))
gg
```

\clearpage

```{r lagging-only, fig.cap="Correlation of the difference in WIS with the  laggingness of the indicator at the target date, stratified by up, down, or flat period. Compare to Figure 5 in the manuscript."}
# Deprecated
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(
    leadingness = cor(AR_wis - wis, leading, use = "pairwise.complete.obs"),
    laggingness = cor(AR_wis - wis, lagging, use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, laggingness, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors[1:5], guide = guide_legend(nrow = 1)) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

\clearpage

```{r diff-in-lead-lag, fig.cap = "Correlation of the difference between leadingness and laggingness with the difference in WIS. The relationship is essentially the same as described in the manuscript and shown in Figure 5."}
# Deprecated
df2 %>%
  filter(period != "jm") %>%
  group_by(forecaster, udf) %>%
  summarise(cor = cor(AR_wis - wis, leading - lagging, 
                      use = "pairwise.complete.obs")) %>%
  ggplot(aes(udf, cor, fill = forecaster)) +
  geom_bar(width = 0.6, position = position_dodge(width=0.6),
           stat = "identity") +
  geom_hline(yintercept = 0) +
  scale_fill_manual(values = fcast_colors[1:5]) +
  theme_bw() +
  ylab("Correlation") +
  xlab("") +
  theme(legend.position = "bottom", legend.title = element_blank())
```

\clearpage

<!-- Figures in the manuscript in their own scripts -->

```{r revisions-dv-jhu, eval=FALSE}
# Figure 1 in the manuscript
source(here("code", "figures", "revisions.R"))
```

```{r ny-trajectory, eval=FALSE}
# Figure 2 in the manuscript
source(here("code", "figures", "trajectory.R"))
gg
```

<!-- Examining data in 2021 -->

```{r fcast-alldates, fig.cap="Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021."}
plotter(fcasts_honest,
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

\clearpage

```{r fcast-alldates-adjusted, fig.cap="Forcast performance over all periods aggregaged with the geometric mean. Again, the inclusion of data in 2021 leads to improved performance."}
plotter(fcasts_honest, 
        "wis", GeoMean, scaler = "strawman_wis", 
        order_of_operations = c("scale","aggr")) +
  ylab("Geometric mean of WIS (relative to baseline)")
```

\clearpage

```{r hot-alldates, fig.cap="Area under the curve for hotspot predictions including data in 2021. Performance degrades relative to the period in 2020. However, there are far fewer hotspots during this period as case rates declined in much of the country."}
plotter_hotspots(hotspots_honest)
```

\clearpage

<!-- Cumulative Mean -->

```{r cumulative-mean, fig.cap="This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model. The shaded background shows" }
cumulatives <- fcasts_honest %>%
  group_by(forecaster, forecast_date) %>%
  summarise(mw = Mean(wis),
            smw = Mean(strawman_wis),
            gmw = GeoMean(wis / strawman_wis)) %>%
  arrange(forecast_date) %>%
  mutate(`Cumulative Mean` = cummean(mw) / cummean(smw),
         `Cumulative Sum` = cumsum(mw) / cumsum(smw),
         `Cumulative Geo Mean` = exp(cummean(log(gmw))),
         `GM Regret` = cumsum(gmw) - cumsum(smw/smw),
         `AM Regret` = cumsum(mw) - cumsum(smw),
         `14 day trailing average` = RcppRoll::roll_meanl(mw / smw, n = 14L))

us_cases <- covidcast::covidcast_signal(
  "jhu-csse", "confirmed_7dav_incidence_num", "2020-03-01", "2021-01-01",
  geo_type = "nation") %>% as_tibble() %>%
  mutate(value = (value / max(value)) * .25 +  .75) %>%
  filter(time_value > "2020-06-01")

ggplot(cumulatives %>% filter(forecast_date < "2021-01-01"), 
       aes(forecast_date, color = forecaster)) + 
  geom_line(aes(y = `Cumulative Sum`)) +
  annotate(x = us_cases$time_value, ymax = us_cases$value, ymin = -Inf,
           geom = "ribbon", alpha = .1) +
  geom_hline(yintercept = 1) +
  xlab("forecast date") +
  scale_color_manual(values = fcast_colors, guide = guide_legend(nrow = 1)) +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

```{r errs-in-space, fig.cap="Percent improvement in WIS relative to the AR forecaster by region (negative numbers indicate improved performance, positives indicate worsening)."}
# fig.height = 9.6, fig.width = 6, out.width = "5in", out.height="8in", 
by_hrr <- fcasts_honest %>% 
  filter(period != "jm") %>%
  group_by(forecaster, geo_value) %>% 
  summarize(value = GeoMean(wis)) %>%
  mutate(time_value = Sys.Date()) %>%
  ungroup()
by_hrr <- left_join(
  by_hrr %>% filter(forecaster != "AR"), 
  by_hrr %>% 
    filter(forecaster == "AR") %>% 
    rename(AR = value) %>%
    select(geo_value, AR)) %>%
  mutate(value = (value - AR) / AR * 100) %>%
  group_by(forecaster) %>%
  group_split()

cc_sigs <- purrr::map(by_hrr,
                   ~ as.covidcast_signal(
                     .x, signal = "Improvement relative to AR",
                     data_source = .x$forecaster[1],
                     geo_type = "hrr"))
br <- c(2.5, 5, 10, 25, 50)
br <- c(-rev(br), 0, br)
colours <- brewer_pal(palette = "RdBu")

maps <- purrr::map(
  cc_sigs, 
  ~plot(.x, title = .x$forecaster[1], 
        choro_params = list(breaks = br), 
        choro_col = rev(colours(11))) +
    theme(legend.position = "none"))

map_plot <- cowplot::plot_grid(plotlist = maps, ncol = 3)
map_leg <- get_legend(maps[[1]] + theme(legend.position = "bottom"))
cowplot::plot_grid(map_plot, map_leg, ncol = 1, rel_heights = c(1, .1))
```

\clearpage

<!-- All code below is deprecated -->

```{r fcast-gs-locations, eval=FALSE}
plotter(fcasts_gs %>% filter(period != "jm"),
        "wis", Mean, scaler = "strawman_wis", 
        order_of_operations = c("aggr","scale")) +
  ylab("Mean WIS (relative to baseline)")
```

```{r fcast-gs-locations-adjusted, eval=FALSE}
plotter(fcasts_gs %>% filter(period != "jm"), 
        "wis", GeoMean, scaler = "strawman_wis", 
        order_of_operations = c("scale","aggr")) +
  ylab("Geometric mean of WIS (relative to baseline)")
```

```{r hot-gs-locations, eval=FALSE}
plotter_hotspots(hotspots_gs %>% filter(period != "jm")) +
  geom_hline(yintercept = 0.5)
```

```{r traj-data, eval = FALSE}
source(here("code", "deprecated", "trajectory_plot_funs.R"))
preds <- readRDS(here("data", "predictions_honest.RDS"))
traj_best <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab, 
                                  "only2020", "best")
traj_worst <- get_trajectory_plots(fcasts_honest, preds, actuals, hrr_tab,
                                   "only2020", "worst")
rm(preds)
```

```{r trajectory-plots, eval = FALSE}
for (nm in names(traj_best)) print(trajectory_panel(nm, traj_best, traj_worst))
```