### 2021-09-27 forecast/paper/supplement.tex

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}

\input{../../common/prelim.tex}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
  Prediction?}

\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison J. Hu, Nat DeFries,
  Sangwon Hyun, Natalia L. Oliveira, James Sharpnack, Jingjing Tang, Robert
  Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J. Tibshirani} 
\correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}

\begin{document}

\maketitle

\SItext

\input{supplement-text.tex}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/fcast-finalized-1}

}

\caption{Forecasting performance using finalized data. Compare to Figure 3 in the manuscript.}\label{fig:fcast-finalized}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/hot-finalized-1}

}

\caption{Hotspot prediction performance using finalized data. Compare to Figure 4 in the manuscript.}\label{fig:hot-finalized}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/fcast-honest-v-finalized-1}

}

\caption{Relative forecast WIS with vintage compared to finalized data. Using finalized data leads to overly optimistic performance.}\label{fig:fcast-honest-v-finalized}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/hot-honest-v-finalized-1}

}

\caption{Relative AUC with vintage compared to finalized data. Using finalized data leads to overly optimistic hotspot performance.}\label{fig:hot-honest-v-finalized}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/wis-densities-1}

}

\caption{Weighted interval score appears to more closely resemble a log-Gaussian distribution.}\label{fig:wis-densities}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/fcast-adjusted-1}

}

\caption{Relative forecast performance using vintage data and summarizing with the more robust geometric mean.}\label{fig:fcast-adjusted}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/sign-test-1}

}

\caption{P-values for a sign test that the WIS of the AR forecaster is smaller than that of the indicator-assisted model. Each P-value corresponds a particular forecast date.}\label{fig:sign-test}
\end{figure}

\begin{table}

\caption{\label{tab:dm-test}$P$-values for a Diebold-Mariano test for differences in forecast error. The test is for the null hypothesis of equal performance against the alternative that the indicator assisted model is better.}
\centering
\begin{tabular}[t]{lrrrrr}
\toprule
metric & CHNG-CLI & CHNG-COVID & CTIS-CLIIC & DV-CLI & Google-AA\\
\midrule
Geometric Mean Relative WIS & 0.141 & 0.076 & 0.000 & 0.059 & 0.008\\
Mean Relative WIS & 0.186 & 0.137 & 0.002 & 0.074 & 0.064\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/fcast-booted-1}

}

\caption{Forecast performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model.}\label{fig:fcast-booted}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/fcast-booted-adjusted-1}

}

\caption{Forecast performance as measured with the geometric mean when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model.}\label{fig:fcast-booted-adjusted}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/hot-booted-1}

}

\caption{Hotspot prediction performance when indicators are replaced with samples from their empirical distribution. Performance is largely similar to the AR model.}\label{fig:hot-booted}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/upswing-summary-1}

}

\caption{Average difference between the WIS of the AR model and the WIS of the other forecasters. The indicator-assisted forecasters do best during down and flat periods.}\label{fig:upswing-summary}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/hotspots-upswing-downswing-1}

}

\caption{Classification and loglikelihood separated into periods of upswing, downswing, and flat cases. Like the analysis of the forecasting task in the main paper (see Figure 7), performance is better during down and flat periods.}\label{fig:hotspots-upswing-downswing}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/cor-wis-ratio-1}

}

\caption{Histograms of the Spearman correlation between the ratio of AR to AR WIS with the percent change in smoothed case rates relative to 7 days earlier.}\label{fig:cor-wis-ratio}
\end{figure}

\clearpage

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/upswing-corr-table-1}

}

\caption{Correlation of the difference in WIS with the difference in median predictions for the AR model relative to the indicator-assisted forecaster. In down periods, improvements in forecast risk are highlycorrelated with lower median predictions. The opposite is true in up periods. This suggests, as one might expect that improved performance of the indicator-assisted model is attributable to being closer to the truth then the AR model. This conclusion is stronger in down periods then in up periods.}\label{fig:upswing-corr-table}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/ccf-dv-finalized-1}

}

\caption{Illustration of the cross-correlation function between DV-CLI and cases. The left panel shows the standardized signals over the period from August 1 to September 28 (as of May 15, 2021). The right panel shows $\CCF_{\ell}(a)$ for different values of $a$ as vertical blue bars. The orange dashed lines indicate the 95\% significance threshold. By our leadingness/laggingness metric, DV-CLI is leading (but notlagging) cases over this period.}\label{fig:ccf-dv-finalized}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/lagging-only-1}

}

\caption{Correlation of the difference in WIS with the  laggingness of the indicator at the target date, stratified by up, down, or flat period. Compare to Figure 5 in the manuscript.}\label{fig:lagging-only}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/diff-in-lead-lag-1}

}

\caption{Correlation of the difference between leadingness and laggingness with the difference in WIS. The relationship is essentially the same as described in the manuscript and shown in Figure 5.}\label{fig:diff-in-lead-lag}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/fcast-alldates-1}

}

\caption{Forecast performance over all periods. Performance largely improves for all forecasters with the inclusion of data in 2021.}\label{fig:fcast-alldates}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/fcast-alldates-adjusted-1}

}

\caption{Forcast performance over all periods aggregaged with the geometric mean. Again, the inclusion of data in 2021 leads to improved performance.}\label{fig:fcast-alldates-adjusted}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/hot-alldates-1}

}

\caption{Area under the curve for hotspot predictions including data in 2021. Performance degrades relative to the period in 2020. However, there are far fewer hotspots during this period as case rates declined in much of the country.}\label{fig:hot-alldates}
\end{figure}

\clearpage

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/cumulative-mean-1}

}

\caption{This figure shows the cumulative sum of WIS for each forecaster divided by the cumulative sum of WIS for the baseline model. The shaded background shows}\label{fig:cumulative-mean}
\end{figure}

\begin{figure}

{\centering \includegraphics[width=\textwidth]{fig/errs-in-space-1}

}

\caption{Percent improvement in WIS relative to the AR forecaster by region (negative numbers indicate improved performance, positives indicate worsening).}\label{fig:errs-in-space}
\end{figure}

\clearpage

\FloatBarrier

\dataset{predictions.zip}{Archived {\tt .RDS} (R objects) files containing
  all predictions for forecasting and hotspots using vintage and finalized data.
Persistent DOI to be added at publication.}

\dataset{evaluations.zip}{Archived {\tt .RDS} (R objects) files containing
  all evaluations for forecasting and hotspots using vintage and finalized data.
  Persistent DOI to be added at publication.}

\dataset{analysis.zip}{Archived {\tt .RDS} (R objects) files containing
  additional data used to produce graphics and conclusions in the manuscript.
  Persistent DOI to be added at publication.}

\dataset{code.zip}{R script files containing
  all code used to reproduce the analysis described in the manuscript.
  Persistent DOI to be added at publication.}

\bibliography{../../common/covidcast.bib}

\end{document}