### 2021-10-05 forecast/paper/supplement-text.tex
Pass over supplement
- Just formatting and copyedits
- Comment out all eval results from 2021 (but I can't seem to
  compile the Rmd because I'm missing the data files)

\section{Finalized Versus Vintage Data}

The goal of this section is to quantify the effect of not properly accounting
for the question of ``what was known when'' in performing retrospective
evaluations of forecasters.  Figures \ref{fig:fcast-finalized} and
\ref{fig:hot-finalized} show what Figures 3 and 4 in the main paper would have
looked like if we had simply trained all models using the finalized data rather
than using vintage data.  This comparison can be seen more straightforwardly in
Figures \ref{fig:fcast-honest-v-finalized} and \ref{fig:hot-honest-v-finalized},
which show the ratio in performance between the vintage and finalized versions.
When methods are given the finalized version of the data rather than the version
available at the time that the forecast would have been made, all methods appear
(misleadingly) to have better performance than they would have had if run
prospectively.  For example, for forecasting case rates 7-days ahead, the WIS of
all methods is at least 8\% larger than what would have been achieved using
finalized data.  This effect diminishes as the forecasting horizon increases,
reflecting the fact that longer-horizon forecasters rely less heavily on recent
data than very short-horizon forecasters.  Crucially, some methods are
``helped'' more than others by the less scrupulous retrospective evaluation,
underscoring the difficulty of avoiding misleading conclusions when performing
retrospective evaluations of forecasters.

\chngcli~(and, to a lesser extent, the other claims-based signals) is the most
affected by this distinction, reflecting the latency in claims-based reporting.
This underscores the importance of efforts to provide ``nowcasts'' for claims
signals (which corresponds to a 0-ahead forecast of what the claims signal's
value will be once all data has been collected). Looking at the \chngcli~and
\dv~curves in Figure \ref{fig:fcast-finalized}, we can see that they perform
very similarly when trained on the finalized data.  This is reassuring because
they are, in principle, measuring the same thing (namely, the percentage of
outpatient visits that are primarily about COVID-related symptoms).  The
substantial difference in their curves in Figure 3 of the main paper must
therefore reflect their having very different backfill profiles.

While using finalized rather than vintage data affects \dv~the least for
forecasting, it is one of the most affected methods for the hotspot problem.
This is a reminder that the forecasting and hotspot problems are fundamentally
distinct.  For example, the hotspot problem does not measure the ability to
distinguish between flat and downward trends.

Even the \ar~model is affected by this distinction, reflecting the fact that the
case rates themselves (i.e., the response values) are also subject to revision.
The forecasters based on indicators are thus affected both by revisions to the
indicators and by revisions to the case rates.  In the case of the \gs~model, in
which we only used finalized values for the \gs~indicator, the difference in
performance can be wholly attributed to revisions of case rates.

\section{Aggregation with Geometric Mean}

In this section, we consider using the geometric mean instead of the arithmetic
mean when aggregating the weighted interval score (WIS) across location-time
pairs.  There are three reasons why using the geometric mean may be desirable.
\begin{enumerate}
\item WIS is right-skewed, being bounded below by zero and having occasional 
  very large values.  Figure \ref{fig:wis-densities} illustrates that the
  densities appear roughly log-Gaussian.  The geometric mean is a natural choice  
  in such a context since the relative ordering of forecasters is determined by
  the arithmetic mean of the {\em logarithm} of their WIS values.
\item In the main paper, we report the ratio of the mean WIS of a forecaster to
  the mean WIS of the baseline forecaster. Another choice could be to take the
  mean of the ratio of WIS values for the two methods. This latter choice would
  penalize a method less for doing poorly where the baseline forecaster also
  does poorly.  Using instead the geometric mean makes the order of aggregation
  and scaling immaterial since the ratio of geometric means is the same as the 
  geometric mean of ratios.
\item If one imagines that a forecaster's WIS is composed of multiplicative
  space-time effects $S_{\ell,t}$ shared across all forecasters,
  i.e., $\WIS(F_{\ell,t,f},Y_{\ell,t})=S_{\ell,t}E_{f,t}$ with $E_{f,t}$ a
  forecaster-specific error, then taking the ratio of two forecasters' geometric
  mean WIS values will effectively cancel these space-time effects. 
\end{enumerate}

Figure \ref{fig:fcast-adjusted} uses the geometric mean for aggregation.
Comparing this with Figure 3 of the main paper, we see that the main conclusions
are largely unchanged; however, \chngcli~now appears better than \ar.  This
behavior would be expected if \chngcli's poor performance is attributable to a
relatively small number of large errors (as opposed to a large number of
moderate errors).  Indeed, Figure 5 of the main paper further corroborates this,
in which we see the heaviest left tails occuring for \chngcli.

\section{Statistical Significance}

In the Introduction of the main paper, we give some reasons that we avoid making
formal statements about statistical significance, preferring instead to examine
the stability of our results in different contexts. There are strong reasons to
avoid model-based significance tests because the necessary assumptions about
stationarity, independence, and the  model being true (or at least approximately
true) are certainly violated. With those caveats in mind, we undertake two
relatively assumption-lean investigations in this section. The first is a
sign-test for whether the difference between the AR model’s relative WIS  and
each other model’s relative WIS is centered at zero. By ``relative WIS'' we mean
scaled by the strawman as displayed in Figure 3. To mitigate the dependence
across time (which intuitively seems to matter more than that across space), we
computed these tests in a stratified way, where for each forecast date we run a
sign test on the scaled errors between two models over all 440 counties. The
results are plotted as histograms in Figure \ref{fig:sign-test}. In this case,
we use the total relative WIS over all aheads, but the histograms are largely
similar for individual target horizons. If there were no difference, we would
expect to see a uniform distribution. However, for all indicator-assisted
models, we see many more small p-values than would be expected if the null
hypothesis (that the AR model is better) were true.

Another relatively assumption-lean method of testing for differences in forecast
accuracy is the Diebold-Mariano test \cite{Diebold:2002, Diebold:2015,
Harvey:1997}. Essentially, the differences between forecast losses are assumed
to have a constant mean and a covariance that depends on time. Under these
conditions, the asymptotic distribution for the standardized mean of the
differences is limiting normal provided that a heteroskedasticity and
autocorrelation robust estimate of the variance is used. Using the loss as
weighted interval score across all HRRs and horizons (7 to 21 days ahead), we
perform the DM test using both the mean relative to the strawman (as reported in
the manuscript) and the geometric mean relative to the strawman as described
above. The first two rows of Table \ref{tab:dm-test} displays p-values for the
test that the indicator-assisted model is no-better than the AR model. Only the
CHNG-CLI model's p-value exceeds conventional statistical significance
thresholds.

\section{Bootstrap Results}

As explained in Section 2.B. of the main paper, a (somewhat cynical) hypothesis
for why we see benefits in forecasting and hotspot prediction is that the
indicators are not actually providing useful information but they are instead
acting as a sort of ``implicit regularization,'' leading to shrinkage on the
autoregressive coefficients and therefore to less volatile predictions.  To
investigate this hypothesis, we consider fitting ``noise features'' that in
truth should have zero coefficients.  Recall (from the main paper) that at each
forecast date, we train a model on 6,426 location-time pairs.  Indicator models
are based on six features, corresponding to the three autoregressive terms and
the three lagged indicator values.  To form noise indicator features, we replace
their values with those from a randomly chosen time-space pair (while keeping
the autoregressive features fixed).  In particular, at each location $\ell$ and
time $t$, for the forecasting task we replace the triplet $(X_{\ell,t},
X_{\ell,t-7}, X_{\ell,t-14})$ in Eq. (3) of the main paper with the triplet
$(X_{\ell^*,t^*}, X_{\ell^*,t^*-7}, X_{\ell^*,t^*-14})$, where $(\ell^*,t^*)$ is
a location-time pair sampled with replacement from the 6,426 location-time
pairs.  Likewise in the hotspot prediction task, we replace the triplet
$(X_{\ell,t}^\Delta, X_{\ell,t-7}^\Delta, X_{\ell,t-14}^\Delta)$ in Eq. (5) of
the main paper with $(X_{\ell^*,t^*}^\Delta, X_{\ell^*,t^*-7}^\Delta,
X_{\ell^*,t^*-14}^\Delta)$.  Figures
\ref{fig:fcast-booted}--\ref{fig:hot-booted} show the results.  No method
exhibits a noticeable performance gain over the \ar~method, leading us to
dismiss the implicit regularization hypothesis.

\section{Upswings and Downswings}

In this section we provide extra details about the upswing / flat / downswing
analysis described in the main text. Figure \ref{fig:upswing-summary} shows the
overall results, examining the average difference $\WIS(AR) - \WIS(F)$ in
period. Figure \ref{fig:hotspots-upswing-downswing} shows the same information
for the hotspot task. On average, during downswings and flat periods, the
indicator-assisted models have lower classification error and higher log
likelihood than the AR model. For hotspots, both Google-AA and CTIS-CLIIC
perform better than the AR model during upswings, in contrast to the forecasting
task, where only Google-AA improves. For a related analysis, Figure
\ref{fig:cor-wis-ratio} shows histograms of the Spearman correlation (Spearman's
$\rho$, a rank-based measure of association) between the $\WIS(F)/\WIS(AR)$ and
the magnitude of the swing. Again we see that case rate increases are positively
related to diminished performance of the indicator models.

One hypothesis for diminished relative performance during upswings is that the
AR model tends to overpredict downswings and underpredict upswings. Adding
indicators appears to help avoid this behavior on the downswing but not as much
on upswings. Figure \ref{fig:upswing-corr-table} shows the correlation between
between $\WIS(AR) - \WIS(F)$ and the difference of their median forecasts.
During downswings, this correlation is large, implying that improved relative
performance of $F$ is related to making lower forecasts than the AR model. The
opposite is true during upswings. This is largely to be expected. However, the
relationship attenuates in flat periods and during upswings. That is, when
performance is better in those cases, it may be due to other factors than simply
making predictions in the correct direction, for example, narrower confidence
intervals.

\section{Leadingness and Laggingness}

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness at location $\ell$, we use the cross correlation
function (CCF) between the two time series. The $\CCF_\ell(a)$ of an indicator
$X_{\ell}$ and case rates $Y_{\ell}$ is defined as
% \[
% \CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
%   \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
%   \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
% \]
% where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
% shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
% are the sample mean and
% standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
their Pearson correlation where $X_\ell$ has been aligned with the values of
$Y_{\ell}$ that occurred $a$ days earlier.  Thus, for any $a>0$,
$\CCF_{\ell}(a)>0$ indicates that $Y_{\ell,t}$ is moving together with
$X_{\ell,t+a}$.  In this case we say that $X_{\ell}$ is lagging $Y_{\ell}$. For
$a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is positively correlated with
$X_{\ell,t-a}$, so we say that $X_{\ell}$ leads $Y_{\ell}$.

Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
containing Charlotte, North Carolina, from August 1, 2020 until the end of
September. These are the same signals shown in Figure 1 in the manuscript but
using finalized data. To define ``leadingness'' we compute $\CCF_{\ell}(a)$ (as
implemented with the R function {\tt ccf()}) for each $a\in \{-15,\ldots,15\}$
using the 56 days leading up to the target date. This is the same amount of data
used to train the forecasters: 21 days of training data, 21 days to get the
response at $a=21$, and 14 days for the longest lagged value. The orange dashed
horizontal line represents the 95\% significance threshold for correlations
based on 56 observations. Any correlations larger in magnitude than this value
are considered statistically significant under the null hypothesis of no
relationship. We define leadingness to be the sum of the significant
correlations that are leading (those above the dashed line with $a<0$) while
laggingness is the same but for $a>0$. In the figure, there are three
significant correlations on the ``leading'' side (at $a = -5, -4, -3$), so
leadingness will be the sum of those values while laggingness is 0: on September
28 in Charlotte, \dv~is leading cases leading but not lagging.

Figure \ref{fig:lagging-only} shows the correlation between laggingness and the
difference in indicator WIS and AR WIS. Unlike leadingness (Figure 5 in the
manuscript) there is no obvious relationship that holds consistently across
indicators. This is heartening as laggingness should not aid forecasting
performance. On the other hand, if an indicator is more lagging than it is
leading, this may suggest diminished performance.
Figure~\ref{fig:diff-in-lead-lag} shows the correlation of the difference in
leadingness and laggingness with the difference in WIS. The pattern here is
largely similar to the pattern in leadingness described in the manuscript: the
relationship is strongest in down periods and weakest in up periods with the
strength diminishing as we move from down to flat to up for all indicators.

In calculating the CCF and the associated leadingness and laggingness scores, we
have used the finalized data, and we look at the behavior at the target date of
the forecast. That is we are using the same data to evaluate predictive accuracy
as to determine leadingness and laggingness. It should be noted that the
leadingness of the indicator at the time the model is trained may also be
important. Thus, we could calculate separate leadingness and laggingness scores
for the trained model and for the evaluation data and examine their combination
in some way. We do not pursue this combination further and leave this
investigation for future work.

% \section{Examining data in 2021}

% In this section, we investigate the sensitivity of the results to the
% period over which we train and evaluate the models.  In the main
% paper, we end all evaluation on December 31, 2020.  Figures
% \ref{fig:fcast-alldates}--\ref{fig:hot-alldates} show how the
% results would differ if we extended this analysis through March
%   31, 2021. Comparing Figure \ref{fig:fcast-alldates} to Figure 3 of
% the main paper, one sees that as ahead increases most methods now improve
% relative to the baseline forecaster. When compared to other methods, 
% \chngcli~appears much better than
% it had previously; however, all forecasters other than \chngcov~and
% \dv~are performing less well relative to the baseline than before.
% These changes are likely due to the differing nature of the pandemic
% in 2021, with flat and downward trends much more common than upward
% trajectories.  Indeed, the nature of the hotspot prediction problem is
% quite different in this period.  With a 21-day training window, it is
% common for there to be many fewer hotspots in training.

\section{Aggregation Over Time and Space}

Following the suggestion of an anonymous reviewer, we investigate two
disaggregated versions of the main forecasting result shown in Figure 3 on the
manuscript. The first (Figure \ref{fig:cumulative-mean}) displays the cumulative
error (summed over all horizons) of each forecaster divided by the cumulative
error of the baseline.  This perspective should illustrate how the models
perform over time, drawing attention to any nonstationary behavior.  During the
initial increase in cases in July 2020, CTIS-CLIIC and Google-AA gain a lot
relative to the AR model. All the indicators do much better than the AR model
during the following downturn (the ebb of the second wave). The AR model
actually improves over the indicators in October 2020, before losing a bit in
late November.

Figure \ref{fig:errs-in-space} examines the spatial behavior over the entire
period of the indicator-assisted models relative to the AR. For ease of
comparison, we show the percent improvement in each HRR. Negative numbers (blue)
mean that the indicator helped while positives (red) mean that the indicator
hurt forecasting performance. The clear pattern is that in most HRRs, the
indicators improved performance, though usually by small amounts (2.5\%--10\%).
In some isolated HRRs, performance was markedly worse, though there does not
appear to be any particular pattern to these locations.

% \section{Comparisons to state-of-the-art forecasters}

% Since mid-July 2020, a number of forecasting teams (48 in total) have submitted
% 1--4 week ahead state-level case forecasts to the COVID19 Forecast Hub
% \cite{ForecastHub}. To compare those forecasts to those created in this paper,
% we selected only those forecasters who submitted at least 300 forecasts for the
% 2-week-ahead target before
% January 1, 2021 (approximately 1 forecast for each state for 6 weeks). This left
% 32 forecasters, one of which, the COVIDhub-4-week-ensemble, was an exact
% duplicate of another. To make the comparison fair, we scaled by the
% COVIDhub-baseline as in Figure 3 of the manuscript. Figure
% \ref{fig:compare-to-hub-mean} shows all submitted forecasts (available 7, 14,
% and 21 days ahead) overlaid on Figure 3. We have highlighted the
% COVIDhub-ensemble in light blue for comparison. All the models developed in this
% paper beat the baseline, and would rank in the top 8 of 31 submitted forecasts
% over this period.

% A more fair comparison that directly adjusts for spatiotemporal difficulty is to
% use the geometric mean scaled by the baseline. This is the same idea shown in
% \ref{fig:fcast-adjusted}. We have also remade this figure overlaying submitted
% forecasts available on the Forecast Hub. It occasionally happens that a
% submitted forecast achieved a WIS of 0, so in this case we use the WIS$+ 1$
% rather than WIS for all forecasters. The result is shown in
% \ref{fig:compare-to-hub-geomean}. By this metric, the methods employed in this
% paper would rank in the top 2, beating the COVIDhub-ensemble at all forecast
% horizons. The improvement of the best indicator assisted model over the AR model
% is roughly the same as the improvement of the AR model over the COVIDhub-ensemble.