### 2021-10-18 perspective/paper.tex
Perspective piece

\documentclass{article}
\input{ryantibs}

\title{Epidemic Tracking and Forecasting: Lessons Learned from a Tumultuous Year}
\author{Roni Rosenfeld \and Ryan J. Tibshirani}

\begin{document}
\maketitle

\section{Introduction}

Epidemic forecasting has garnered increasing interest in the last decade,
nurtured and scaffolded by various forecasting challenges organized by groups
within the U.S. federal government, including the CDC \citep{Biggerstaff:2016,
  Biggerstaff:2018, Reich:2019}, OSTP \citep{Johansson:2019}, DARPA
\citep{DelValle:2018}, and elsewhere \citep{Ajelli:2018, Viboud:2018}.  In 2017,
after several years of experimentation with flu forecasting in academic groups,
the CDC decided to incorporate influenza forecasting into its normal operations,
including weekly public communications citep{FluSight} and briefing to
higher-ups.  To provide more reliable infrastructure and support for its
forecasting needs, the CDC in 2019 designated two national Centers of Excellence
for Influenza Forecasting, one at the University of Massachusetts at
Amherst\footnote{\url{https://reichlab.io/people}} and one at Carnegie Mellon
University\footnote{\url{https://delphi.cmu.edu/about/center-of-excellence/}}.

Not unrelatedly, the last decade has also seen a rise in the importance of
\emph{digital surveillance} streams in public health, with improving epidemic
tracking and forecasting models being a key application of these data.  Digital
streams, such as search and social media trends, have constituted a large part
of the focus \citep{Ginsberg:2009, Brownstein:2009, Salathe:2012,
Kass-Hout:2013, Santillana:2015, Paul:2017}; however, even more broadly, data
from \emph{auxiliary} streams that operate outside of traditional public health
reporting, such as online surveys, medical devices, or electronic medical
records, have received considerable attention as well \citep{Kass-Hout:2011,
Carlson:2013, Viboud:2014, Smolinski:2015, Santillana:2016, Charu:2017,
Koppeschaar:2017, Yang:2019, Leuba:2020, Radin:2020, Ackley:2020}.

The Carnegie Mellon Delphi group, which the two of us co-lead, has worked in
both of these emerging disciplines---epidemic forecasting, and building relevant
auxiliary signals to aid such forecasting models---since 2012.  In 2020, as the
pandemic broke out, we struggled like many other groups to find ways to
contribute to the national efforts to respond to the pandemic.  We ended up
shifting our focus to lie nearly entirely on the data end of the spectrum,
pursuing several different directions in order to build and make available to
the public a variety of new indicators that reflect real-time COVID-19 activity
in the U.S.  Three papers in this collection describe this work from three
different perspectives.  A fourth describes international work by some of our
collaborators that parallels our group's work on online surveys in the U.S.

\section{Papers in This Collection}

Here is a very brief summary of the papers in this collection.

\begin{enumerate}
\item \citet{Reinhart:2021} describes our group's (ongoing) effort in building
  and maintaining COVIDcast: an open repository of real-time, geographically
  detailed COVID-19 indicators in the U.S.  These indicators (a term we use
  interchangeably with signals) are derived from a diverse set of data sources:
  medical testing devices, medical insurance claims, internet search trends,
  app-based mobility data, and online surveys, among others.  Many indicators
  are demonstrated to have meaningful statistical relationships with what have
  become the pandemic's ``topline'' numbers (reported cases, hospitalizations,
  deaths), whereas others uniquely reflect certain activity (not available in
  other publicly-available data sources) that may drive or affect the spread of
  COVID-19.  The paper demonstrates through a number of examples that the
  indicators in the COVIDcast repository can improve on the timeliness,
  robustness, and scope of traditional public health reporting streams.

\item \citet{McDonald:2021} provides a detailed analysis of whether a core set
  of the indicators in the COVIDcast repository can be used to improve the
  accuracy of COVID-19 short-term forecasting and hotspot detection models.
  This speaks to the quantitative utility of the indicators in a way that is
  directly tied to the benefits observed in relevant downstream modeling tasks.
  The paper finds that time series models (that are already competitive with top
  forecasters from the COVID-19 Forecast Hub \citep{ForecastHub}) improve in
  predictive accuracy when they are supplemented with any of the five indicators
  under consideration.  However, the magnitude of these improvements varies with
  the pandemic's local dynamics at prediction time, and only one indicator
  appears to offer a nontrivial improvement during upswings.

\item \citet{Salomon:2021} focuses on the U.S. COVID-19 Trends and Impact Survey
  (CTIS), an (ongoing) online survey operated by our group, in partnership with
  Facebook.  This is a very rich source of data about the pandemic and its
  effect on people, only partially reflected by the indicators (derived from the
  survey) in the COVIDcast repository; the full data set of individual,
  anonymized survey responses is available to researchers under a data use
  agreement.  The paper presents descriptive analyses that reflect the unique
  value of CTIS as an important supplement to public health reporting, in
  particular, as an instrument to measure key information about behaviors,
  attitudes, economic impacts, and other topics not covered in traditional
  public health streams.

\item \emph{Authors' note to the Editor: this is to be filled out later.  The
    Astley et al. paper is still undergoing a significant revision and its main
    messages may change as a result.}
\end{enumerate}

\section{Lessons Learned}

We now take the opportunity to reflect on some ``lessons learned'' from our work
over the past year and a half.  Some of the observations below are described in
more depth in the papers in the collection, and others extend beyond the papers
in the collection (but we give references to relevant articles with more details
in the discussion below).

\paragraph{Deceptively simple data labels often belie the data's true meaning
  and complexity.}

Labels such as ``COVID-19 cases'' or ``COVID-19 hospitalizations'' hide an
enormous amount of complexity and potential ambiguity, especially when applied
to data at fine geographic and temporal resolutions.  We elaborate on this and
several other examples in what follows.

\begin{itemize}
\item Cases may be lab-confirmed only or also suspected (with the definition of
  ``suspected'' varying across jurisdictions and time); they may be listed by
  date reported on the jurisdiction's website, by date reported to the public
  health authority, by date tested, by specimen collection date, or occasionally
  by symptom onset date (most informative but often unavailable or
  inapplicable). A casual review of many websites of local and state health
  departments suggests there is great heterogeneity in what is being reported
  \citep{Simon:2021}.

\item The term ``hospitalizations'' is used ambiguously, sometimes referring to
  incidence (hospital admissions) and sometimes to prevalence (hospital bed
  occupancy).  These two quantities cannot easily be mapped to one another, 
  because COVID-19 hospital discharges are rarely if ever reported.
  Furthermore, people admitted without a COVID-19 diagnosis may acquire the
  infection and/or the diagnosis at any time during their hospital stay.

\item Hospitalizations may be reported by the location (typically county) of the
  patient's residence, but are more often reported by the county of the
  reporting hospital.  This is an important distinction, as many rural COVID-19
  patients in need of advanced care travel to the nearest secondary or tertiary
  hospital, often at a nearby urban county.  For example, the hospitals in
  Pittsburgh, located in Allegheny County, Pennsylvania (population 1.2
  million), treat patients from throughout a 13 county region in Southwestern
  Pennsylvania (population over 4 million).  For tracking and forecasting
  hospitalization burden then, the geographic units of hospital referral regions
  (HRRs, \citep{DartmouthHRR}) may be most appropriate.  Alas, these units do
  not conform to county boundaries, which complicates the projection of cases to
  hospitalizations.

\item Deaths are usually reported by county where they occurred, which for
  hospitalized patients may differ from their county of residence \citep{NCHS}.

\item Hospitalization or deaths with COVID-19 are significantly different from
  hospitalization or deaths due to COVID-19 (as captured by e.g., a
  COVID-related chief complaint or primary ICD-10 code).  The proportion of the
  two varies significantly by age groups and across time \citep{Fillmore:2021}.

\item Test positivity rates are most often reported by lumping together all
  tests performed regardless of the reason for performing them.  Tests taken
  following positive diagnosis, due to symptoms, or due to being a contact of a
  confirmed case, are all likely to have a much higher positivity rate than that
  of the general population.  Screening tests are most likely to reflect the
  true prevalence in the screened population.  Sadly, very few jurisdictions
  report or even retain the breakdown of the test results by reason for testing,
  thereby losing forever valuable information.

\item Medical insurance claims offer rich, detailed information about COVID-19
  and other health conditions, but are not without their weaknesses.  Claims are
  often not filed until weeks and months after the medical encounter.  As such,
  signals derived from claims are usually subject to regular and considerable
  revisions up to 60 days after a given date of service, because signals must be
  updated each time new claims for that day of service are received (specific
  statistics on how this affects signal values are given below).  This tends to
  make projections challenging, especially at finer geographic units such as
  counties, since there tends to be a high degree of heterogeneity across
  locations.

\item Medical claims contain information about lab tests taken, but not their
  results.  More generally, for understandable HIPAA reasons, medical claims
  contain only information necessary for adjudicating and auditing claims.   
\end{itemize}

Data definitions must be disambiguated, clarified, and made consistent to the
greatest extent possible, and remaining inconsistencies must be documented and
saliently communicated.

\paragraph{Understanding the data generation process is critical for downstream
  applications.}

Both traditional public health surveillance data streams and newer digital
surveillance streams are the result of often complex processes, some having to
do with the underlying health status or activities being monitored, others with
the reporting process itself.  Understanding the entire ``data generation
process'' for each data source can be challenging, but is absolutely essential
for proper modeling and effective use of the data.  Some examples are as
follows.

\begin{itemize}
\item In medical claims, relevant diagnoses and comorbidities may not be
  reflected if they are not directly relevant to the charges incurred.  On the
  other hand, because medical claim coding determines reimbursement levels, some
  codes may be over-represented relative to their medical significance.

\item Some populations and some healthcare settings are not reflected in the
  commercial claims stream.  These include the healthcare systems of the
  Department of Defense, Indian Health Services, Veterans Affairs, prison
  systems, and other systems that do not reimburse by procedure or service, as
  well as Medicare fee-for-service and Medicaid.  This can cause significant
  bias in the signals relative to the prevalence in the general population.

\item Public health reporting data are often subject to backlogs and reporting
  delays, and estimates for any particular date can be revised over time as
  errors are found or additional data becomes available.  During the pandemic,
  audits, corrections, and the clearing of backlogs has frequently resulted in
  huge artificial spikes and drops \citep{ArvisaisAnhalt:2021}.  Data
  aggregators like Johns Hopkins CSSE \citep{Dong:2020} have worked tirelessly
  to correct such anomalies after first publication (they attempt to
  back-distribute a spike or dip, by working with a local authority to figure
  out how this should best be done).

\item Data revisioning (also known as ``backfill'') is pervasive not only in
  traditional public health reporting, but also in many (though not all) digital
  surveillance sources.  As already alluded to above, signals based on medical
  claims typically undergo regular revisions because many claims (on which these 
  signals are based) get submitted and processed late; for many COVID-related
  claims-based signals, the median relative error between initial reports and
  final values is over 10\%, and only after 30 days or so do estimates typically
  match finalized values within 5\% \citep{Reinhart:2021}.  However, the
  systematic nature of these revisions suggests that, with suitable historical
  data, statistical models could be fit to estimate the final values from
  preliminary reports.  By comparison, revisions to public health reports during
  the pandemic (the spikes and dips just described) have been much less
  systematic and much less predictable.

\item Traditionally, public health agencies do not publish provisional data
  until it meets a level of stability.  For example, NCHS data on the percentage
  of deaths due to pneumonia, influenza and COVID-19 is not released until at
  least 20\% of the expected deaths in a jurisdiction have been reported
  \citep{PIMortality}, a process that may take several weeks.  However, for
  modeling and forecasting purposes, even highly provisional data can be very 
  informative, as long as sufficient historical provisional data is collected so
  that the statistical relationships between provisional values and finalized
  values can be modeled.

\item Calendar effects permeate not only the reporting process but also
  health-seeking behavior and the epidemic process itself, with the effects on
  these three processes not easy to disentangle.  Major holidays and other
  national or regional events are associated with significant travel, social
  mixing, and other distinct behaviors affecting disease transmission.  However,
  holidays and weekends also affect health-seeking behavior via reduced
  non-emergency healthcare capacity (doctors' offices and labs being partially
  or fully closed).  Perhaps the strongest calendar effects are on reporting,
  including claim filing and hospital reporting.  Using 1- or 2-week trailing
  averages eliminates weekend effects, but at a cost of reduced temporal
  resolution, and it leaves unsolved the effects of holidays.  A better approach
  might be to explicitly model and correct for calendar effects. 
\end{itemize}

\paragraph{Mandated reporting in a time of emergency can be burdensome and
  inflexible.}

COVID-19 reporting by hospitals, as mandated by HHS during the pandemic,
consisted of many dozens of data elements and imposed a significant burden on
the nation's 6,000 or so hospitals at a time that they were already stretched to
their limits.  It also took a huge effort to formulate, communicate,
disambiguate, and monitor for quality assurance and uniformity of
interpretation.  In light of this, it is not surprising that it took a long
time, and pressure, for most hospitals to comply (near universal compliance was
not achieved until December 2020).  When changes needed to be made to the
collected statistics, an arduous and time consuming process of approvals,
re-formulation, re-communication, re-implementation, and re-assessment had to be
followed.

While some aspects of mandated reporting are likely to remain irreplaceable,
effective alternative surveillance sources can be of great use: they can improve
on the timeliness, scope, robustness, and utility of mandated reporting data,
while being less burdensome to collect.  This is a common theme that runs
through all four papers in the collection, but is perhaps most directly
addressed in \citet{Reinhart:2021} (which focuses on the ecosystem of signals
broadly).  That said, we have far from saturated the utility of auxiliary
surveillance.  Much more needs to be developed in this area in order to usher
epidemic tracking into its next phase of reliability, accuracy, and
transparency.  To us, electronic medical records hold a great promise for
surveillance streams, and we elaborate on this in the next section.

\paragraph{Human behavior and its impact on the progression of epidemics is hard
  to measure and hard to model.}

In the nearly 10 years of government-organized epidemic forecasting exercises in
the U.S., efforts were focused on modeling the natural history and likely
evolution of the pathogen, with adaptation of human behavior playing a secondary
role (if any role at all).  The pandemic demonstrated that our forecasting
models must pay closer attention to reactive human behavior, even more so if we
are to consider interventions.  Unfortunately, many highly relevant aspects of
human behavior are not measured by publicly available data streams, like
compliance with policies and recommendations (with perhaps surveys and mobility
reports providing our best glimpse into these hard-to-observe aspects of
behavior \citep{Bilinski:2021}), and even the very process of promulgating these 
policies and recommendations.  Furthermore, even if we had these data in hand,
incorporating their effects will require significant and new cognitive and
behavioral modeling, with uncertain success.  The tragic breakdown and
fragmentation of trust in governments, public health officials, and healthcare
professionals are perhaps the hardest factors to measure and model, yet they
played an undeniable role in the progression of the current pandemic in the
U.S. and other countries.

\section{The Road Ahead}

The focus of the Delphi group during the initial, critical period of the
pandemic (February 2020 -- March 2021) was on short-term goals: trying to
provide signals, analysis, and decision support to federal, state and local
public health officials, as well as to fellow researchers, data journalists, and
the general public.  In spring of 2021, equipped with the hard lessons learned
during this tumultuous year, we turned our attention back to the original vision
of the Delphi group, and asked ourselves: given where we are and what we know
now, what is needed to be able to take a major step forward in epidemic tracking
and forecasting?  In this section, we list some ideas which we hope will elicit
further public discussion and, most importantly, experimentation. Because our
expertise lies in modeling and forecasting, not in public health surveillance,
our perspective and recommendations are necessarily limited to those aspects of
surveillance that are needed to realize our vision.

\paragraph{EMR as a key missing component for epidemic tracking and
  forecasting.}

The success of nowcasting, analytics and forecasting depends crucially on the
availability of rich, real-time data sources.  In light of the limitations of
mandated reporting discussed above, we must consider the complementary value of
other data sources.  Chief among these are electronic medical records (EMR), as
are being created and used daily by inpatient and outpatient healthcare
providers, medical laboratories, and pharmacies.  The advantage of these data
resources: they are rich, real-time, and already being generated (found ``in the
wild'').  The challenges: they are highly fragmented in the U.S., with its
approximately 6,000 hospitals and 100,000 outpatient care facilities.  One
promising avenue for countering this fragmentation are Health Information
Exchanges (HIEs), which were set up in the early 2000s with the support of the
federal government to coordinate the sharing of healthcare information among
healthcare providers in a given region, and eventually nationally.  The primary
goal of the HIEs has been patient continuity of care, but public health
surveillance is recognized as an additional worthy goal.  In the context of
health surveillance, HIEs hold the promise of reducing fragmentation from a
hundred thousand partners to only a few hundred.

Other formidable challenges to using EMR are legal, ethical, commercial, and
operational.  Who owns the data, who has access to it, and who has use rights
are all complex and often open questions.  An overriding concern is of course
patient privacy.  We must find a way to use this highly promising data for the
common good without compromising the privacy of individuals.  Fortunately, a
technological solution appears feasible, in the form of \emph{federated
  surveillance}. An outline is as follows.

\begin{itemize}
\item A common API is developed for querying all participating EMR custodians.

\item EMR databases are queried daily with an agreed upon set of queries, and
  return aggregated counts.

\item These counts are further aggregated across multiple providers and
  localities, and then fused with all other available data sources to provide
  alerts, nowcasts, and forecasts.

\item These model outputs are then shared back to the contributing EMR
  custodians, as well as to CDC and other federal and state agencies.

\item Done in this way, no personal health information (PHI) ever leaves the
  custodians' premises, while aggregated statistics can be combined to increase
  statistical power, thereby shortening alert latency and improving detection
  and prediction capabilities.
\end{itemize}

A successful example of federated querying (albeit designed for research rather
than real-time tracking and forecasting) has been recently demonstrated in the
UK \citep{Williamson:2020}.

One advantage of this approach to health surveillance is that when a new
emerging health crisis is identified, or when specific syndromic signatures are
discovered (e.g., ageusia and anosmia for COVID-19), a new query can be
developed, approved, and deployed literally overnight, allowing us to ``shine a
light'' on it on very short notice.  This can be contrasted with traditional,
legally-mandated public health reporting, which could take weeks and months to
develop, approve, negotiate, disseminate, implement, monitor, and assure quality
of, as has happened during the current pandemic.  In the slightly longer term,
demonstrating the effectiveness and superiority of federated surveillance could
obviate the need for crisis-time mandated reporting, alleviating the reporting
burden on hospitals and other healthcare providers during these difficult
times.

In an interpandemic period, an important use case for federated surveillance is
detection of trends and anomalies.  A set of queries can be designed to
continuously test for unusual recent spikes or trends in any number of
diagnoses, symptoms, lab tests, or prescriptions.  The aggregation of evidence
across many systems, localities, and data streams will make detection both more
sensitive and more robust.  Such a system would likely have detected the opioid
epidemic years before it was actually noticed.

One open technical challenge with the federated surveillance approach is
semantic heterogeneity: the use of emerging Health IT standards like HL7's FHIR
\citep{HL7FHIR} can enable a unified view of EMR data elements, but different 
healthcare systems often have different operational definitions for supposedly
universal concepts like ``high blood pressure'' or ``low blood oxygenation''.
Combining counts of such events across different healthcare systems may be a bit
like adding apples to oranges.  It will take some work to harmonize semantics
across so many diverse data custodians, but this is both doable and well worth
doing.  Note that this problem is less severe for the many surveillance and
anomaly detection tasks where the focus is on changes in a signal (in a given
location) over time, rather than on its absolute meaning.

\paragraph{Different phases of epidemic surveillance call for different analytic
  tools.}

It is important to discuss analytic needs separately for each of three different 
phases of epidemic surveillance and tracking, since each poses different
technical challenges and requires different analytical tools.

In the \emph{interpandemic phase}, the main activity is threat scanning for
threats, namely monitoring data streams and events throughout the world for
disconcerting developments.  Relevant statistical tools include anomaly
detection and scan statistics to help decide when an epidemiological
investigation is warranted.  While it may be possible to rank the risks of
different outbreak triggering events (like species jumping or point mutations)
in different locations, which could in turn be used to inform surveillance
resource allocation, conventional forecasting has a limited role to play in this
phase, as such events have large inherent uncertainty.

In the \emph{containment phase}, a discovered threat must be intensely
monitored, continuously assessed, and ultimately contained.  The analytical,
data-driven tools required in this phase include real-time estimation of
critical epidemiological parameters such as R0, infection fatality rates, the
incubation period, the serial interval, and so on.  These real-time estimates
are necessarily based on provisional data, highlighting the value of modeling
the data generation process discussed above.  In this phase, forecasting still
has a limited role, since the outbreak is still local, its fundamental dynamics
unclear, and point events can have large consequences down the road.

If containment fails, during the \emph{mitigation phase} the goals of analytics
expand significantly to include informing mitigation policies and planning.
Real-time tracking (nowcasting) and short term-forecasting (a few weeks
ahead) can play critical roles in these activities, and indeed have been the
focus of our group's work since its inception.  While there is still important
work to be done and advances to be made in this area, we believe that these
advances are likely to be incremental until we see major progress in either (1)
supporting data streams (e.g., better standardization and cleaning of public
health reporting data, identification of leading indicators from, say, EMR); or
(2) our collective scientific understanding of the real-world geotemporal
dynamics of epidemics (discussed next).

\paragraph{Useful, reliable longer-term forecasting remains an aspiration.}

Influenza forecasting exercises in the last several years demonstrated that it
is often possible to usefully quantify uncertainty over the remainder of an
ongoing flu season \citep{Reich:2019}.  But this success was based mostly on
observing the behavior of seasonable epidemics over several decades.  To
reliably forecast the progression of pandemics, where relevant historical data
is almost nonexistent, we must have a detailed quantitative understanding of how
different, diverse factors affect disease transmissibility.  Such an 
understanding is currently grossly lacking, as evidenced by our collective
failure to predict \citep{Reich:2021} (or even understand, post-hoc
\citep{Hawre:2021}) the high-level temporal and geographic contours of the main
pandemic waves in the U.S.  Yet this very pandemic, the most instrumented in
human history, is also a rare opportunity to attempt this vital scientific and
technological goal.

\subsection*{Acknowledgements}

\emph{Authors' note: to be filled out later.}

\newpage
\bibliographystyle{plainnat}
\bibliography{../common/covidcast}

\end{document}