### 2021-10-30 forecast/paper/paper.tex
Second push done
- Incorporate sentence update in intro from Jacob
- Take a careful pass through rest of the paper, various small
  edits or copyedits
- Restore some of the original discussion, but try to avoid having
  it sound like "use my method"

\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnasresearcharticle} % Choose template 
% {pnasresearcharticle} = Template for a two-column research article
% {pnasmathematics} %= Template for a one-column mathematics article
% {pnasinvited} %= Template for a PNAS invited submission

\input{../../common/prelim.tex}

\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
  Prediction?}

% Use letters for affiliations, numbers to show equal authorship (if applicable)
% and to indicate the corresponding author

% Writing group
\author[a,1]{Daniel J. McDonald}
\author[b,2]{Jacob Bien}
\author[c,2]{Alden Green}
\author[c,d,2]{Addison J. Hu}

% Middle authors
\author[d]{Nat DeFries}
\author[b]{Sangwon Hyun}
\author[c,d]{Natalia L. Oliveira}
\author[e]{James Sharpnack}
\author[f]{Jingjing Tang}
\author[g,h]{Robert Tibshirani}
\author[c]{Val{\'e}rie Ventura}
\author[c,d]{Larry Wasserman}

% Senior authors
\author[c,d]{Ryan J. Tibshirani}

\affil[a]{Department of Statistics, University of British Columbia}
\affil[b]{Department of Data Sciences and Operations, University of Southern
  California} 
\affil[c]{Department of Statistics \& Data Science, Carnegie Mellon University}
\affil[d]{Machine Learning Department, Carnegie Mellon University}
\affil[e]{Department of Statistics, University of California, Davis}
\affil[f]{Computational Biology Department, Carnegie Mellon University}
\affil[g]{Department of Statistics, Stanford University}
\affil[h]{Department of Biomedical Data Science, Stanford University}

% Please give the surname of the lead author for the running footer
\leadauthor{McDonald}

% Please add a significance statement to explain the relevance of your work
\significancestatement{Validated forecasting methodology should be a vital  
  element in the public health response to any fast-moving epidemic or 
  pandemic. A widely-used model for predicting the future spread of a temporal
  process, based on past data, is an autoregressive (AR) model. While basic,
  such an AR model (properly trained) is already competitive with the top models
  in operational use for COVID-19 forecasting. In this paper, we exhibit five
  auxiliary indicators---based on de-identified medical insurance claims,
  self-reported symptoms via online surveys, and COVID-related Google
  searches---that further improve the predictive accuracy of an AR model in
  COVID-19 forecasting. The most substantial gains appear to be in quiescent   
  times; but the Google search indicator appears to also offer improvements
  during upswings in pandemic activity.   
}

% Please include corresponding author, author contribution and author
% declaration information

\authorcontributions{D.J.M., J.B., A.G., A.J.H., and R.J.T. wrote the manuscript,
produced figures, and developed software necessary to perform the analyses. All
authors performed research and provided intellectual contributions. R.J.T.
acquired funding.}
\authordeclaration{The authors have no competing interests to declare.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be
  addressed. E-mail: daniel@stat.ubc.ca} 
\equalauthors{\textsuperscript{2}J.B., A.G., and A.J.H. contributed equally to 
  this work.}

% At least three keywords are required at submission. Please provide three to
% five keywords, separated by the pipe symbol. 
\keywords{COVID-19 $|$ forecasting $|$ hotspot prediction $|$ time series | 
  digital surveillance}

\begin{abstract}
  Short-term forecasts of traditional streams from public health reporting (such 
  as cases, hospitalizations, and deaths) are a key input to public health
  decision-making during a pandemic. Since early 2020, our research group has
  worked with data partners to collect, curate, and make publicly available
  numerous real-time COVID-19 indicators, providing multiple views of pandemic
  activity in the U.S.  This paper studies the utility of five such
  indicators---derived from de-identified medical insurance claims,
  self-reported symptoms from online surveys, and COVID-related Google search
  activity---from a forecasting perspective. For each indicator, we ask whether
  its inclusion in an autoregressive (AR) model leads to improved predictive
  accuracy relative to the same model excluding it. Such an AR model, without
  external features, is already competitive with many top COVID-19 forecasting
  models in use today. Our analysis reveals that (a) inclusion of each of these
  five indicators improves on the overall predictive accuracy of the AR model;
  (b) predictive gains are in general most pronounced during times in which
  COVID cases are trending in ``flat'' or ``down'' directions; (c) one
  indicator, based on Google searches, seems to be particularly helpful
  during ``up''  trends.      
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

\begin{document}

\maketitle
\thispagestyle{firststyle}
\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, correctly formulating public policy
responses, and rationally planning future public health resource needs.  A
companion paper \cite{Reinhart:2021} describes our research group's efforts,
beginning in April 2020, in curating, and maintaining a database of real-time
indicators that track COVID-19 activity and other relevant phenomena. The
signals (a term we use synonomously with ``indicators'') in this database are
accessible through the COVIDcast API \cite{CovidcastAPI}, as well as associated
R \cite{CovidcastR} and Python \cite{CovidcastPy} packages, for convenient data  
fetching and processing. In the current paper, we quantify the utility provided
by a core set of these indicators for two fundamental prediction tasks:
probabilistic forecasting of COVID-19 case rates and prediction of future
COVID-19 case hotspots (defined by the event that a relative increase in
COVID-19 cases exceeds a certain threshold).

At the outset, we should be clear that our intent in this paper is \textit{not}
to provide an authoritative take on cutting-edge COVID-19 forecasting methods.
Similarly, some authors, e.g., \cite{ioannidis2020a}, have pointed out numerous
mishaps of forecasting during the pandemic and it is not our general intent to
fix them here. Instead, we start with a basic and yet reasonably effective
predictive model for predicting future trends in COVID-19 cases, and present a
rigorous, quantitative assessment of the added value provided by auxiliary
indicators, that are derived from data sources that operate outside of 
traditional public health streams. In particular, we consider five indicators
derived from de-identified medical insurance claims, self-reported symptoms from
online surveys, and COVID-related Google searches.

To assess this value in as direct terms as possible, we base our study around a
very simple basic model: an autoregressive model, in which COVID cases in the
near future are predicted using a linear combination of COVID cases in the near
past. Forecasting carries a rich literature, offering a wide range of
sophisticated techniques, see, e.g., \cite{Hyndman:2018} for a review; however,
we purposely avoid enhancements such as order selection, correction of
outliers/anomalies in the data, and inclusion of regularization or
nonlinearities. Similarly, we do not account for other factors that may well aid
in forecasting, such as age-specific effects, holiday adjustments, and the
effects of public health mandates. All that said, despite its simplicity, the
basic autoregressive model that we consider in this paper exhibits competitive
performance (see the Supplementary Materials for details) with many of the top
COVID-19 case forecasters submitted to the U.S.\ COVID-19 Forecast Hub
\cite{ForecastHub}, which is the official source of forecasts used in public
communications by the U.S. CDC. The strong performance of the autoregressive
model here is in line with the fact that simple, robust models have also
consistently been among the best-performing ones for COVID-19 death
forecasting \cite{Cramer:2021}.

In the companion paper \cite{Reinhart:2021}, we analyze correlations between
various indicators and COVID case rates. These correlations are natural
summaries of the contemporaneous association between an indicator and COVID
cases, but they fall short of delivering a satisfactory answer to the question
that motivates the current article: is the information contained in an indicator
demonstrably useful for the prediction tasks we care about? Note that even
lagged correlations cannot deliver a complete answer. Demonstrating
\textit{utility for prediction} is a much higher standard than simply asking
about correlations; to be useful in forecast or hotspot models, an indicator
must provide relevant information that is not otherwise contained in past values
of the case rate series itself (cf.\ the pioneering work on Granger  causality
\cite{Granger:1969, engle1987a}, as well as the further references given 
below). We assess this directly by inspecting the difference in predictive  
performance of simple autoregressive models trained with and without access to  
past values of a particular indicator.

We find that each of the five indicators we consider--- two based on
COVID-related outpatient visits from medical insurance claims, one on
self-reported symptoms from online surveys, and one on Google searches for
anosmia or ageusia---provide an overall improvement in accuracy when
incorporated into the autorgressive model. This is true both for COVID-19 case 
forecasting and hotspot prediction. Further analysis reveals that the gains in
accuracy depend on the pandemic's dynamics at prediction time: the biggest  
gains in accuracy appear during times in which cases are ``flat'' or trending
``down''; but the indicator based on Google searches offers a most notable
improvement when cases are trending ``up''.

Careful handling of data revisions plays a key role in our analysis. Signals 
computed from surveillance streams are often subject to latency and/or 
revision. For example, a signal based on aggregated medical insurance claims may
be available after just a few days, but it can then be substantially revised
over the next several weeks as additional claims are submitted and/or processed
late. Correlations between such a signal and case rates calculated ``after the
fact'' (i.e., computed retrospectively, using the finalized values of this
signal) will not deliver an honest answer to the question of whether this signal
would have been useful in real time. Instead, we build predictive models using
only the data that would have been available \textit{as of} the prediction date,
and compare the ensuing predictions in terms of accuracy. The necessity of
real-time data for honest forecast evaluations has been recognized in
econometrics for a long time, see \cite{howrey1978a, harvey1983a, mankiw1986a,
  trivellato1986a, diebold1991a, patterson1995a, koenig2003a, croushore2006a,
  clark2009a, faust2009a, croushore2011a}; but it is often overlooked in
epidemic forecasting despite its critical importance, see, e.g.,
\cite{chakraborty2018know}.

Finally, it is worth noting that examining the importance of additional features
for prediction is a core question in inferential statistics and econometrics,
with work dating back to at least \cite{Granger:1969}. Still today, drawing
rigorous inference based on predictions, without (or with lean) assumptions, is
an active field of research from both the applied and theoretical angles;
see, e.g., \cite{Diebold:2002, McCraken:2007, Diebold:2015, Stokes:2017,
  Lei:2018, Rinaldo:2019, Williamsom:2020, Zhang:2020, Dai:2021, Fryer:2021}.
Our take in the current work is in line with much of this literature; however,
in order to avoid making any explicit assumptions, we do not attempt to make
formal  significance statements, and instead, broadly examine the stability of
our conclusions with respect to numerous modes of analysis.

\section{Methods}

\subsection{Signals and Locations}

We consider prediction of future COVID-19 case rates or case hotspots (to be
defined precisely shortly).  By case rate, we mean the case count per 100,000
people (the standard in epidemiology).  We use reported case data aggregated by 
JHU CSSE \cite{Dong:2020}, which, like the auxiliary indicators that we use to
supplement the basic autoregressive models, is accessible through the COVIDcast
API \cite{CovidcastAPI}.

The indicators we focus on provide information not generally available from
standard public health reporting. Among the many auxiliary indicators collected
in the API, we study the following five:  
\begin{itemize}
\item Change Healthcare COVID-like illness (CHNG-CLI): The percentage of
  outpatient visits that are primarily about COVID-related symptoms, based on
  de-identified Change Healthcare claims data.
\item Change Healthcare COVID (CHNG-COVID): The percentage of outpatient visits
  with confirmed COVID-19, based on the same claims data.
\item COVID Trends and Impact Survey COVID-like illness in the community
  (CTIS-CLI-in-community): The estimated percentage of the population who know
  someone in their local community who is sick, based on Delphi's COVID Trends
  and Impact Survey, in partnership with Facebook.
\item Doctor Visits COVID-like illness (DV-CLI): The same as CHNG-CLI, but
  computed based on de-identified medical insurance claims from other health
  systems partners.  
\item Google search trends for anosmia and ageusia (Google-AA): A measure of  
  Google search volume for queries that relate to anosmia or ageusia (loss of
  smell or taste), based on Google's COVID-19 Search Trends data set.
\end{itemize}
We choose these indicators because, conceptually speaking, they measure aspects
of an individual's disease progression that would plausibly precede the  
occurence of (at worst, co-occur  with) the report of a positive COVID-19 test,
through standard public health reporting streams.

For more details on the five indicators (including how these are precisely
computed from the underlying data streams) we refer to
\url{https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html},
and the companion paper on the COVIDcast API and its signals
\cite{Reinhart:2021}. For CTIS in particular, we refer to the companion paper
\cite{Salomon:2021}. For the Google COVID-19 Search Trends data set, see
\cite{GoogleSymptoms}; see also \cite{Klopfen:2020, Vaira:2020} for a
justification of the relevance of anosmia or ageusia to COVID infection.

As for geographic resolution, we consider the prediction of COVID-19 case rates
and hotspots aggregated at the level of an individual \textit{hospital referral
  region} (HRR). HRRs correspond to groups of counties in the United States
within the same hospital referral system. The Dartmouth Atlas of Healthcare 
Policy~\cite{DartmouthHRR}, defines these 306 regions based on a number of
characteristics. They are contiguous regions such that most of the hospital
services for the underlying population are performed by hospitals within the
region. Each HRR also contains at least one city where major procedures
(cardiovascular or neurological) are performed. The smallest HRR has a
population of about 125,000. While some are quite large (such as the one
containing Los Angeles, which has more than 10 million people), generally HRRs  
are much more homogenous in size than the (approximately) 3200 U.S. counties,   
and they serve as a nice middle ground in between counties and states.

HRRs, by their definition, would be most relevant for forecasting hospital
demand.  We have chosen to focus on cases (forecasting and predicting
hotspots) at the HRR level because the indicators considered should be more 
useful in predicting case activity rather than hospital demand, as the former is  
intuitively more contemporaneous to the events that are measured by
the given five indicators. Predicting case rates (and hotspots) at the HRR level
is still a reasonable goal in its own right; and moreover, it could be used to
feed predicted case information into downstream hospitalization models.

\subsection{Vintage Training Data}

In this paper, all models are fit with ``vintage'' training data. This means
that for a given prediction date, say, September 28, 2020, we train models 
using data that would have been available to us \textit{as of}
September 28 (imagine that we can ``rewind'' the clock to September 28 and query
the COVIDcast API to get the latest data it would have had available at that
point in time.)  This is possible because of the COVIDcast API's comprehensive
data versioning system (described in more detail in the companion paper
\cite{Reinhart:2021}).  We also use the \texttt{evalcast} R package
\cite{EvalcastR}, which streamlines the process of training arbitrary
prediction models over a sequence of prediction dates, by constructing the
proper sequence of vintage training data sets.

\begin{figure}[tb!]
  \includegraphics[width=\columnwidth]{fig/revisions.pdf}
  % RJT: Ensuring we start with a "declarative title" :) 
  % https://www.internationalscienceediting.com/how-to-write-a-figure-caption/
  \caption{Revision behavior for two indicators in the HRR containing Charlotte, 
    North Carolina.  Each colored line corresponds to the data as reported on a
    particular date (\textit{as of} dates varying from September 28 through
    October 19). The left panel shows the DV-CLI signal, which was regularly
    revised throughout the period, though the effects fade as we look further
    back in time. In contrast, the right panel shows case rates reported by JHU
    CSSE (smoothed with a 7-day trailing average), which remain ``as reported''
    on September 28, with a spike towards the end of this period, until a major 
    correction is made on October 19, which brings this down and affects all
    prior data as well.}  
  \label{fig:vintage}
\end{figure}

Vintage training data means different things, in practice, for different
signals. The three signals based on medical claims, CHNG-CLI, CHNG-COVID, and
DV-CLI, are typically 3-5 days latent, and subject to a considerable but
regular degree of revision or ``backfill'' after their initial publication date.
The survey-based signal, CTIS-CLI-in-community, is 2 days latent, and rarely
undergoes any revision at all.  The target variable itself, reported COVID-19
case rates, is 1 day latent, and exhibits frequent, unpredictable revisions
after intial publication.  Compared to the pattern of revisions in the medical 
claims signals, which are much more systematic in nature, revisions in case
reports can be highly erratic. Big spikes or other anomalies can occur in the
data as reporting backlogs are cleared, changes in case definitions are made,
etc. Groups like JHU CSSE then work tirelessly to correct such anomalies after
first publication (e.g., they will attempt to back-distribute a spike when a
reporting backlog is cleared, by working with a local authority to figure out
how this should best be done), which can result in very nontrivial revisions.
See Figure~\ref{fig:vintage} for an example.

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then; the latency of 
the signal was always at least one week through 2020.  However, unlike the
claims-based signals, there is no reason for revisions to occur after initial
publication, and furthermore, the latency of the signal is not an unavoidable
property of the data type, so we simply use finalized signal values, with zero
latency, in our analysis.

\subsection{Analysis Tasks}

\begin{table*}[t]
\centering
\caption{Summary of forecasting and hotspot prediction tasks considered in
  this paper.}
\begin{tabular}{l p{2.75in} p{2.75in}}
  \toprule
  & \textbf{Forecasting} & \textbf{Hotspot prediction} \\
  \midrule
  Response variable & $Y_{\ell,t}$ (7-day trailing average of COVID-19 case 
incidence rates, per location $\ell$ and time $t$) & $Z_{\ell,t} =
\indicator{Y_{\ell,t} \geq 1.25 \cdot Y_{\ell,t-7}}$ (indicator that 
$Y_{\ell,t}$ grows by more than 25\% relative to the preceding week) \\ 
  Geographic resolution & Hospital referral region (HRR) & Hospital referral
region (HRR) \\ 
  Forecast period & June 9--December 31, 2020 & June 16--December 31, 2020 \\  
  Model type & Quantile regression & Logistic regression \\
  Evaluation metric & Weighted interval score (WIS) & Area under curve (AUC) \\
  \bottomrule
\end{tabular}
\label{tab:analysis_tasks}
\end{table*}

To fix notation, let $Y_{\ell,t}$ denote the 7-day trailing average of COVID-19
case incidence rates in location (HRR) $\ell$ and at time (day) $t$.  To be
clear, this is the number of new daily reported cases per 100,000
people, averaged over the 7-day period $t-6, \ldots, t$. The first task we
consider---\textit{forecasting}---is to predict $Y_{\ell,t+a}$ for each
``ahead'' value $a=7,\ldots,21$.  The second task---\textit{hotspot
  prediction}---is to predict a binary variable defined in terms of the relative
change of $Y_{\ell,t+a}$ (relative to its value one week prior,
$Y_{\ell,t+a-7}$), again for each $a=7,\ldots,21$.

Why do we define the response variables via 7-day averaging? The short answer 
is robustness: averaging stabilizes the case time series, and moderates
uninteresting artifacts like day-of-the-week effects in the data.  We note
that we can also equivalently view this (equivalent up to a constant factor) as 
predicting the HRR-level case incidence rate \textit{summed} over some 7-day
period in the future, and predicting a binary variable derived from this.

In what follows, we cover more details on our two analysis
tasks. Table~\ref{tab:analysis_tasks} presents a summary.

\paragraph{Dynamic Retraining}

For each prediction date $t$, we use a 21-day trailing window of data to train
our forecast or hotspot prediction models (so, e.g., the trained models will
differ from those at prediction date $t-1$).  This is done to account for
(potential) nonstationarity.  For simplicity, the forecasting and hotspot
prediction models are always trained on data across all HRRs (i.e., the
coefficients in the models do not account for location-specific effects).

\paragraph{Prediction Period}

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data and lagged values 
of signals for features, as we will detail shortly) June 9, 2020 for
forecasting, and June 16, 2020 for hotspot prediction. (The bottleneck here is
the CTIS-CLI-in-community signal, which does not exist before early April 2020,
when the survey was first launched).

\paragraph{Forecasting Models}

Recall $Y_{\ell,t}$ denotes the 7-day trailing average of COVID-19 case
incidence rates in location $\ell$ and at time $t$.  Separately for each
$a=7,\ldots,21$, to predict $Y_{\ell,t+a}$ for ahead value $a$, we consider a
simple probabilistic forecasting model of the form:     
\begin{equation}
\label{eq:forecast_ar}
\Quantile_\tau(Y_{\ell,t+a} \ | \ Y_{\ell,s},\ s \leq t)   
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j}.  
\end{equation}
This model uses current case rates, and the case rates 7 and 14 days ago, in
order to predict (the quantiles of) case rates in the future.  We consider a
total of 7 quantile levels (chosen in accordance with the county-level quantile
levels suggested by the COVID-19 Forecast Hub),      
\begin{equation}
\label{eq:quantile_levels}
\tau \in \{0.025,\ 0.1,\ 0.25,\ 0.5,\ 0.75,\ 0.9,\ 0.975 \}.
\end{equation}
We fit \eqref{eq:forecast_ar} using \textit{quantile regression}
\cite{Koenker:1978, Koenker:2005, Koenker:2006} separately for each $\tau$, 
using data from all 306 HRRs, and within each HRR, using the most recent 21 days
of training data.  This gives us 6,426 training samples for each quantile 
regression problem.

In addition to this pure autoregressive model, we also consider five
probabilistic forecasting models of the form:  
\begin{multline}
\label{eq:forecast_ar_x}
\Quantile_\tau(Y_{\ell,t+a} \ | \ Y_{\ell,s},\ X_{\ell,s},\ s \leq t) 
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y_{\ell,t-7j} + 
\sum_{j=0}^2 \gamma^{a,\tau}_j X_{\ell,t-7j},
\end{multline}
where $X_{\ell,t}$ denotes any one of the five auxiliary indicators---CHNG-CLI, 
CHNG-COVID, CTIS-CLI-in-community, DV-CLI, or Google-AA---at location $\ell$ and
time $t$. Note that we apply the same lags (current value, along with the values
7 and 14 days ago) for the auxiliary indicators as we do for the case
rates. Training then proceeds just as before: we use the same 7 quantile levels
in \eqref{eq:quantile_levels}, and fit quantile regression separately for each 
level $\tau$, using data from all 306 HRRs and a trailing window of 21 days of
training data.

At prediction time, in order to avoid crossing violations (that is, for two
levels $\tau' > \tau$, the predicted quantile at level $\tau$ exceeds the
predicted quantile at level $\tau'$), we apply a simple post hoc sorting.  See 
Figure~\ref{fig:trajectory} for an example forecast.

\begin{figure}[t]
\includegraphics[width=\columnwidth]{fig/trajectory.pdf}
\caption{Forecast for the HRR containing New York City from an autoregressive
  model made on October 15 (vertical line).  The fan displays 50\%, 80\% and
  95\% intervals while the orange curve shows the median forecast. The black
  curve shows ``finalized'' data, as reported in May 2021.}   
\label{fig:trajectory}
\end{figure}

\paragraph{Hotspot Prediction Models}

Define the binary indicator:
$$
Z_{\ell,t} = \indicator{Y^\Delta_{\ell,t} \geq 0.25},
$$
where we use the notation \smash{$Y^\Delta_{\ell,t} = (Y_{\ell,t} - Y_{\ell,
    t-7})/(Y_{\ell,t-7})$}. In other words, $Z_{\ell,t}=1$ if the number of
newly reported cases over the past 7 days has increased by at least 25\%
compared to the preceding week.  When this occurs, we say location $\ell$ is a
\textit{hotspot} at time $t$.  Empirically, this rule labels about 27\% of
location-time pairs as hotspots, during the prediction period (June 16--December
31, 2020).

We treat hotspot prediction as a binary classification problem and use a setup 
altogether quite similar to the forecasting setup described previously.
Separately for each $a=7,\ldots,21$, to predict $Z_{\ell,t+a}$, we consider a
simple logistic model:
\begin{equation}
\label{eq:hotspot_ar}
\logit \big( \P(Z_{\ell,t+a} = 1 \ | \ Y_{\ell,s},\ s \leq t) \big) 
= \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j},
\end{equation}
where $\logit(p) = \log(p/(1-p))$, the log-odds of $p$.

In addition to this pure autoregressive model, we also consider five
logistic models of the form:  
\begin{multline}
\label{eq:hotspot_ar_x}
\logit \big( \P(Z_{\ell,t+a} = 1 \ | \ Y_{\ell,s},\ X_{\ell,s},\ s \leq t) \big)
= \\ \alpha^{a,\tau} + \sum_{j=0}^2 \beta^{a,\tau}_j Y^\Delta_{\ell,t-7j} +  
\sum_{j=0}^2 \gamma^{a,\tau}_j X^\Delta_{\ell,t-7j},
\end{multline}
where we use \smash{$X^\Delta_{\ell,t} = (X_{\ell,t} - X_{\ell,
    t-7})/(X_{\ell,t-7})$}, and again $X_{\ell,t}$ stands for any of the five  
auxiliary indicators at location $\ell$ and time $t$.  We fit the above models,
\eqref{eq:hotspot_ar}, \eqref{eq:hotspot_ar_x}, using logistic regression,   
pooling all 306 HRRs and using a 21-day trailing window for the training data.

An important detail is that in hotspot prediction we remove all data from
training and evaluation where, on average, fewer than 30 cases (this refers to a
count, not a rate) are observed over the prior 7 days. This avoids having to
make arbitrary calls for a hotspot (or lack thereof) based on small counts.

\subsection{Evaluation Metrics}

For forecasting, we evaluate the probabilistic forecasts produced by the
quantile models in \eqref{eq:forecast_ar} and \eqref{eq:forecast_ar_x} using
\textit{weighted interval score} (WIS), a quantile-based scoring rule; see,
e.g., \cite{Gneiting:2007}.  WIS is a proper score, which means that its
expectation is minimized by the population quantiles of the target variable.
The use of WIS in COVID-19 forecast scoring was proposed by \cite{Bracher:2021}; 
WIS is also the main evaluation metric used in the COVID-19 Forecast Hub. More
broadly, the specific form of WIS used here is a standard metric in the
forecasting community for evaluating quantile-based probabilistic forecasts,
just as mean squared forecast error is standard for point forecasts.

WIS is typically defined for quantile-based forecasts where the quantile levels
are symmetric around 0.5.  This is the case for our choice in
\eqref{eq:quantile_levels}.  Let $F$ be a forecaster comprised of predicted
quantiles $q_\tau$ parametrized by a quantile level $\tau$.  In the case of
symmetric quantile levels, this is equivalent to a collection of central
prediction intervals $(\ell_\alpha, u_\alpha)$, parametrized by an exclusion
probability $\alpha$. The WIS of the forecaster $F$, evaluated at the target
variable $Y$, is defined by:
\begin{equation}
\label{eq:wis_intervals}
\WIS(F,Y) = \sum_\alpha \Big\{\alpha(u_\alpha - \ell_\alpha) + 2 \cdot
\dist(Y, [\ell_\alpha, u_\alpha])\Big\},  
\end{equation}
where $\dist(a, S)$ is the distance between a point $a$ and set $S$ (the
smallest distance between $a$ and an element of $S$).  Note that, corresponding  
to \eqref{eq:quantile_levels}, the exclusion probabilities are $\alpha \in
\{0.05,\ 0.2,\ 0.5,\ 1\}$, resulting in 4 terms in the above sum.  By 
straightforward algebra, it is not hard to see WIS has an alternative 
representation in terms of the predicted quantiles themselves:
\begin{equation}
\label{eq:wis_quantiles}
\WIS(F,Y) = 2 \sum_\tau \phi_\tau(Y - q_\tau), 
\end{equation}
where $\phi_\tau(x) = \tau |x|$ for $x \geq 0$ and $\phi_\tau(x) = (1-\tau)
|x|$ for $x<0$, which is often called the ``tilted absolute'' loss.  While
\eqref{eq:wis_quantiles} is more general (it can accomodate asymmetric quantile
levels), the first form in \eqref{eq:wis_intervals} is typically preferred in
presentation, as the score nicely decouples into a ``sharpness'' component
(first term in each summand) and an ``under/overprediction'' component (second
term in each summand).  But the second form given in \eqref{eq:wis_quantiles} is 
especially noteworthy in our current study because it reveals WIS to be the same
as the quantile regression loss that we use to train our forecasting models.
%(i.e., we fit by optimizing WIS averaged over the training data).

For hotspot prediction, we evaluate the probabilistic classifiers produced by
the logistic models in \eqref{eq:hotspot_ar} and \eqref{eq:hotspot_ar_x} using
the area under the curve (AUC) of their true positive versus false positive rate 
curve (which is traced out by varying the discrimination threshold).

The primary aggregation scheme that we will use in model evaluation and
comparisons will be to average WIS per forecaster at ahead value $a$ over all
forecast dates $t$ and locations $\ell$; and similarly, to compute AUC per
classifier at ahead value $a$ over all forecast dates $t$ and locations
$\ell$.

\subsection{Other Considerations}

\paragraph{Missing Data Imputation}

Over the prediction period, all auxiliary indicators are available (in the
proper vintage sense) for all locations and prediction times, except for the
Google-AA signal, which is only observed for an average of 105 (of 306) HRRs.
Such missingness occurs because the COVID-19 search trends data is
constructed using differential privacy methods \cite{Bavadekar:2020}, and a 
missing signal value means that the level of noise added in the differential
privacy mechanism is high compared to the underyling search count.  In other
words, values of the Google-AA signal are clearly \textit{not} missing at
random.  It seems most appropriate to impute missing values by zero, and this is
what we do in our analysis.

\paragraph{Backfill and Nowcasting}

As described previously, the auxiliary indicators defined in terms of medical
claims (CHNG-CLI, CHNG-COVID, and DV-CLI) undergo a significant and systematic
pattern of revision, or ``backfill'', after their initial publication.  Given
their somewhat statistically-regular backfill profiles, it would be reasonable
to attempt to estimate their finalized values based on vintage data---a problem
we refer to as \textit{nowcasting}---as a pre-processing step before using them
as features in the models in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}. 
Nowcasting is itself a highly nontrivial modeling problem, and we do not attempt
it in this paper (it is a topic of ongoing work in our research group), but we
note that nowcasting would likely improve the performance of the models
involving claims-based signals in particular.

\paragraph{Spatial Heterogeneity}

Some signals have a significant amount of spatial heterogeneity, by which we
mean their values across different geographic locations are not comparable.
This is the case for the Google-AA signal (due to the way in which the
Google search trends time series is self-normalized, see
\cite{GoogleSymptoms}) and the claims-based signals (due to market-share 
differences, and/or differences in health-seeking behavior).  Such spatial
heterogeneity likely hurts the performance of the predictive models that rely on
these signals, because we train the models on data pooled over all locations.
In the current paper, we do not attempt to address this issue, and we simply
note that location-specific effects (or pre-processing to remove spatial bias)
would likely improve the performance of the models involving Google-AA and the
claims-based indicators.

\section{Results}

Here, and in what follows, we will use ``AR'' to refer to the pure 
autoregressive model both in forecasting, \eqref{eq:forecast_ar}, and in hotspot
prediction, \eqref{eq:hotspot_ar} (the reference to the prediction task should
always be clear from the context). We will also use the name of an auxiliary 
indicator---namely ``CHNG-CLI'',  ``CHNG-COVID'', ``CTIS-CLI-in-community'',
``DV-CLI'', or ``Google-AA''---interchangeably with the model in forecasting, 
\eqref{eq:forecast_ar_x}, or hotspot prediction, \eqref{eq:hotspot_ar_x}, that
uses this same indicator as a feature (the meaning should be clear from
the context).  So, for example, the CHNG-CLI model in forecasting is the one in  
\eqref{eq:forecast_ar_x} that sets $X_{\ell,t}$ to be the value of the CHNG-CLI
indicator at location $\ell$ and time $t$.  Finally, we use the term ``indicator
model'' to refer to any one of the ten models of the form
\eqref{eq:forecast_ar_x} or \eqref{eq:hotspot_ar_x} (five from each of the
forecasting and hotspot prediction tasks).

Below is summary of the high-level conclusions.

\begin{itemize}
\item Stratifying predictions by the ahead value ($a=7,\ldots,21$), and
  aggregating results over the prediction period (early June through end of
  December 2020), we find that each of the indicator models generally gives a
  boost in predictive accuracy over the AR model, in both the forecasting and
  hotspot prediction tasks.  The gains in accuracy generally attenuate as the
  ahead value grows.

\item In the same aggregate view, CHNG-COVID and DV-CLI offer the biggest 
  gains in both forecasting and hotspot prediction.  CHNG-CLI is
  inconsistent: it provides a large gain in hotspot prediction, but little gain  
  in forecasting (it seems to be hurt by a notable lack of robustness, due
  to backfill).  CTIS-CLI-in-community and Google-AA each provide decent 
  gains in forecasting and hotspot prediction.  The former's performance in 
  forecasting is notable in that it clearly improves on AR even at the largest
  ahead values.

\item In a more detailed analysis of forecasting performance, we find that the
  indicator models tend to be better than AR when case rates are flat or
  decreasing (most notable in CHNG-COVID and CTIS-CLI-in-community), but can be
  worse than AR when case rates are increasing (this is most notable in CHNG-CLI
  and DV-CLI). More rarely does an indicator model tend to beat AR when case
  rates are increasing, but there appears to be evidence of this for the
  Google-AA model.   
  % In an increasing period, when an indicator model is worse than AR, it tends
  % to underpredict relative to AR (its median forecast is lower).  In a flat
  % or decreasing period, when an indicator model is better than AR, AR tends
  % to overpredict relative to it (the median AR forecast is higher).

\item In this same analysis, when an indicator model performs better than AR 
  in a decreasing period, this tends to co-occur with instances in which the 
  indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
  given location, its behavior mimics that of future case rates).  On
  the other hand, if an indicator model does better in periods of
  increase, or worse in periods of increase or decrease, its performance is
  not as related to leadingness.   
\end{itemize}

Finally, to quantify the importance of training and making predictions using
proper vintage data, we ran a parallel set of forecasting and hotspot prediction
experiments using finalized data. The results, given in the Supplementary
Materials, show that training and making predictions on finalized data can
result in overly optimistic estimates of true test-time performance (up to 10\%
better in terms of average WIS or AUC). Furthermore, since indicators can have
greatly different backfill profiles, the use of finalized data in retrospective
evaluations changes the relative ranking of models.  For example, CHNG-CLI and
DV-CLI, when trained on finalized data, perform very similarly in forecasting.
This makes sense since they are both claims-based indicators that are supposedly
measuring the same thing.  However, DV-CLI outperforms CHNG-CLI on vintage data,
reflecting its has a less severe backfill profile.

The Supplement provides a number of other additional analyses; for example, we
examine two assumption-lean methods for assessing the statistical significance
of our main results.

Code to reproduce all results can be found at
\url{https://github.com/cmu-delphi/covidcast-pnas/tree/main/forecast/code}.

\subsection{Aggregate Results by Ahead Value}

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.95\textwidth]{fig/fcast-hot-combo-1.pdf}
  \caption{Main results for both tasks. Left: average WIS for each forecast
    model, over all forecast dates and all HRRs, divided by the average WIS
    achieved by a baseline model (a probabilistic version of the flat-line
    forecaster).  Right: area under the curve for each hotspot prediction model,
    calculated over all prediction dates and all HRRs.  Here and in all figures
    we abbreviate CTIS-CLI-in-community by CTIS-CLIIC.  
  }  
  \label{fig:forecast}
\end{figure*}

Figure~\ref{fig:forecast} (left panel) displays evaluation results for
forecasting, stratified by ahead value and averaged over all HRRs and forecast
dates.  Shown is the average WIS for each forecast model divided by that from a
baseline model. The baseline model is a flat-line forecaster which forms its
median forecast by using the most recent value $Y_{\ell,t}$ for all aheads
$Y_{\ell,t+a}$, with predicted quantiles defined by the empirical distribution
of the residuals from this median forecast over recent history. This is the same
baseline model as in the COVID-19 Forecast Hub. Here, we use the baseline model
in order to scale mean WIS to put it on an interpretable, unitless
scale.  In the figure, we can see that all curves are below 1, which means
(smaller WIS is better) that all of the models, including AR, outperform the
baseline on average over the forecasting period.  On the other hand, the models
deliver at best an improvement of about 20\% in average WIS over the baseline
model, with this gap narrowing to about 10\% at the largest ahead values,
illustrating the difficulty of the forecasting problem.

We can also see from the figure that CHNG-COVID and DV-CLI offer the biggest
gains over AR at small ahead values, followed by CTIS-CLI-in-community and
Google-AA, with the former providing the biggest gains at large ahead values.
The CHNG-CLI model performs basically the same as AR.  This is likely due to the
fact that CHNG-CLI suffers from volatility due to backfill.  The evidence for
this explanation is twofold: (1) the CHNG-CLI model benefits from a more robust
method of aggregating WIS (geometric mean; shown in the Supplementary
Materials); and (2) when we train and make predictions on finalized data, it
handily beats AR, on par with the best-performing models (also shown in the
Supplement).

Figure~\ref{fig:forecast} (right panel) displays the results for hotspot
prediction, again stratified by ahead value and averaged over all HRRs and
prediction dates.  We can see many similarities to the forecasting results
(note: now larger AUC is better). For example, CHNG-COVID and DV-CLI offer the
biggest improvement over AR, and all models, including AR, degrade in
performance towards the baseline (in this context, a classifier based on random
guessing, which achieves an AUC of 0.5) as the ahead values grow, illustrating
the difficulty of the hotspot prediction problem. A clear difference, however,
is that the CHNG-CLI model performs quite well in hotspot prediction, close to
the best-performing indicator models for many of the ahead values. This may be
because volatility in the CHNG-CLI indicator plays less of a role in the
associated logistic model's predicted probabilities (in general, a sigmoid
function can absorb a lot of the variability in its input).

\subsection{Implicit Regularization Hypothesis}

One might ask if the benefits observed in forecasting and hotspot
prediction have anything to do with the actual auxiliary indicators
themselves. A plausible alternative explanation is that the indicators are
just providing \textit{implicit regularization} on top of the basic AR model,
in the same way any noise variable might, if we were to use it to create lagged
features in \eqref{eq:forecast_ar_x} and \eqref{eq:hotspot_ar_x}.

To test this hypothesis, we reran all of the prediction experiments but with
$X_{\ell,t}$ in each indicator model replaced by suitable random noise
(bootstrap samples from a signal's history).  The results, shown and explained
more precisely in the Supplement, are vastly different (i.e., worse) than the
original set of results.  In both forecasting and hotspot prediction, the
``fake'' indicator models offer essentially no improvement over the pure AR
model, which---informally speaking---strongly rejects the implicit
regularization hypothesis.

On the topic of regularization, it is also worth noting that the use of 
$\ell_1$ regularization (tuned using cross-validation) in fitting any of the
models in \eqref{eq:forecast_ar}, \eqref{eq:forecast_ar_x},
\eqref{eq:hotspot_ar}, or  \eqref{eq:hotspot_ar_x} did not generally improve
their performance (experiments not shown).  This is likely due to the fact that
the number of training samples is large compared to the number of features
(6,426 training samples and only 3--6 features).  
% DJM: and that proper CV requires longer training windows, exacerbating the
% nonstationarity, a scenario in which its behavior is poorly understood.

\subsection{Evaluation in Up, Down, and Flat Periods}

\begin{figure*}[t]
  \includegraphics[width=\textwidth]{fig/upswing-histogram-1.pdf}
  \caption{Histogram of the difference in WIS for the AR model and that for each
    indicator model, stratified by up, down, or flat period, measured in terms
    of case trends. Note that \textit{larger} differences here are better for
    each indicator model.  The y-axis is on the log scale to emphasize tail 
    behavior.}   
  \label{fig:up_down_flat}
\end{figure*}

The course of the pandemic has played out quite differently across space and
time. Aggregating case rates nationally shows three pronounced waves, but the
behavior is more nuanced at the HRR level.  Figure \ref{fig:trajectory} is a
single example of a forecast in a period of relatively flat case trends, as New
York City enters what would become its second wave. The AR forecaster's 50\%
prediction interval contains this upswing, but its forecasted median is clearly
below the finalized case data.  Unfortunately, this behavior is fairly typical
of all forecasters: during upswings, the forecasted median tends to fall below 
the target, while the reverse is true during downswings.

Figure~\ref{fig:up_down_flat} shows histograms of the differences in WIS of the 
AR model and each indicator model, where we stratify these differences by
whether the target occurs during a period of increasing cases rates (up),
decreasing case rates (down), or flat case rates (flat). To define the
increasing period, we use the same definition we used for the hotspot task in
Table \ref{tab:analysis_tasks}.  Therefore, all hotspots are labeled ``up'',
while all non-hotspots are either ``flat'' or ``down''.  For the ``down''
scenario, we simply use the opposite of the hotspot definition: $Y_{\ell,t}$
decreases by more than 20\% relative to the preceding week.

While the performance of all models, including AR, generally degrades in up
periods, different models exhibit different and interesting patterns. CHNG-CLI,
CHNG-COVID, Google-AA, and especially CTIS-CLI-in-community have large right
tails (showing improvements over AR) during the down periods.  Google-AA and
CTIS-CLI-in-community have large right tails during the flat periods. CHNG-CLI
and DV-CLI have large left tails (poor forecasts relative to AR) in flat and up
periods.  Google-AA is the only model that outperforms the AR model, on average,
in up periods.  Overall, the indicators seem to help more during flat or down
periods than during up periods, with the exception of Google-AA.

The Supplement pursues this analysis further.  For example, we examine
classification accuracy and log-likelihood for the hotspot task and find a
similar phenomenon: the indicators considerably improve accuracy or
log-likelihood during flat or down periods, with more mixed behavior during up
periods when CHNG-CLI, CHNG-COVID, and DV-CLI, in particular, lead to decreased  
performance.

\subsection{Effects of Leading or Lagging Behavior}
\label{sec:leading-lagging}

\begin{figure}[t]
  \includegraphics[width=\columnwidth]{fig/leading-only-1.pdf}
  \caption{Correlation of the difference in WIS with the ``leadingness'' of the
    indicator at the target date, stratified by up, down, or flat period.} 
  \label{fig:leading}
\end{figure}

As described in the methods section, each of the indicators we examine could be
said to measure aspects of disease progression that would precede a positive
test. That is, we imagine that these signals should ``lead'' cases. It is
entirely reasonable to imagine that, prior to an increase of confirmed COVID-19
tests reported by a public health authority in a particular location, we would
see an increase in medical insurance claims for COVID-related outpatient
visits. However, it may well be the case that such behavior is different
during different periods. In fact, we find empirically that the ``leadingness''
of an indicator (degree to which it leads case activity) tends to be more
pronounced in down or flat periods than in up periods, a plausible explanation 
for the decreased performance in up periods noted above.

In the supplement, we define a quantitative score to measure the 
leadingness of an indicator, at any time $t$ and any location $\ell$, based on 
cross correlations to case rates over a short time window around $t$.
The higher this score, the greater it ``leads'' case activity. This analysis is
closely related to Granger causality \cite{engle1987a} and draws on a large
body of prior work that measures leadingness in economic time-series, e.g., 
\cite{mitchell1938a, koopmans1947a, shiskin1968a, granger1969a, yeats1972a, 
  sargent1977a, auerbach1982a, koch1988a, StockWatson1989, diebold1989a,
  lahiri1991a, emerson1996a, hamilton1996a}. Figure \ref{fig:leading} displays
correlations between the leadingness score of an indicator and the WIS
difference (AR model minus an indicator model), stratified by whether the target
is classified as up, down, or flat.  One would naturally expect that the WIS
difference would be positively correlated with leadingness.  Somewhat
surprisingly, this relationship turns out to be strongest in down periods and
weakest in up periods.  In fact, it is very nearly the case that for each
indicator, the strength of correlations only decreases as we move from down to
flat to up periods.  In the supplement, we extend this analysis by studying
analogous ``laggingness'' scores, but we do not find as clear patterns.

\section{Discussion}

Can auxiliary indicators improve COVID-19 forecasting and hotspot prediction
models?  Our answer, based on analyzing five auxiliary indicators from the
COVIDcast API (defined using from medical insurance claims, internet-based
surveys, and internet search trends) is undoubtedly ``yes''. However, there are
levels of nuance to such an answer that must be explained.  None of the
indicators that we have investigated are transformative, rendering the
prediction problem easy when it was once hard (in the absence of auxiliary
information). Rather, the gains in accuracy from the indicator models (over an 
autoregressive model based only on past case rates) appear to be nontrivial, and
consistent across modes of analysis, but modest.  In forecasting, the indicator
models are found to be most useful in periods in which case rates are flat or
trending down, rather than periods in which case rates are trending up (as one
might hope to see is the benefit being provided by a hypothetical ``leading 
indicator'').

As described previously, it is likely that we could improve the indicator models
by using location-specific effects, as well as using nowcasting techniques to
estimate finalized indicator values before we use them as features (to account
for backfill in the claims-based signals in particular).  Beyond this, it is
certainly possible that more sophisticated models for forecasting or hotspot
prediction would lead to different results, and possibly even different
insights. Natural directions to explore include using multiple indicators in a
single model, allowing for interaction terms, and leveraging HRR demographics or
mobility patterns.  That said, we are doubtful that more sophisticated modeling
techniques would change the ``topline'' conclusion---that auxiliary indicators
can provide nontrivial, consistent, but modest gains in forecasting and hotspot
prediction. Whether a more sophisticated model would be able to leverage the
indicators in such a way as to change some of the finer conclusions (e.g., by 
offering clear improvements in periods in which cases are trending up) is less 
clear to us.

We reiterate the importance of using vintage data for rigorous backtesting. Data
sources that are relevant to public health surveillance are often subject to
revision, sometimes regularly (as in medical claims data) and sometimes
unpredictably (such as COVID-19 case reports).  When analyzing models that are
designed to predict future events, if we train these models and make predictions
using finalized data, then we are missing a big part of the story.  Not only
will our sense of accuracy be unrealistic, but certain models may degrade by a
greater or lesser extent when they are forced to work with vintage data, so 
backtesting with finalized data may lead us to make modeling decisions that are
suboptimal for true test-time performance.

In this paper, we have chosen to consider only very simple forecasting models,
while devoting most of our effort to accounting for as much of the complexity of
the underlying data and evaluation as possible. In fact, our paper is not only
about providing rigorous answers to questions about model comparisons in
COVID-19 forecasting and hotspot prediction, but also about demonstrating how 
one might go about answering such questions in general. We hope that others will
leverage our work, and build on it, to guide advances on the frontier of
predictive modeling for epidemics and pandemics.

\acknow{We thank Matthew Biggerstaff, Logan Brooks, Johannes Bracher, Michael 
  Johansson, Evan Ray, Nicholas Reich, and Roni Rosenfeld for several
  enlightening conversations about forecasting, scoring, and evaluation.  
  This material is based on work supported by gifts from Facebook, Google.org,
  the McCune Foundation, and Optum; Centers for Disease Control and Prevention
  (CDC) grant U01IP001121; the Canadian Statistical Sciences Institute; National 
  Sciences and Engineering Research Council of Canada (NSERC) grant
  RGPIN-2021-02618; and National Science Foundation Graduate Research
  Fellowship Program (NSF GRFP) award DGE1745016.}

\showacknow{} % Display the acknowledgments section

% Bibliography
\bibliography{../../common/covidcast.bib}

\end{document}