### 2021-11-04 cover/cover.Rmd
Add Astley paragraph

---
title: Playground for PNAS cover graphic
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.path = "fig/", cache = TRUE,
                      autodep = TRUE, cache.comments = TRUE,
                      message = FALSE, warning = FALSE)

library(covidcast)
library(ggplot2)
library(dplyr)
library(purrr)
```

```{r get-data, eval = FALSE}
sources <- c("usa-facts",
             "doctor-visits",
             "fb-survey",
             "google-symptoms")

signals <- c("confirmed_7dav_incidence_prop",
             "smoothed_adj_cli",
             "smoothed_whh_cmnty_cli",
             "sum_anosmia_ageusia_smoothed_search")

start <- as.Date("2020-04-15")
end <- as.Date("2021-10-22")
dates <- seq.Date(start, end, by = "1 day")
geos <- tolower(state.abb)

x <- covidcast_signals(
  sources, signals,
  start, end,
  geo_type = "state", 
  geo_values = geos) %>% 
  map(function(df) df %>%
        select(data_source, signal, geo_value, time_value, value))
```

```{r save-all, eval = FALSE}
save(list = ls(), file = "all.rda")
```

```{r load-all}
load(file = "all.rda")
```

```{r load-all}
# Inspect missingness: google has lots (privacy threshold)
do.call(rbind, x) %>%
  filter(geo_value %in% tolower(state.abb)) %>%
  group_by(signal, geo_value) %>%
  summarize(nas = length(dates) - n()) %>%
  ungroup() %>% group_by(signal) %>%
  filter(nas > 0)

y <- tibble(geo_value = rep(geos, each = length(dates)),
            time_value = rep(dates, length(geos)))

x[[4]] %>%
  full_join(y, by = c("geo_value", "time_value")) %>%
  group_by(time_value) %>%
  summarize(nas = sum(is.na(value))) %>% 
  filter(nas > 40)
```

```{r trend-plots}
names <- c("Reported COVID cases",
           "COVID symptoms\nin medical claims",
           "Self-reported COVID\nsymptoms in surveys",
           "Searches for COVID\nsymptoms on Google")
colors <- c("#FFFFFF", "#E69F00", "#56B4E9", "#CC79A7")

g <- vector(mode = "list", length = length(x))
delta <- 1 / length(geos) * 0.1
alpha <- 0.35
line_width <- 0.3
label_size <- 3
xlim <- c(start - 10, end)
ylim <- c(-0.05, 1.05 + length(geos) * delta)
xp <- xlim[1] - 8
yp <- mean(ylim)

for (i in 1:length(x)) {
  g[[i]] <- x[[i]] %>%
    full_join(y, by = c("geo_value", "time_value")) %>%
    mutate(geo_num = as.integer(as.factor(geo_value))) %>%
    group_by(geo_value) %>%
    mutate(value = (value - Min(value)) / (Max(value) - Min(value)) +
             delta * (geo_num - 1)) %>%
    ungroup() %>%
    ggplot(aes(x = time_value, y = value, group = geo_value)) +
    labs(y = names[i]) + coord_cartesian(xlim = xlim, ylim = ylim) +
    geom_line(alpha = 0.5, color = colors[i], size = line_width) +
    theme_void() +
    annotate("text", x = xp, y = yp, label = names[i],
             color = colors[i], angle = 90, size = label_size) +
    theme(panel.background = element_rect(fill = "black"))
}

gg <- do.call(gridExtra::arrangeGrob, list(grobs = g, ncol = 1))
ggsave(file = "cover-graphic.pdf", gg, height = 22.5, width = 21, units = "cm")
```