### 2021-11-05 forecast/paper/pnas-suppl-template.tex
add proper DOIs to forecast supplement

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}

\input{../../common/prelim.tex}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

$for(header-includes)$
$header-includes$
$endfor$

\title{Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot  
  Prediction?}

\author{Daniel J. McDonald, Jacob Bien, Alden Green, Addison J. Hu, Nat DeFries,
  Sangwon Hyun, Natalia L. Oliveira, James Sharpnack, Jingjing Tang, Robert
  Tibshirani, Valerie Ventura, Larry Wasserman, and Ryan J. Tibshirani} 
\correspondingauthor{Daniel J. McDonald.\\E-mail: daniel@stat.ubc.ca}

\begin{document}

\maketitle

\SItext

$body$

\FloatBarrier

\dataset{Code}{Code for reproducibility. {\tt R} script files and small
  datasets. 
  Persistent DOI: \url{https://doi.org/10.5281/zenodo.5639567}}

\dataset{Data}{Large intermediate data sets. Archived {\tt .RDS} (R objects)
  files 
  containing all evaluations for forecasting and hotspots using vintage and
  finalized data. Can be used to produce the graphics and conclusions in the
  manuscript without rerunning the entire pipeline. 
  Persistent DOI: \url{https://doi.org/10.5683/SP3/UW4VTC}}

\bibliography{$for(bibliography)$$bibliography$$sep$,$endfor$}

\end{document}