### 2021-11-08 forecast/code/deprecated/build_map.Rmd
update forecast readme

---
title: Build the mapping from HRRs to states
---

```{r, message = FALSE}
library(covidcast)
library(dplyr)
library(tidyr)

start_day <- "2020-03-01"
end_day <- "2021-10-15"

y <- covidcast_signal(data_source = "jhu-csse", 
                      signal = "confirmed_incidence_prop",
                      start_day = start_day, 
                      end_day = end_day,
                      geo_type = "state") %>%
  select(geo_value, time_value, value) %>%
  pivot_wider(names_from = geo_value)

dim(y)
head(y)

x <- covidcast_signal(data_source = "jhu-csse", 
                      signal = "confirmed_incidence_prop",
                      start_day = start_day, 
                      end_day = end_day,
                      geo_type = "hrr") %>%
  select(geo_value, time_value, value) %>%
  pivot_wider(names_from = geo_value)

dim(x)
head(x)

# Check we have the same time values
all(y$time_value == x$time_value)

# Convert to matrices, compute regression coefficients
library(MASS) # Using ginv() because I got singularity errors with solve()
x_mat = as.matrix(x[,-1])
y_mat = as.matrix(y[,-1])
b = t(cov(y_mat, x_mat) %*% ginv(cov(x_mat)))

# Sanity check: check against first 4 individual regressions
par(mfrow = c(2, 2))
for (i in 1:4) plot(b[,i], coef(lm(y_mat[,i] ~ x_mat + 0)))

# Sanity check: look at reconstruction error by state
err = colMeans((y_mat - x_mat %*% b)^2)
barplot(err)
```